---
title: Early life history respirometry (photosynthesis & respiration) analysis and plotting
author: "AS Huffmyer"
date: '2020'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
---
This script analyzes and plots data for Early Life History Energetics 2020 respirometry data. Plots are displayed in Plotting section. Results are provided in Analysis section along with summaries of potential implications. See Conclusions section for general overview of findings.    

## **Setup**  

Set up workspace, set options, and load required packages.    
```{r, echo=TRUE, show=FALSE}
rm(list=ls(all=TRUE)) 
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("scales" %in% rownames(installed.packages()) == 'FALSE') install.packages('scales') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("effects" %in% rownames(installed.packages()) == 'FALSE') install.packages('effects') 

#load packages
library("ggplot2")
library("tidyverse")
library('car')
library('lme4')
library('lmerTest')
library('scales')
library('cowplot')
library('effects')
```


## **Data visualization and manipulation**  

Load data from LoLinR.   
```{r, warning=FALSE, message=FALSE}
PRdata<-read.csv("../../Ahya2020/Output/Respiration/oxygen_P_R_calc.csv") #load data
```

Separate project specific data.  
```{r, warning=FALSE, message=FALSE}
#remove all rows of wells that did not have samples or blanks
PRdata<-PRdata[!is.na(PRdata$Type),]

#format columns
PRdata$dpf<-as.factor(PRdata$dpf)

```

Calculate a P:R ratio using gross photosynthesis.  
```{r, warning=FALSE, message=FALSE}
PRdata$ratio<-(PRdata$GP.nmol.org.min)/abs(PRdata$R.nmol.org.min) #calculate ratio with absolute values
#remove outliers detected by values of P:R ratio data
```

## **Plotting: Larvae**    

Generate plots by days post fertilization.  

### *Respiration*  

Examine data.  

```{r}
hist(PRdata$R.nmol.org.min)
```

Generate dot plot.  

```{r, results=TRUE, warning=FALSE, message=FALSE}
Rplot1<-ggplot(data=PRdata, aes(x=dpf, y=R.nmol.org.min, colour=Stage, group=interaction(Stage, dpf))) +
  geom_boxplot(position=position_dodge(0.9), lwd=1)+
  geom_jitter(shape=16, position=position_dodge(0.9), size=3)+
  geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
  scale_color_brewer(palette = "Dark2", breaks=c("Gastrula", "Larvae"))+
  scale_y_continuous(limits=c(-0.1, 0.1), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
  xlab("Days post fertilization")+
  ylab(expression(bold(paste("R (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
  theme_classic() +
  theme(text = element_text(size = 18, color="black"),
        axis.text = element_text(size = 18, color="black"), 
        legend.position = "top",
        axis.title = element_text(size = 18, color="black", face="bold"), 
        legend.title=element_blank(), 
        legend.text = element_text(size=18), 
        plot.margin=unit(c(1,1,1,1), "cm"), 
        axis.title.y = element_text(margin = margin(t = 0, r = 1, b = 0, l = 1)), 
        axis.title.x = element_text(margin = margin(t = 3, r = 0, b = 0, l = 0)));Rplot1

Rplot2<-Rplot1+theme(legend.position="none")

ggsave("../../Ahya2020/Figures/Respiration/ahya_R.pdf", plot=Rplot1, height=8, width=7, units = c("in"), dpi=300) #output figure
```

Generate mean table  

```{r, results=TRUE}
meanR <- plyr::ddply(PRdata, c("Stage", "dpf"), summarise,
                   N    = length(R.nmol.org.min[!is.na(R.nmol.org.min)]),
                   mean = mean(R.nmol.org.min, na.rm=TRUE),
                   sd   = sd(R.nmol.org.min, na.rm=TRUE),
                   se   = sd / sqrt(N)
); meanR
``` 

### *Net Photosynthesis*   

Examine data.  

```{r}
hist(PRdata$P.nmol.org.min)
```

Generate plots.  

```{r, results=TRUE, warning=FALSE, message=FALSE}
Pplot1<-ggplot(data=PRdata, aes(x=dpf, y=P.nmol.org.min, colour=Stage, group=interaction(Stage, dpf))) +
  geom_boxplot(position=position_dodge(0.9), lwd=1)+
  geom_jitter(shape=16, position=position_dodge(0.9), size=3)+
  geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
  scale_color_brewer(palette = "Dark2", breaks=c("Gastrula", "Larvae"))+
  scale_y_continuous(limits=c(-0.1, 0.1), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
  xlab("Days post fertilization")+
  ylab(expression(bold(paste("Net P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
  theme_classic() +
  theme(text = element_text(size = 18, color="black"),
        axis.text = element_text(size = 18, color="black"), 
        legend.position = "right",
        axis.title = element_text(size = 18, color="black", face="bold"), 
        legend.title=element_blank(), 
        legend.text = element_text(size=18), 
        plot.margin=unit(c(1,1,1,1), "cm"), 
        axis.title.y = element_text(margin = margin(t = 0, r = 1, b = 0, l = 1)), 
        axis.title.x = element_text(margin = margin(t = 3, r = 0, b = 0, l = 0))); Pplot1

Pplot2<-Pplot1+theme(legend.position="none")

ggsave("../../Ahya2020/Figures/Respiration/ahya_P.pdf", plot=Pplot2, height=8, width=7, units = c("in"), dpi=300) #output figure
```
 
Generate mean table.   

```{r, results=TRUE}
meanP <- plyr::ddply(PRdata, c("Stage", "dpf"), summarise,
                   N    = length(P.nmol.org.min[!is.na(P.nmol.org.min)]),
                   mean = mean(P.nmol.org.min, na.rm=TRUE),
                   sd   = sd(P.nmol.org.min, na.rm=TRUE),
                   se   = sd / sqrt(N)
); meanP
```

### *Gross Photosynthesis*   

Examine data.  

```{r}
hist(PRdata$GP.nmol.org.min)
```

```{r, results=TRUE, warning=FALSE, message=FALSE}
GPplot1<-ggplot(data=PRdata, aes(x=dpf, y=GP.nmol.org.min, colour=Stage, group=interaction(Stage, dpf))) +
  geom_boxplot(position=position_dodge(0.9), lwd=1)+
  geom_jitter(shape=16, position=position_dodge(0.9), size=3)+
  geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
  scale_color_brewer(palette = "Dark2", breaks=c("Gastrula", "Larvae"))+
  scale_y_continuous(limits=c(-0.1, 0.1), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
  xlab("Days post fertilization")+
  ylab(expression(bold(paste("Gross P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
  theme_classic() +
  theme(text = element_text(size = 18, color="black"),
        axis.text = element_text(size = 18, color="black"), 
        legend.position = "right",
        axis.title = element_text(size = 18, color="black", face="bold"), 
        legend.title=element_blank(), 
        legend.text = element_text(size=18), 
        plot.margin=unit(c(1,1,1,1), "cm"), 
        axis.title.y = element_text(margin = margin(t = 0, r = 1, b = 0, l = 1)), 
        axis.title.x = element_text(margin = margin(t = 3, r = 0, b = 0, l = 0))); GPplot1

GPplot2<-GPplot1+theme(legend.position="none") 

ggsave("../../Ahya2020/Figures/Respiration/ahya_GP.pdf", plot=GPplot2, height=8, width=7, units = c("in"), dpi=300) #output figure
```
 
Generate mean table.  

```{r, results=TRUE}
meanGP <- plyr::ddply(PRdata, c("Stage", "dpf"), summarise,
                   N    = length(GP.nmol.org.min[!is.na(GP.nmol.org.min)]),
                   mean = mean(GP.nmol.org.min, na.rm=TRUE),
                   sd   = sd(GP.nmol.org.min, na.rm=TRUE),
                   se   = sd / sqrt(N)
); meanGP
```

### *Gross Photosynthesis:Respiration ratio *   

Examine data and remove outliers.  

```{r}
hist(PRdata$ratio)

PRdata<-PRdata%>%filter(ratio > -15 & ratio < 25)

hist(PRdata$ratio)
```

```{r, results=TRUE, warning=FALSE, message=FALSE}
PRplot1<-ggplot(data=PRdata, aes(x=dpf, y=ratio, colour=Stage, group=interaction(Stage, dpf))) +
  geom_boxplot(position=position_dodge(0.9), lwd=1)+
  geom_jitter(shape=16, position=position_dodge(0.9), size=3)+
  scale_color_brewer(palette = "Dark2", breaks=c("Gastrula", "Larvae", "Juvenile"))+
  geom_hline(yintercept=1, linetype="dashed", color="black", size=0.75)+
  scale_y_continuous(limits=c(-15, 10), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
  ylab(expression(bold(paste("P : R ")))) +
  xlab("Days Post Fertilization") +
  theme_classic() +
  theme(text = element_text(size = 18, color="black"),
        axis.text = element_text(size = 18, color="black"), 
        legend.position = "top",
        axis.title = element_text(size = 18, color="black", face="bold"), 
        legend.title=element_blank(), 
        legend.text = element_text(size=22), 
        plot.margin=unit(c(1,1,1,1), "cm"), 
        axis.title.y = element_text(margin = margin(t = 0, r = 1, b = 0, l = 1)), 
        axis.title.x = element_text(margin = margin(t = 3, r = 0, b = 0, l = 0))); PRplot1

PRplot2<-PRplot1+theme(legend.position="right") 

ggsave("../../Ahya2020/Figures/Respiration/ahya_PR.pdf", plot=PRplot1, height=8, width=8, units = c("in"), dpi=300) #output figure
``` 


Generate mean table.    

```{r, results=TRUE}
meanPR <- plyr::ddply(PRdata, c("Stage", "dpf"), summarise,
                   N    = length(ratio[!is.na(ratio)]),
                   mean = mean(ratio, na.rm=TRUE),
                   sd   = sd(ratio, na.rm=TRUE),
                   se   = sd / sqrt(N)
); meanPR
```

### *Plot Panels*  

Combine plots above into a single panel.  

```{r, results=TRUE, warning=FALSE, message=FALSE}
ahya_respiration<-plot_grid(Rplot2, Pplot2, GPplot2, PRplot2, labels = c("A", "B", "C", "D"), label_size=18, ncol=4, nrow=1, rel_heights= c(1,1,1,1), rel_widths = c(0.75,0.75,0.75,1), align="h")

ggsave(filename="../../Ahya2020/Figures/Respiration/ahya_PR_panel.pdf", plot=ahya_respiration, dpi=300, width=24, height=6, units="in")
```



## **Analysis: Larvae**  

### *Respiration*     

Build linear mixed effect model and examine for Respiration.     
```{r, results=TRUE, warning=FALSE, message=FALSE}
Rmodel1<-lmer(R.nmol.org.min~Stage*dpf + (1|Run/dpf), data=PRdata) #run nested within day
anova(Rmodel1, type="II")
```

Check assumptions of model for residual normality and variance. 

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(Rmodel1))
hist(residuals(Rmodel1))
leveneTest(residuals(Rmodel1)~Stage * dpf, data=PRdata)
```

There is no significant effect of stage or dfp on respiration. 

### *Net Photosynthesis*  

Build linear mixed effect model and examine for Net Photosynthesis.  
```{r, results=TRUE, warning=FALSE, message=FALSE}
Pmodel1<-lmer(P.nmol.org.min~Stage*dpf + (1|Run/dpf), data=PRdata)
anova(Pmodel1, type="II")
```
 
Check assumptions of model for residual normality and variance. 

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(Pmodel1))
hist(residuals(Pmodel1))
leveneTest(residuals(Pmodel1)~Stage * dpf, data=PRdata)
```

Stage and dpf are not signficant in analysis of net photosynthesis.    

### *Gross Photosynthesis*  

Build linear mixed effect model and examine for Gross Photosynthesis.  
```{r, results=TRUE, warning=FALSE, message=FALSE}
GPmodel1<-lmer(GP.nmol.org.min~Stage*dpf + (1|Run/dpf), data=PRdata)
anova(GPmodel1, type="II")
```
 
Check assumptions of model for residual normality and variance.  

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(GPmodel1))
hist(residuals(GPmodel1))
leveneTest(residuals(GPmodel1)~Stage * dpf, data=PRdata)
```

Stage and dpf are not significant for gross photosynthesis.  

### *P:R*  
 
Build linear mixed effect model and examine for P:R ratio (gross photosynthesis : respiration)   
```{r, results=TRUE, warning=FALSE, message=FALSE}
PRmodel1<-lmer(ratio~Stage*dpf + (1|Run/dpf), data=PRdata)
anova(PRmodel1, type="II")
```

Check assumptions of model for residual normality and variance. Assumptions are violated, revisit.  

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(PRmodel1))
hist(residuals(PRmodel1))
leveneTest(residuals(PRmodel1)~Stage * dpf, data=PRdata)
```

No significant effect of lifestage or dpf on P:R.  

## **Conclusions**  

(1) There was no positive photosynthesis signal, as expected from a horizontal transmitting species that was not infected.  
(2) No metric had a significant relationship with dfp or stage.  
(3) Respiration rates are low, but this may be expected for such small larvae.  
