---
title: Early life history respirometry (photosynthesis & respiration) with LoLinR
author: "AS Huffmyer"
date: '2020'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
---
## Setup  

Set up workspace, set options, and load required packages.    
```{r, echo=TRUE, show=FALSE}
rm(list=ls(all=TRUE)) 
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("devtools" %in% rownames(installed.packages()) == 'FALSE') install.packages('devtools') 
library(devtools)
if ("segmented" %in% rownames(installed.packages()) == 'FALSE') install.packages('segmented') 
if ("plotrix" %in% rownames(installed.packages()) == 'FALSE') install.packages('plotrix') 
if ("gridExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('gridExtra') 
if ("LoLinR" %in% rownames(installed.packages()) == 'FALSE') install_github('colin-olito/LoLinR') 
if ("lubridate" %in% rownames(installed.packages()) == 'FALSE') install.packages('lubridate') 
if ("chron" %in% rownames(installed.packages()) == 'FALSE') install.packages('chron') 
if ("plyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('plyr') 
if ("dplyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('dplyr') 
if ("stringr" %in% rownames(installed.packages()) == 'FALSE') install.packages('stringr') 
if ("Rmisc" %in% rownames(installed.packages()) == 'FALSE') install.packages('Rmisc') 


#load packages
library("ggplot2")
library("segmented")
library("plotrix")
library("gridExtra")
library("LoLinR")
library("lubridate")
library("chron")
library('plyr')
library('dplyr')
library('stringr')
library('Rmisc')

```

## Respirometry in LoLinR for MicroSens data SDR 641 and SDR 710

Modified from K. Wong, S. Gurr, H. Putnam scripts  

Set the path of all respirometry files within the R project.   

```{r, warning=FALSE, message=FALSE}
path.r<-"../../Ahya2020/Data/Respiration/Runs/R" #location of R files
path.p<-"../../Ahya2020/Data/Respiration/Runs/P" #location of P files
```

Generate respiration data frames that will be populated by LolinR.

```{r, warning=FALSE, message=FALSE}
#generate a 6 column dataframe with specific column names
#respiration
Resp.R <- data.frame(matrix(NA, ncol=6))
colnames(Resp.R) <- c("Date", "Run", "Sample.ID","Chamber.ID","Intercept", "mmol.L.min")

Resp.Rb <- data.frame(matrix(NA, ncol=6))
colnames(Resp.Rb) <- c("Date", "Run", "Sample.ID","Chamber.ID","Intercept", "mmol.L.min")

#photosynthesis 
Photo.R <- data.frame(matrix(NA, ncol=6))
colnames(Photo.R) <- c("Date", "Run", "Sample.ID","Chamber.ID","Intercept", "mmol.L.min")

Photo.Rb <- data.frame(matrix(NA, ncol=6))
colnames(Photo.Rb) <- c("Date", "Run", "Sample.ID","Chamber.ID","Intercept", "mmol.L.min")
```

Grab phase (either P or R) from file names that will be used to sort files for separate P and R analysis.  
```{r}
Pfiles <-list.files(path.p)
Rfiles <-list.files(path.r)
```

Load in the sample information file for P and R files. For the PresSens, It is important to have chambers in the order "A1, B1, C1, D1", rather than "A1, A2, A3, A4". BUT the MicroSens orders in "A1, A2, A3, A4". So the Sample Info should be in this order for the MicroSens. 

```{r, warning=FALSE, message=FALSE}
Sample.Info<-read.csv("../../Ahya2020/Data/Respiration/Resp_Sample_Info.csv") #information on life stage, number of individuals, runs, etc.
rename <- Sample.Info$Chamber.ID #Grab info on chamber and sample ID from the sample list. Since P and R are replicate files this list applies to both P and R files
samp <- Sample.Info$Sample.ID

Rrun <- str_sub(Rfiles, 10, 17) #grab run from file name
Rdate <- str_sub(Rfiles, 1, str_length(Rfiles)-15) #grab date from file name

Prun <- str_sub(Pfiles, 10, 17) #grab run from file name
Pdate <- str_sub(Pfiles, 1, str_length(Pfiles)-15) #grab date from file name
```

Verify that time conversion works that is used in the loop below with a test file. 
```{r, results=FALSE}
test<-read.table("../../Ahya2020/Data/Respiration/Runs/P/20201111_Run1-641_P.csv", skip = 47, header=T, sep=",", na.strings="NA", fill = TRUE, as.is=TRUE, fileEncoding="latin1")
#reads in correctly 

test$Time.Min <- as.matrix(read.table(text = test[,4], sep = ":")) %*% c(60, 1, 1/60) #convert hh:mm:ss to minutes works!
test$Time.Min
```

Run loop to extract slopes from respiration data. Subset for respiration data by indicating "Rfiles" as grabbed above. 

```{r, results=FALSE, warning=FALSE, message=FALSE}

for(i in 1:length(Rfiles)) { # for every file in list start at the first and run this following function
  Resp.Data <-read.table(file.path(path.r,Rfiles[i]), skip = 46, header=T, sep=",", na.strings="NA", fill = TRUE, as.is=TRUE, fileEncoding="latin1") #reads in the data files, skip first 47 lines, has header, csv comma separated, assign NAs
  Resp.Data$Time.Min <- as.matrix(read.table(text = Resp.Data[,4], sep = ":")) %*% c(60, 1, 1/60) #set time as "relative time" column in data sheet and convert hh:mm:ss to decimal minutes
  Resp.Data <- Resp.Data %>% #filters data by removing first 10 min as temp/plate acclimates and removing last 5 min to prevent hypoxic effects
     filter(Time.Min >= 10 & Time.Min <= 25) 
  Resp.Data.N <- Resp.Data[,53:76] #subset desired columns in datasheet
  #Resp.Data.N[is.na(Resp.Data.N)] <- 0.1 # turn na's into 0.1's as a dummy variable, make sure this is only populating the missing column, not other cells in datasheet, has to be a value <1

  for(j in 1:(ncol(Resp.Data.N))){
    model <- rankLocReg(
      xall=Resp.Data$Time.Min, yall=as.numeric(Resp.Data.N[, j]), 
      alpha=0.4, method="pc", verbose=TRUE) #extract slopes, percentile rank method with minimum window size of 0.4. This means that in order to fit a slope, it has to encompass at least 40% of available datapoints. 
    pdf(paste0("../../Ahya2020/Output/Respiration/RespirationPlots/",Rdate[i], "_",Rrun[i],"_",rename[j],"_regression_trunc.pdf")) #generate output file names
    plot(model)
    dev.off()
    
    Resp.Rb[j,1] <- as.character(Rdate[i]) #stores the date
    Resp.Rb[j,2] <- as.character(Rrun[i]) #stores the run number
    Resp.Rb[j,3] <- as.character(samp[j+(i-1)*ncol(Resp.Data.N)]) #stores the sample ID 
    Resp.Rb[j,4] <- as.character(rename[j]) #stores the chamber ID
    Resp.Rb[j,5:6] <- model$allRegs[i,c(4,5)] #inserts slope and intercept in the dataframe
    
  }
  Resp.R <- rbind(Resp.R, Resp.Rb) #bind final data frame
}

```

Save respiration data frames.  

```{r, results=FALSE, warning=FALSE, message=FALSE}
Resp.R <- Resp.R[-1,] #remove empty row

write.csv(Resp.R, paste0("../../Ahya2020/Output/Respiration/Respiration_rates_Ahya2020.csv", sep="")) #save respiration rate data
```

Run loop to extract slopes from photosynthesis data (Pfiles). 

```{r, warning=FALSE, message=FALSE}
for(i in 1:length(Pfiles)) { # for every file in list start at the first and run this following function
  Photo.Data <-read.table(file.path(path.p,Pfiles[i]), skip = 46, header=T, sep=",", na.strings="NA", fill = TRUE, as.is=TRUE, fileEncoding="latin1") #reads in the data files, skip first 47 lines, has header, csv comma separated, assign NAs
  Photo.Data$Time.Min <- as.matrix(read.table(text = Photo.Data[,4], sep = ":")) %*% c(60, 1, 1/60) #set time as "relative time" column in data sheet and convert hh:mm:ss to decimal minutes
  Photo.Data <- Photo.Data %>% #filters data by removing first 5 min as temp/plate acclimates
     filter(Time.Min >= 5) 
  Photo.Data.N <- Photo.Data[,53:76] #subset desired columns in datasheet
  #Photo.Data.N[is.na(Photo.Data.N)] <- 0.01 # turn na's into 0.1's as a dummy variable, make sure this is only populating the missing column, not other cells in datasheet, has to be a value <1

  for(j in 1:(ncol(Photo.Data.N))){
    model <- rankLocReg(
      xall=Photo.Data$Time.Min, yall=as.numeric(Photo.Data.N[, j]), 
      alpha=0.4, method="pc", verbose=TRUE) #extract slopes, percentile rank method with minimum window size of 0.4. This means that in order to fit a slope, it has to encompass at least 40% of available datapoints. 
    pdf(paste0("../../Ahya2020/Output/Respiration/PhotosynthesisPlots/",Pdate[i], "_",Prun[i],"_",rename[j],"_regression_trunc.pdf")) #generate output file names
    plot(model)
    dev.off()
    
    Photo.Rb[j,1] <- as.character(Pdate[i]) #stores the date
    Photo.Rb[j,2] <- as.character(Prun[i]) #stores the run number
    Photo.Rb[j,3] <- as.character(samp[j+(i-1)*ncol(Photo.Data.N)]) #stores the sample ID 
    Photo.Rb[j,4] <- as.character(rename[j]) #stores the chamber ID
    Photo.Rb[j,5:6] <- model$allRegs[i,c(4,5)] #inserts slope and intercept in the dataframe
    
  }
  Photo.R <- rbind(Photo.R, Photo.Rb) #bind final data frame
}
```

Save photosynthesis data frames.  

```{r, warning=FALSE, message=FALSE}
Photo.R <- Photo.R[-1,] #remove empty row

write.csv(Photo.R, paste0("../../Ahya2020/Output/Respiration/Photosynthesis_rates_Ahya2020.csv", sep="")) #save P rates
```


## Format output data frames  

```{r, warning=FALSE, message=FALSE}
Resp.Rates <- read.csv(file = "../../Ahya2020/Output/Respiration/Respiration_rates_Ahya2020.csv") #read file back in so slopes don't have to be generated every time
Resp.Rates = subset(Resp.Rates, select = -c(X) ) #remove empty column

#format "run" column
Resp.Rates<-Resp.Rates %>%
  mutate(Run = str_sub(Run, 4, 8))
Resp.Rates$Run<-as.factor(Resp.Rates$Run) #format as factor 
```


```{r, warning=FALSE, message=FALSE}
Photo.Rates <- read.csv(file = "../../Ahya2020/Output/Respiration/Photosynthesis_rates_Ahya2020.csv") #import P rates
Photo.Rates = subset(Photo.Rates, select = -c(X) ) #remove empty column

#format "run" column
Photo.Rates<-Photo.Rates %>% #Format as "#" rather than "Run#"
  mutate(Run = str_sub(Run, 4, 8))
Photo.Rates$Run<-as.factor(Photo.Rates$Run) #format as factor

Sample.Info<-read.csv("../../Ahya2020/Data/Respiration/Resp_Sample_Info.csv") #information on life stage, number of individuals, runs, etc.
```

Merge P and R files with sample info for testing and manipulation.  

```{r, warning=FALSE, message=FALSE}
oxygen<-full_join(Sample.Info, Resp.Rates) #add respiration data
colnames(oxygen)[colnames(oxygen) == 'Intercept'] <- 'Resp.Intercept' #rename to specify R
colnames(oxygen)[colnames(oxygen) == 'mmol.L.min'] <- 'R.mmol.L.min' #rename to specify R

oxygen<-full_join(oxygen, Photo.Rates) #add photosynthesis data
colnames(oxygen)[colnames(oxygen) == 'Intercept'] <- 'Photo.Intercept' #rename to specify P
colnames(oxygen)[colnames(oxygen) == 'mmol.L.min'] <- 'P.mmol.L.min' #rename to specify P
```

Account for volume and substract blank values. Average blank calculated for each run for P and R phases. Display mean blank value in order of R and P phases.    
```{r, results=TRUE, warning=FALSE, message=FALSE}
#Account for chamber volume to convert from mmol L-1 m-1 to mmol m-1. This standardizes across water volumes (different because of coral size) and removes per Liter
oxygen$R.mmol.min <- oxygen$R.mmol.L.min * oxygen$Volume #calculate

oxygen$P.mmol.min <- oxygen$P.mmol.L.min * oxygen$Volume #calculate

blank_data <- subset(oxygen, Type == "Blank") #subset to blank data only
plot(blank_data$R.mmol.min , ylab="mmol O2 min-1") #blanks during dark phase
plot(blank_data$P.mmol.min , ylab="mmol O2 min-1") #blanks during light phase

#display mean blankvalues
mean(blank_data$R.mmol.min)*1000*1000 #mean R phase blanks
mean(blank_data$P.mmol.min)*1000*1000 #mean P phase blanks
mean(oxygen$R.mmol.min)*1000*1000 #mean R rates
mean(oxygen$P.mmol.min)*1000*1000 #mean P rates Something off here !!!!!!!!!!!!!!!

resp.blnk <- aggregate(R.mmol.min ~ Date*Run, data=blank_data, mean) #calculate average blank during light for each run
colnames(resp.blnk)[colnames(resp.blnk) == 'R.mmol.min'] <- 'R.Blank.mmol.min' #rename to specify blank for R

photo.blnk <- aggregate(P.mmol.min ~ Date*Run, data=blank_data, mean) #calculate average blank during dark for each run
colnames(photo.blnk)[colnames(photo.blnk) == 'P.mmol.min'] <- 'P.Blank.mmol.min' #rename to specify blank for P

oxygen <- full_join(oxygen, resp.blnk) #add R blanks to master
oxygen <- full_join(oxygen, photo.blnk) #add P blanks to master
```

Subtract blank values to generate a "corrected" value for umol O2 min-1.  

```{r, warning=FALSE, message=FALSE}
oxygen$R.mmol.min.corr<-oxygen$R.mmol.min-oxygen$R.Blank.mmol.min #subtract R blanks
oxygen$P.mmol.min.corr<-oxygen$P.mmol.min-oxygen$P.Blank.mmol.min #subtract P blanks
```

Normalize to biologically relevant measure. Here, normalize to number of larvae. This can be substituted or changed for larval size/volume as well.  

```{r, warning=FALSE, message=FALSE}
oxygen.bio <- oxygen %>% filter(Type == "Sample") #isolate only biological samples and drop unused factor levels 
oxygen.bio <- droplevels(oxygen.bio) #drop unused factor levels

#respiration conversions
oxygen.bio$R.mmol.org.min <- oxygen.bio$R.mmol.min.corr/oxygen.bio$Org.Number #calculate oxygen per organism (larvae or recruit)
oxygen.bio$R.umol.org.min <- oxygen.bio$R.mmol.org.min*1000 #calculate micromoles 
oxygen.bio$R.nmol.org.min <- oxygen.bio$R.umol.org.min*1000 #calculate nanmoles 

#respiration conversions
oxygen.bio$P.mmol.org.min <- oxygen.bio$P.mmol.min.corr/oxygen.bio$Org.Number #calculate oxygen per organism (larvae or recruit)
oxygen.bio$P.umol.org.min <- oxygen.bio$P.mmol.org.min*1000 #calculate micromoles 
oxygen.bio$P.nmol.org.min <- oxygen.bio$P.umol.org.min*1000 #calculate nanmoles 

#gross photosynthesis
oxygen.bio$GP.mmol.min.corr <- oxygen.bio$P.mmol.min.corr - oxygen.bio$R.mmol.min.corr #calculate gross photosynthesis by subtracting out oxygen consumption from respiration
oxygen.bio$GP.mmol.org.min <- oxygen.bio$GP.mmol.min.corr/oxygen.bio$Org.Number #calculate oxygen per organism (larvae or recruit)
oxygen.bio$GP.umol.org.min <- oxygen.bio$GP.mmol.org.min*1000 #calculate micromoles 
oxygen.bio$GP.nmol.org.min <- oxygen.bio$GP.umol.org.min*1000 #calculate nanmoles 

```

Save as .csv file.  
```{r, warning=FALSE, message=FALSE}
write.csv(oxygen.bio, paste0("../../Ahya2020/Output/Respiration/oxygen_P_R_calc.csv", sep="")) #save final file
```

This point ends the general use script for P and R rates. 

