---
title: "Larval size analysis - Mcap 2020"
author: "AS Huffmyer"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

# Set Up  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r setup, include = FALSE}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("RColorBrewer")) install.packages("RColorBrewer")
if (!require("car")) install.packages("car")
if (!require("emmeans")) install.packages("emmeans")
if (!require("multcomp")) install.packages("multcomp")
if (!require("knitr")) install.packages("knitr")
if (!require("kableExtra")) install.packages("kableExtra")

# load packages
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(car)
library(emmeans)
library(multcomp)
library(knitr)
library(kableExtra)
```


# Import and manipulate data
```{r}
# Larval size data
size <- read_csv("Mcap2020/Data/Physiology/Size/larval_size.csv")

#metadata
metadata <- read_csv("Mcap2020/Data/lifestage_metadata.csv")

size <- left_join(size, metadata)
size$hpf <- as.factor(size$hpf)
```

## Prep data frame, considering each individual larvae as a replicate here, not summarising by tube  
```{r}
# Calculate mean counts for each sample
df <- size %>%
  dplyr::select(tube.ID, lifestage, replicate, `area (mm)`, hpf, group)%>%
  drop_na()%>% #remove na's that could not be measured
  rename(area=`area (mm)`)
  #group_by(tube.ID) %>%
  #summarise(mean_area = mean(`area (mm)`))

#match in identifying information
#df$lifestage<-size$lifestage[match(df$tube.ID, size$tube.ID)]
#df$hpf<-size$hpf[match(df$tube.ID, size$tube.ID)]
#df$group<-size$group[match(df$tube.ID, size$tube.ID)]
df$lifestage<-as.factor(df$lifestage)
```

# Plotting  

## Plot data with mean and standard error
```{r}
df %>%
  ggplot(aes(x = lifestage, y = area, color = lifestage)) +
  labs(x = "",y = "Mean Larval Size (mm^2)") +
geom_jitter(width = 0.1) +                                            # Plot all points
  stat_summary(fun.data = mean_cl_normal, fun.args = list(mult = 1),    # Plot standard error
               geom = "errorbar", color = "black", width = 0.5) +
  stat_summary(fun.y = mean, geom = "point", color = "black") + # Plot mean
  theme_classic()
```

Present means and standard error of each group and save summary table

```{r}
df%>%
  group_by(lifestage, hpf)%>%
  summarize(n=length(area),
            Mean=format(round(mean(area), 3), 3), 
            SE=format(round(sd(area)/sqrt(length(area)),3),3))%>%
  rename(Lifestage=lifestage, HPF=hpf)%>%
  kbl(caption="Descriptive statistics of larval size across ontogeny")%>%
  kable_classic(full_width=FALSE, html_font="Arial")%>%
  row_spec(0, bold = TRUE) 

#need to output to csv 
df%>%
  group_by(lifestage, hpf)%>%
  summarize(n=length(area),
            Mean=format(round(mean(area), 3), 3), 
            SE=format(round(sd(area)/sqrt(length(area)),3),3))%>%
  rename(Lifestage=lifestage, HPF=hpf)%>%
  write_csv(., "Mcap2020/Output/Physiology/larval_size_table.csv")
```


## Plot data as a scatterplot   
```{r}
df$hpf<-as.factor(df$hpf)
size_plot<-df %>%
    ggplot(., aes(x = hpf, y = area)) +
    #geom_boxplot(outlier.size = 0) +
    geom_smooth(method="loess", se=TRUE, fullrange=TRUE, level=0.95, color="black") +
    geom_point(aes(fill=group, group=group), pch = 21, size=4, position = position_jitterdodge(0.1)) + 
    xlab("Hours Post-Fertilization") + 
    scale_fill_manual(name="Lifestage", values=c("#8C510A", "#DFC27D","#80CDC1", "#003C30"))+
    ylab(expression(bold(paste("Larval Size (mm"^2, ")"))))+
    ylim(0,1)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); size_plot

#EGG: #8C510A
#EMBRYO: #DFC27D
#LARVAE: #80CDC1
#RECRUIT: #003C30

#COLORS TO TRY FROM BRBG: "#543005" "#8C510A" "#BF812D" "#DFC27D" "#F6E8C3" "#F5F5F5" "#C7EAE5" "#80CDC1" "#35978F" "#01665E" "#003C30"

#ggsave("Mcap2020/Figures/Physiology/SymbiontCounts.png", symb_plot, dpi=300, w=8.5, h=5, units="in")
```

## Plot data as box plot  
```{r}

size_plot2<-df %>%
    ggplot(., aes(x = hpf, y = area)) +
    geom_boxplot(aes(color=group), outlier.size = 0, lwd=1) +
    #geom_smooth(method="loess", se=TRUE, fullrange=TRUE, level=0.95, color="black") +
    geom_point(aes(fill=group), pch = 21, size=2, position = position_jitterdodge(0.1)) + 
    xlab("Hours Post-Fertilization") + 
    scale_fill_manual(name="Lifestage", values=c("#8C510A", "#DFC27D","#80CDC1", "#003C30"))+
    scale_color_manual(name="Lifestage", values=c("#8C510A", "#DFC27D","#80CDC1", "#003C30"))+
    ylab(expression(bold(paste("Larval Size (mm"^2, ")"))))+
    ylim(0, 0.8)+
    theme_classic() + 
    #geom_text(label="A", x=1, y=2500, size=4, color="black")+ #egg
    #geom_text(label="A", x=2, y=2500, size=4, color="black")+ #embryo 1
    #geom_text(label="A", x=3, y=2500, size=4, color="black")+ #larvae 1
    #geom_text(label="AB", x=4, y=4100, size=4, color="black")+ #larvae 2
    #geom_text(label="AB", x=5, y=4100, size=4, color="black")+ #larvae 3
    #geom_text(label="AB", x=6, y=4100, size=4, color="black")+ #larvae 4
    #geom_text(label="BC", x=6.8, y=4500, size=4, color="black")+ #larvae 5
    #geom_text(label="CD", x=7.2, y=6500, size=4, color="black")+ #recruit 1
    #geom_text(label="D", x=7.8, y=6500, size=4, color="black")+ #larvae6
    #geom_text(label="D", x=8.2, y=8700, size=4, color="black")+ #recruit2
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14)
      ); size_plot2

ggsave("Mcap2020/Figures/Physiology/LarvalSize_Box.png", size_plot2, dpi=300, w=8.5, h=5, units="in")
```

# Statistical analysis  

Run ANOVA on cells per larvae by sampling point, specified by sequence of samples taken (life stage, hpf).    

```{r}
model<-aov(area~lifestage, data=df)
qqPlot(residuals(model))
leveneTest(residuals(model)~lifestage, data=df)
```

Violation in normality and variance assumptions. Conduct non-parametric test (Kruskal Wallis).  

```{r}
kruskal.test(area~lifestage, data=df)
```

Significant effect of lifestage on larval size.  

View posthoc comparisons for differences between lifestages.  

```{r}
emm = emmeans(model, ~ lifestage)
cld(emm, Letters=c(LETTERS)) #letter display
```


