---
title: "Physiology Analysis: symbiont densities and larval size"
author: "AS Huffmyer"
date: "2023" 
output:
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

This script analyzes larval size and symbiont density as physiology metrics in the *M. capitata* developmental time series project.    

# **Setup**  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r setup, include = FALSE}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("RColorBrewer")) install.packages("RColorBrewer")
if (!require("car")) install.packages("car")
if (!require("emmeans")) install.packages("emmeans")
if (!require("multcomp")) install.packages("multcomp")
if (!require("knitr")) install.packages("knitr")
if (!require("kableExtra")) install.packages("kableExtra")
if (!require("lme4")) install.packages("lme4")
if (!require("GGally")) install.packages("GGally")
if (!require("lmerTest")) install.packages("lmerTest")
if (!require("cowplot")) install.packages("cowplot")

# load packages
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(car)
library(emmeans)
library(multcomp)
library(knitr)
library(kableExtra)
library(lme4)
library(GGally)
library(lmerTest)
library(cowplot)
library(Hmisc)
```

In this project, physiological assays include symbiont cell density and larval volume.  

# **1. Larval Size/Volume**  

First, plot and analyze larval size measured as volume. Volume is calculated as an elliptical sphere using the length and width measurements of each individual sampled across the developmental time series.     

## Import and manipulate data

Load in data, match in metadata about lifestage, and rename levels of lifestage factor. 
```{r}
# Larval size data
size <- read_csv("Mcap2020/Data/Physiology/Size/larval_volume.csv")

#metadata
metadata <- read_csv("Mcap2020/Data/lifestage_metadata.csv")

# select columns
size <- size %>%
  dplyr::select(tube.ID, code, replicate, volume)%>%
  drop_na() #remove na's that could not be measured

size$replicate<-as.factor(size$replicate)
size$tube.ID<-as.factor(size$tube.ID)
size$code<-as.factor(size$code)

size$hpf<-metadata$hpf[match(size$code, metadata$code)]
size$group<-metadata$group[match(size$code, metadata$code)]
size$lifestage<-metadata$lifestage[match(size$code, metadata$code)]

#change metamorphosed recruit to metamorphosed polyp 
size$group <- if_else(size$group == "Metamorphosed Recruit", "Metamorphosed Polyp", size$group)
size$lifestage <- if_else(size$lifestage == "Metamorphosed Recruit (231 hpf)", "Metamorphosed Polyp (231 hpf)", size$lifestage)
size$lifestage <- if_else(size$lifestage == "Metamorphosed Recruit (183 hpf)", "Metamorphosed Polyp (183 hpf)", size$lifestage)

size$hpf <- as.factor(size$hpf)
size$group <- as.factor(size$group)
size$lifestage <- as.factor(size$lifestage)

#reorder lifestage factor 
size$lifestage <- factor(size$lifestage, levels=c("Egg (1 hpf)", "Embryo (5 hpf)", "Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf)", "Larvae (163 hpf)", "Larvae (183 hpf)", "Larvae (231 hpf)", "Metamorphosed Polyp (183 hpf)", "Metamorphosed Polyp (231 hpf)"))
```

Remove na's from data frame.  
```{r}
# Calculate mean counts for each sample
size <- size %>%
  dplyr::select(tube.ID, lifestage, replicate, volume, hpf, group, code)%>%
  drop_na()
```

## Plot larval volume    

Plot data with mean and standard error for each lifestage.  
```{r}
size %>%
  ggplot(aes(x = lifestage, y = volume, color = group)) +
  labs(x = "",y = "Mean Larval Volume (mm^3)") +
  geom_jitter(width = 0.1) +                                            # Plot all points
  stat_summary(fun.data = mean_cl_normal, fun.args = list(mult = 1),    # Plot standard error
               geom = "errorbar", color = "black", width = 0.5) +
  stat_summary(fun = mean, geom = "point", color = "black") + # Plot mean
  theme_classic()
```

Present means and standard error of each group and save summary table.  

```{r}
size%>%
  group_by(lifestage, hpf)%>%
  summarise(n=length(volume),
            Mean=format(round(mean(volume), 3), 3), 
            SE=format(round(sd(volume)/sqrt(length(volume)),3),3))%>%
  rename(Lifestage=lifestage, HPF=hpf)%>%
  kbl(caption="Descriptive statistics of larval volume across ontogeny")%>%
  kable_classic(full_width=FALSE, html_font="Arial")%>%
  row_spec(0, bold = TRUE) 

#need to output to csv 
size%>%
  group_by(lifestage, hpf)%>%
  summarise(n=length(volume),
            Mean=format(round(mean(volume), 3), 3), 
            SE=format(round(sd(volume)/sqrt(length(volume)),3),3))%>%
  rename(Lifestage=lifestage, HPF=hpf)%>%
  write_csv(., "Mcap2020/Output/Physiology/larval_volume_table.csv")
```

Plot data as a box plot.     
```{r}
size_plot<-size %>%
    ggplot(., aes(x = lifestage, y = volume, colour=group)) +
    geom_boxplot(aes(colour=group), lwd=1) +
    geom_point(aes(fill=group, group=group), colour="black", pch = 21, size=4, position = position_jitterdodge(0.5), alpha=0.6) + 
    xlab("") + 
    scale_colour_manual(name="Lifestage", values=c("#8C510A", "#DFC27D","#80CDC1", "#003C30"))+
    scale_fill_manual(name="Lifestage", values=c("#8C510A", "#DFC27D","#80CDC1", "#003C30"))+
    ylab(expression(bold(paste("Volume (mm"^3, ")"))))+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12), 
      axis.text.x=element_text(angle=45, hjust=1)
      ); size_plot

ggsave(plot=size_plot, "Mcap2020/Figures/Physiology/size.png",height=8, width=8)
```

## Statistical analysis  

Run ANOVA test on larval size. 
```{r}
size_model<-aov(volume~lifestage, data=size)
summary(size_model)
qqPlot(residuals(size_model))
leveneTest(residuals(size_model)~lifestage, data=size)
```

There appears to be a V=violation in normality and variance assumptions. Conduct non-parametric test (Kruskal Wallis).  

```{r}
kruskal.test(volume~lifestage, data=size)
```

There is a significant difference in size (larval volume) between lifestages.   

View posthoc comparisons for differences between lifestages.  

```{r}
emm = emmeans(size_model, ~ lifestage)
cld(emm, Letters=c(LETTERS)) #letter display
pairs(emm)
```

Metamorphosed polyps are smaller in volume than larvae within each respective time point. Size is quite variable between life stages.   

# **2. Symbiont Density**  

Symbiont density was measured across life stages. For egg to metamorphosed polyp stages, this was done by sampling individuals into a tube. We are able to normalize cell densities to both per individual (individual normalized) and per unit individual size (volume normalized). For attached recruit time points, the individuals were scraped from plugs and then homogenized and used for cell density analysis. Prior to scraping, we quanitified surface area of the recruits on plugs. For these timepoints, cell density is only able to be normalized to surface area (surface area normalized). For this reason, we will plot these datasets separately and cannot compare cell densities between attached recruits and other life stages.  

## Symbiont density per individual  

First, look at symbiont cell densities per individual.  

### Import and manipulate data  
```{r}
# Cell count data
sym_counts <- read_csv("Mcap2020/Data/Physiology/CellDensity/symbiont.counts.csv")

sym_counts<-sym_counts%>%
  rename(code=lifestage)
```

Calculate cells and normalize to either planar size (eggs through metamorphosed recruits) or surface area (attached recruits)   
```{r}
# Calculate mean counts for each sample
df <- sym_counts %>%
  dplyr::select(tube.ID, num.squares, matches("count[1-6]")) %>%
  gather("rep", "count", -tube.ID, -num.squares) %>%
  group_by(tube.ID, num.squares) %>%
  summarise(mean_count = mean(count, na.rm = TRUE))

#match in identifying information
df$code<-sym_counts$code[match(df$tube.ID, sym_counts$tube.ID)]
df$total.volume.ul<-sym_counts$total.volume.ul[match(df$tube.ID, sym_counts$tube.ID)]
df$num.individuals<-sym_counts$num.individuals[match(df$tube.ID, sym_counts$tube.ID)]
df$surface.area<-sym_counts$surface.area[match(df$tube.ID, sym_counts$tube.ID)]

#add in metadata
df$hpf<-metadata$hpf[match(df$code, metadata$code)]
df$group<-metadata$group[match(df$code, metadata$code)]
df$lifestage<-metadata$lifestage[match(df$code, metadata$code)]

df$group <- if_else(df$group == "Metamorphosed Recruit", "Metamorphosed Polyp", df$group)
df$lifestage <- if_else(df$lifestage == "Metamorphosed Recruit (231 hpf)", "Metamorphosed Polyp (231 hpf)", df$lifestage)
df$lifestage <- if_else(df$lifestage == "Metamorphosed Recruit (183 hpf)", "Metamorphosed Polyp (183 hpf)", df$lifestage)

# Normalize counts by homogenate volume and surface area
df <- df %>%
  mutate(cells.mL = mean_count * 10000 / num.squares,
         cells = cells.mL * (total.volume.ul/1000),
         cells.ind = cells / num.individuals, 
         cells.mm = cells / surface.area)

sym_counts<-df

# order factor 
sym_counts$lifestage<-factor(sym_counts$lifestage, levels=c("Egg (1 hpf)", "Embryo (5 hpf)", "Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf)", "Larvae (163 hpf)", "Larvae (183 hpf)", "Larvae (231 hpf)", "Metamorphosed Polyp (183 hpf)", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf)", "Attached Recruit (231 hpf)", "Attached Recruit (255 hpf)"))
```

Note that in this dataframe there is a column for individual normalized, volume normalized, and surface area normalized, depending on the lifestage as described above.    

### Plotting  

First, plot mean and standard error for cell density normalized to individual (per larva).  

Display cells per individual.  
```{r}
sym_counts %>%
  filter(!group=="Attached Recruit")%>%
  ggplot(aes(x = lifestage, y = cells.ind, color = group)) +
  labs(x = "",y = "Cell Density per larva") +
geom_jitter(width = 0.1) +                                            # Plot all points
  stat_summary(fun.data = mean_cl_normal, fun.args = list(mult = 1),    # Plot standard error
               geom = "errorbar", color = "black", width = 0.5) +
  stat_summary(fun = mean, geom = "point", color = "black") + # Plot mean
  theme_classic()
```

Display cell density per mm2 in attached recruit plugs normalized to surface area. 
```{r}
sym_counts %>%
  filter(group=="Attached Recruit")%>%
  ggplot(aes(x = lifestage, y = cells.mm, color = lifestage)) +
  labs(x = "",y = "Cell Density per mm2") +
geom_jitter(width = 0.1) +                                            # Plot all points
  #stat_summary(fun.data = mean_cl_normal, fun.args = list(mult = 1),    # Plot standard error
               #geom = "errorbar", color = "black", width = 0.5) +
  #stat_summary(fun.y = mean, geom = "point", color = "black") + # Plot mean
  theme_classic()
```

Present means and standard error of each group and save summary table.  

```{r}
sym_counts%>%
  group_by(group, hpf, lifestage)%>%
  summarise(n=length(cells.ind),
            Mean=format(round(mean(cells.ind), 0), 0), 
            SE=format(round(sd(cells.ind)/sqrt(length(cells.ind)),0),0))%>%
  rename(Lifestage=group, HPF=hpf)%>%
  kbl(caption="Descriptive statistics of Symbiodiniaceae cell densities per larva across ontogeny")%>%
  kable_classic(full_width=FALSE, html_font="Arial")%>%
  row_spec(0, bold = TRUE) 

sym_counts%>%
  group_by(group, hpf, lifestage)%>%
  summarise(n=length(cells.mm),
            Mean=format(round(mean(cells.mm), 0), 0), 
            SE=format(round(sd(cells.mm)/sqrt(length(cells.mm)),0),0))%>%
  rename(Lifestage=group, HPF=hpf)%>%
  kbl(caption="Descriptive statistics of Symbiodiniaceae cell densities per mm2 across ontogeny")%>%
  kable_classic(full_width=FALSE, html_font="Arial")%>%
  row_spec(0, bold = TRUE) 

#need to output to csv 
sym_counts%>%
  group_by(group, hpf, lifestage)%>%
  summarise(n=length(cells.ind),
            Mean=format(round(mean(cells.ind), 0), 0), 
            SE=format(round(sd(cells.ind)/sqrt(length(cells.ind)),0),0))%>%
  rename(group=group, HPF=hpf)%>%
  write_csv(., "Mcap2020/Output/Physiology/cell_density_table.csv")
```

Plot data as box plot for egg through metamorphosed polyp stages normalized to cells per individual.    
```{r}
symb_plot2<-sym_counts %>%
    filter(!group=="Attached Recruit")%>%
    droplevels()%>%
    ggplot(., aes(x = lifestage, y = cells.ind)) +
    #geom_boxplot(aes(color=group), outlier.size = 0, lwd=1) +
    #geom_smooth(method="loess", se=TRUE, fullrange=TRUE, level=0.95, color="black") +
    geom_point(aes(fill=group), pch = 21, size=4, position = position_jitterdodge(0.3)) + 
    xlab("") + 
    scale_fill_manual(name="Lifestage", values=c("#8C510A", "#DFC27D","#80CDC1", "#003C30"))+
    scale_color_manual(name="Lifestage", values=c("#8C510A", "#DFC27D","#80CDC1", "#003C30"), guide="none")+
    ylab(expression(bold(paste("Symbiont cells individual"^-1))))+
    ylim(0,9000)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      axis.text.x=element_text(angle=45, hjust=1)
      ); symb_plot2

```

### Statistical analysis  

Run ANOVA for cell density per individual.  

```{r}
sym_ind_model_data<-sym_counts%>%
      filter(!group=="Attached Recruit")%>%
      droplevels()

sym_ind_model<-aov(cells.ind~lifestage, data=sym_ind_model_data)
summary(sym_ind_model)
qqPlot(residuals(sym_ind_model))
leveneTest(residuals(sym_ind_model)~lifestage, data=sym_ind_model_data)
```

Both normality and homogeneity of variance pass.   

There is a significant difference in cell densities bewteen lifestages. View posthoc comparisons for differences between lifestages.  

```{r}
emm = emmeans(sym_ind_model, ~ lifestage)
cld(emm, Letters=c(LETTERS)) #letter display
pairs(emm)
```

Cell densities increase over development with highest densities at 231 hours post fertilization in both larvae and metamorphosed polyps. 

Output data to file.  

```{r}
sym_counts %>%
  write_csv(., file = "Mcap2020/Output/Physiology/calculated_densities.csv")
```


## Symbiont density per unit size  

### Data manipulation and correlation  

First, test for correlation between symbiont cell density and larval volume to see if there is a relationship.  
Generate data frame with summarized size and cell density information for each time point from eggs to metamorphosed recruits, because we have data for volume and counts for each sample. We do not include attached recruits here yet, because we cannot calculate densities per individual.   

```{r}
#read in data frame generated in previous chunk 
sym_counts<-sym_counts%>%
  dplyr::select(tube.ID, lifestage, group, hpf, cells.ind, cells.mm)

#grab size data
volume<-size%>%
  group_by(tube.ID)%>%
  summarise(mean_volume=mean(volume, na.rm=TRUE))

volume$tube.ID<-as.factor(volume$tube.ID)
sym_counts$hpf<-as.factor(sym_counts$hpf)

corr<-left_join(sym_counts, volume)
```

Generate number of symbiont cells per mm^3 area for each tube.    
```{r}
corr<-corr%>%
  mutate(mean_volume_2=mean_volume*2)%>%
  mutate(counts_volume=cells.ind/mean_volume)%>%
  mutate(counts_volume=ifelse(is.na(counts_volume), cells.mm, counts_volume)) #add attached recruit data already calculated as cells per mm2
```

Plot correlation between cell counts (cells per individual) and volume (mm^3).  

```{r}
correlation<-corr %>%
    filter(!group=="Attached Recruit")%>%
    ggplot(., aes(x = mean_volume, y = cells.ind)) +
    #geom_smooth(method="lm", se=TRUE, fullrange=TRUE, level=0.95, color="black", fill="gray") +
    geom_point(aes(fill=group), pch = 21, size=4) + 
    xlab(expression(bold(paste("Larval Size (mm"^2, ")")))) + 
    scale_fill_manual(name="Lifestage", values=c("#8C510A", "#DFC27D","#80CDC1", "#003C30"))+
    scale_color_manual(name="Lifestage", values=c("#8C510A", "#DFC27D","#80CDC1", "#003C30"))+
    xlab(expression(bold(paste("Individual Volume (mm"^3, ")"))))+
    ylab(expression(bold(paste("Symbiont cells individual"^-1))))+
    #ylim(0, 9000)+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14)
      ); correlation
```

Test relationship with a spearman correlation.  

```{r}
cor.test(corr$mean_volume, corr$cells.ind, method=c("spearman"))
```

There is not a significant correlation between volume and symbiont density.   

### Plotting  

Plot cells per mm^3 as a boxplot.    

```{r}
#order for creating a legend for all plots 
corr$group <- factor(corr$group, levels = c("Egg", "Embryo", "Larvae", "Metamorphosed Polyp", "Attached Recruit"))

cells_size_plot<-corr %>%
    filter(!group=="Attached Recruit")%>%
    ggplot(., aes(x = hpf, y = counts_volume)) +
    geom_boxplot(aes(color=group), outlier.size = 0, lwd=1) +
    geom_point(aes(fill=group), pch = 21, size=4, position = position_jitterdodge(0.4)) + 
    xlab("Hours Post-Fertilization") + 
    scale_fill_manual(name="Lifestage", values=c("#8C510A", "#DFC27D","#80CDC1", "#003C30", "#BA55D3"), guide="none")+
    scale_color_manual(name="Lifestage", values=c("#8C510A", "#DFC27D","#80CDC1", "#003C30", "#BA55D3"))+
    ylab(expression(bold(paste("Volume-normalized cell density (mm"^-3,")"))))+
    #ylim(2000, 35000)+
    theme_classic() + 
    theme(
      legend.position="top",
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14)
      ); cells_size_plot

#making a plot with a legend for use later on 
#we are NOT using this plot for comparisons due to different units for attached recruits that cannot be compared.  
legend_plot<-corr %>%
    #filter(!group=="Attached Recruit")%>%
    ggplot(., aes(x = hpf, y = counts_volume)) +
    geom_boxplot(aes(color=group), outlier.size = 0, lwd=1) +
    geom_point(aes(fill=group), pch = 21, size=4, position = position_jitterdodge(0.4)) + 
    xlab("Hours Post-Fertilization") + 
    scale_fill_manual(name="Lifestage", values=c("#8C510A", "#DFC27D","#80CDC1", "#003C30", "#BA55D3"), guide="none")+
    scale_color_manual(name="Lifestage", values=c("#8C510A", "#DFC27D","#80CDC1", "#003C30", "#BA55D3"))+
    ylab(expression(bold(paste("Symbiont cells mm"^-3))))+
    #ylim(2000, 35000)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14)
      )

```

Plot as linear relationship.  

```{r}
cells_size_plot2<-corr %>%
    filter(!group=="Attached Recruit")%>%
    ggplot(., aes(x = as.numeric(as.character(hpf)), y = counts_volume)) +
    geom_point(aes(fill=group, group=group), pch = 21, size=4, position = position_jitterdodge(0.4)) + 
    geom_smooth(method="lm", se=TRUE, fullrange=TRUE, level=0.95, color="black") +
    xlab("Hours Post-Fertilization") + 
    scale_fill_manual(name="Lifestage", values=c("#8C510A", "#DFC27D","#80CDC1", "#003C30", "#BA55D3"), guide="none")+
    scale_color_manual(name="Lifestage", values=c( "#8C510A", "#DFC27D","#80CDC1", "#003C30", "#BA55D3"))+
    ylab(expression(bold(paste("Symbiont cells mm"^-3))))+
    #ylim(2000, 35000)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14)
      ); cells_size_plot2
```

There is not a correlation between size and cell density. 

Analyze differences in normalized cell counts by timepoint. 

```{r}
model<-corr%>%
  filter(!group=="Attached Recruit")%>%
  droplevels()%>%
  aov(counts_volume~lifestage, data=.)

qqPlot(residuals(model))

corr%>%
  filter(!group=="Attached Recruit")%>%
  droplevels()%>%
  leveneTest(residuals(model)~lifestage, data=.)

summary(model)
```

Cell density normalized to size is significantly different across development.  

View posthoc differences.  
```{r}
emm = emmeans(model, ~ lifestage)
cld(emm, Letters=c(LETTERS)) #letter display

```

Generate summary table of descriptive statistics.  

```{r}
#need to output to csv 
corr%>%
  group_by(group, hpf, lifestage)%>%
  filter(!group=="Attached Recruit")%>%
  droplevels()%>%
  summarise(n=length(counts_volume),
            Mean_sym_mm3=format(round(mean(counts_volume), 0), 0), 
            SE=format(round(sd(counts_volume)/sqrt(length(counts_volume)),0),0))%>%
  rename(Lifestage=group, HPF=hpf)%>%
  write_csv(., "Mcap2020/Output/Physiology/normalized_size_cells_summary.csv")
```

### Attached Recruit Plotting  

Plot as box plot for cells per mm^2.    

```{r}
recruit_size_plot<-corr %>%
    filter(group=="Attached Recruit")%>%
    ggplot(., aes(x = lifestage, y = counts_volume)) +
    #geom_boxplot(aes(color=group), outlier.size = 0, lwd=1) +
    geom_point(aes(fill=group), pch = 21, size=4, position = position_jitterdodge(0.8)) + 
    xlab("") + 
    scale_fill_manual(name="Lifestage", values=c("#BA55D3"), guide="none")+
    scale_color_manual(name="Lifestage", values=c("#BA55D3"))+
    ylab(expression(bold(paste("Symbiont cells (mm"^-2,")"))))+
    ylim(500, 10000)+
    theme_classic() + 
    theme(
      legend.position="top",
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      axis.text.x=element_text(angle=45, hjust=1)
      ); recruit_size_plot
```

View summary of symbiont densities per unit surface area.  

```{r}
#need to output to csv 
corr%>%
  group_by(group, hpf, lifestage)%>%
  filter(group=="Attached Recruit")%>%
  droplevels()%>%
  summarise(n=length(counts_volume),
            Mean_sym_mm2=format(round(mean(counts_volume), 0), 0), 
            SE=format(round(sd(counts_volume)/sqrt(length(counts_volume)),0),0))%>%
  rename(Lifestage=group, HPF=hpf)%>%
  write_csv(., "Mcap2020/Output/Physiology/normalized_size_cells_summary_attached.csv")
```

Test differences with a kruskal wallis test.  

```{r}
corr%>%
  group_by(group, hpf, lifestage)%>%
  filter(group=="Attached Recruit")%>%
  droplevels()%>%
  kruskal.test(counts_volume~lifestage, data=.)
```

Cell density is not significantly different by life stage in attached recruits, likely due to low sample size.  

# **3. Generate Figures**  

Generate a figure that has the symbiont density metrics of interest: counts per individual and symbionts per mm2 for attached recruits. 

```{r}
# extract the legend from one of the plots
legend <- get_legend(
  # create some space to the left of the legend
  legend_plot +  theme(legend.box.margin = margin(1,1,1,1))
)

symb_plot2b<-symb_plot2+theme(legend.position="none")+ theme(axis.title=element_text(size=14, face="bold"))
recruit_size_plot2<-recruit_size_plot+theme(legend.position="none")+theme(axis.title=element_text(size=14))+ylim(0,10000)

#target_plots<- ggdraw(symb_plot2b) +
#  draw_plot(recruit_size_plot2, .16, .6, .35, .35) +
#  draw_plot_label(
#    c("A", "B"),
#    c(0.05, 0.2),
#    c(1, 0.96),
#    size = 16
#  );target_plots

target_plots_legend<-plot_grid(symb_plot2b, recruit_size_plot2, legend, rel_widths = c(1,0.5,0.5), rel_heights=c(1,0.8,1), ncol=3, nrow=1, align="h", axis = "b", labels=c("A", "B", ""))

ggsave(file="Mcap2020/Figures/Physiology/Physiology_figure_densities.png", target_plots_legend, dpi=300, width=12, height=6, units="in")

```


# **4. Generate relative change dataframe**  

## Symbiont density  

Read in data from calculated densities.  

```{r}
den_relchange<-read_csv("Mcap2020/Output/Physiology/calculated_densities.csv")
```

Calculate relative change. Calculate relative change for all lifestages compared to the first time point in this dataset (eggs). Do this for both individual normalized and surface area normalized for all stages. Use NA for stages without this normalization.     

```{r}
time_point_order = c("Egg (1 hpf)", "Embryo (5 hpf)", "Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf)", "Larvae (163 hpf)", "Larvae (183 hpf)", "Larvae (231 hpf)", "Metamorphosed Polyp (183 hpf)", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf)", "Attached Recruit (231 hpf)", "Attached Recruit (255 hpf)") #Set time_point order
den_relchange$lifestage<-factor(den_relchange$lifestage, levels=time_point_order)

cells_rel_change<-den_relchange%>%
  dplyr::select(lifestage, tube.ID, cells.ind, cells.mm)%>%
  group_by(lifestage)%>%
  summarise(mean_cells.in=mean(cells.ind), mean_cells.mm=mean(cells.mm))%>%
  mutate(group=if_else(grepl("Attached", .$lifestage), "surface area", "individual"))%>%
  group_by(lifestage, group)%>%
  arrange(lifestage, .by_group = TRUE)%>%
  group_by(group)%>%
  mutate(sym_density_ind=((mean_cells.in-dplyr::first(mean_cells.in))/abs(dplyr::first(mean_cells.in))), 
         sym_density_mm=((mean_cells.mm-dplyr::first(mean_cells.mm))/abs(dplyr::first(mean_cells.mm))))%>%
  dplyr::select(lifestage, sym_density_ind, sym_density_mm)%>%
  write_csv("Mcap2020/Data/Integration/RelativeChange/cells_relchange.csv")
```


## Larval size  

Read in data from calculated volume  

```{r}
size_relchange<-read_csv("Mcap2020/Output/Physiology/larval_volume_table.csv")
levels(as.factor(size_relchange$Lifestage))
```

Calculate relative change. Calculate relative change for all lifestages compared to the first time point in this dataset (eggs). 

```{r}
time_point_order = c("Egg (1 hpf)", "Embryo (5 hpf)", "Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf)", "Larvae (163 hpf)", "Larvae (183 hpf)", "Larvae (231 hpf)", "Metamorphosed Polyp (183 hpf)", "Metamorphosed Polyp (231 hpf)") #Set time_point order
size_relchange$Lifestage<-factor(size_relchange$Lifestage, levels=time_point_order)

size_rel_change<-size_relchange%>%
  dplyr::select(Lifestage, Mean)%>%
  mutate(size_change=((Mean-dplyr::first(Mean))/dplyr::first(Mean)))%>%
  dplyr::select(Lifestage, size_change)%>%
  write_csv("Mcap2020/Data/Integration/RelativeChange/size_relchange.csv")
```



