---
title: "Symbiont density analysis - Mcap 2020"
output: html_document
---

```{r setup, include = FALSE}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("RColorBrewer")) install.packages("RColorBrewer")
if (!require("car")) install.packages("car")
if (!require("emmeans")) install.packages("emmeans")
if (!require("multcomp")) install.packages("multcomp")

# load packages
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(car)
library(emmeans)
library(multcomp)
```


# Import data
```{r}
# Cell count data
sym_counts <- read_csv("Mcap2020/Data/Physiology/CellDensity/symbiont.counts.csv")

```

# Calculate cells per square centimeter
```{r}
# Calculate mean counts for each sample
df <- sym_counts %>%
  select(tube.ID, num.squares, matches("Count[1-6]")) %>%
  gather("rep", "count", -tube.ID, -num.squares) %>%
  group_by(tube.ID, num.squares) %>%
  summarise(mean_count = mean(count, na.rm = TRUE))

#match in identifying information
df$lifestage<-sym_counts$lifestage[match(df$tube.ID, sym_counts$tube.ID)]
df$total.volume.ul<-sym_counts$total.volume.ul[match(df$tube.ID, sym_counts$tube.ID)]
df$num.individuals<-sym_counts$num.individuals[match(df$tube.ID, sym_counts$tube.ID)]

# Normalize counts by homogenat volume and surface area
df <- df %>%
  mutate(cells.mL = mean_count * 10000 / num.squares,
         cells = cells.mL * (total.volume.ul/1000),
         cells.ind = cells / num.individuals)
```

# Plot data with mean and standard error
```{r}
df %>%
  #filter(!is.na(site)) %>%
  ggplot(aes(x = lifestage, y = cells.ind, color = lifestage)) +
  labs(x = "",y = "Cell Density per larva") +
geom_jitter(width = 0.1) +                                            # Plot all points
  stat_summary(fun.data = mean_cl_normal, fun.args = list(mult = 1),    # Plot standard error
               geom = "errorbar", color = "black", width = 0.5) +
  stat_summary(fun.y = mean, geom = "point", color = "black") + # Plot mean
  theme_classic()
```

# Plot data as box plot  
```{r}
symb_plot<-df %>%
    ggplot(., aes(x = lifestage, y = cells.ind, fill = lifestage)) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    xlab("Lifestage") + 
    scale_fill_brewer()+
    ylab(expression(bold(paste("Symbiont cells individual"^-1))))+
    ylim(0,9000)+
    theme_classic() + 
    geom_text(label="A", x=1, y=2500, size=5, color="black")+ #egg
    geom_text(label="A", x=2, y=2500, size=5, color="black")+ #embryo 1
    geom_text(label="A", x=3, y=2500, size=5, color="black")+ #larvae 1
    geom_text(label="AB", x=4, y=4000, size=5, color="black")+ #larvae 2
    geom_text(label="AB", x=5, y=4000, size=5, color="black")+ #larvae 3
    geom_text(label="AB", x=6, y=4000, size=5, color="black")+ #larvae 4
    geom_text(label="BC", x=7, y=4500, size=5, color="black")+ #larvae 4
    geom_text(label="D", x=8, y=6500, size=5, color="black")+ #larvae 6
    geom_text(label="CD", x=9, y=6500, size=5, color="black")+ #recruit1
    geom_text(label="D", x=10, y=8700, size=5, color="black")+ #recruit2
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black")
      ); symb_plot

ggsave("Mcap2020/Figures/Physiology/SymbiontCounts.pdf", symb_plot, dpi=300, w=8.5, h=5, units="in")
```

# Statistical analysis  

Run ANOVA on cells per larvae by timestage.  

```{r}
model<-aov(cells.ind~lifestage, data=df)
qqPlot(residuals(model))
leveneTest(residuals(model)~lifestage, data=df)
```

Both normality and homogeneity of variance pass. View summary and p-values.  

```{r}
summary(model)
```

There is a significant effect of lifestage on cell densities. View posthoc comparisons for differences between lifestages.  

```{r}
emm = emmeans(model, ~ lifestage) 
cld(emm, Letters=c(LETTERS)) #letter display
```

Output data to file.  

```{r}
df %>%
  write_csv(., file = "Mcap2020/Output/Physiology/calculated_densities.csv")
```

