---
title: "Symbiont density analysis - Mcap 2020"
author: "AS Huffmyer"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

# Set Up  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r setup, include = FALSE}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("RColorBrewer")) install.packages("RColorBrewer")
if (!require("car")) install.packages("car")
if (!require("emmeans")) install.packages("emmeans")
if (!require("multcomp")) install.packages("multcomp")
if (!require("knitr")) install.packages("knitr")
if (!require("kableExtra")) install.packages("kableExtra")

# load packages
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(car)
library(emmeans)
library(multcomp)
library(knitr)
library(kableExtra)
```


# Import and manipulate data
```{r}
# Cell count data
sym_counts <- read_csv("Mcap2020/Data/Physiology/CellDensity/symbiont.counts.csv")

#metadata
metadata <- read_csv("Mcap2020/Data/lifestage_metadata.csv")

sym_counts <- left_join(sym_counts, metadata)
sym_counts$hpf <- as.factor(sym_counts$hpf)
```

## Calculate cells per square centimeter
```{r}
# Calculate mean counts for each sample
df <- sym_counts %>%
  dplyr::select(tube.ID, num.squares, matches("count[1-6]")) %>%
  gather("rep", "count", -tube.ID, -num.squares) %>%
  group_by(tube.ID, num.squares) %>%
  summarise(mean_count = mean(count, na.rm = TRUE))

#match in identifying information
df$lifestage<-sym_counts$lifestage[match(df$tube.ID, sym_counts$tube.ID)]
df$total.volume.ul<-sym_counts$total.volume.ul[match(df$tube.ID, sym_counts$tube.ID)]
df$num.individuals<-sym_counts$num.individuals[match(df$tube.ID, sym_counts$tube.ID)]
df$surface.area<-sym_counts$surface.area[match(df$tube.ID, sym_counts$tube.ID)]
df$hpf<-sym_counts$hpf[match(df$tube.ID, sym_counts$tube.ID)]
df$group<-sym_counts$group[match(df$tube.ID, sym_counts$tube.ID)]
df$lifestage<-as.factor(df$lifestage)
df$group<-as.factor(df$group)

# Normalize counts by homogenat volume and surface area
df <- df %>%
  mutate(cells.mL = mean_count * 10000 / num.squares,
         cells = cells.mL * (total.volume.ul/1000),
         cells.ind = cells / num.individuals, 
         cells.mm = cells / surface.area)
```

# Plotting  

Plot data with mean and standard error for larvae through metamorphosis (these counts have cells/individual). Plot attached recruits separately, these values are in cells per mm2. We will plot cells per unit surface area for all stages in later analyses. 
```{r}
df %>%
  #filter(!is.na(site)) %>%
  filter(!group=="Attached Recruit")%>%
  ggplot(aes(x = lifestage, y = cells.ind, color = lifestage)) +
  labs(x = "",y = "Cell Density per larva") +
geom_jitter(width = 0.1) +                                            # Plot all points
  stat_summary(fun.data = mean_cl_normal, fun.args = list(mult = 1),    # Plot standard error
               geom = "errorbar", color = "black", width = 0.5) +
  stat_summary(fun.y = mean, geom = "point", color = "black") + # Plot mean
  theme_classic()
```

```{r}
df %>%
  #filter(!is.na(site)) %>%
  filter(group=="Attached Recruit")%>%
  ggplot(aes(x = lifestage, y = cells.mm, color = lifestage)) +
  labs(x = "",y = "Cell Density per mm2") +
geom_jitter(width = 0.1) +                                            # Plot all points
  #stat_summary(fun.data = mean_cl_normal, fun.args = list(mult = 1),    # Plot standard error
               #geom = "errorbar", color = "black", width = 0.5) +
  #stat_summary(fun.y = mean, geom = "point", color = "black") + # Plot mean
  theme_classic()
```

Present means and standard error of each group and save summary table

```{r}
df%>%
  group_by(group, hpf, lifestage)%>%
  summarize(n=length(cells.ind),
            Mean=format(round(mean(cells.ind), 0), 0), 
            SE=format(round(sd(cells.ind)/sqrt(length(cells.ind)),0),0))%>%
  rename(Lifestage=group, HPF=hpf)%>%
  kbl(caption="Descriptive statistics of Symbiodiniaceae cell densities per larva across ontogeny")%>%
  kable_classic(full_width=FALSE, html_font="Arial")%>%
  row_spec(0, bold = TRUE) 

df%>%
  group_by(group, hpf, lifestage)%>%
  summarize(n=length(cells.mm),
            Mean=format(round(mean(cells.mm), 0), 0), 
            SE=format(round(sd(cells.mm)/sqrt(length(cells.mm)),0),0))%>%
  rename(Lifestage=group, HPF=hpf)%>%
  kbl(caption="Descriptive statistics of Symbiodiniaceae cell densities per mm2 across ontogeny")%>%
  kable_classic(full_width=FALSE, html_font="Arial")%>%
  row_spec(0, bold = TRUE) 

#need to output to csv 
df%>%
  group_by(group, hpf, lifestage)%>%
  summarize(n=length(cells.ind),
            Mean=format(round(mean(cells.ind), 0), 0), 
            SE=format(round(sd(cells.ind)/sqrt(length(cells.ind)),0),0))%>%
  rename(group=group, HPF=hpf)%>%
  write_csv(., "Mcap2020/Output/Physiology/cell_density_table.csv")
```



Plot data as a scatterplot   
```{r}
df$code<-paste(df$lifestage, "\n (", df$hpf, ")")
df$code <- factor(df$code,levels = c("Egg Fertilized \n ( 1 )", "Embryo 1 \n ( 5 )", "Larvae 1 \n ( 38 )", "Larvae 2 \n ( 65 )", "Larvae 3 \n ( 93 )", "Larvae 4 \n ( 163 )", "Larvae 5 \n ( 183 )", "Recruit 1 \n ( 183 )", "Larvae 6 \n ( 231 )", "Recruit 2 \n ( 231 )", "Plug 1 \n ( 183 )", "Plug 2 \n ( 231 )", "Plug 3 \n ( 255 )"))
df$hpf<-as.numeric(as.character(df$hpf))

symb_plot<-df %>%
    ggplot(., aes(x = hpf, y = cells.ind)) +
    #geom_boxplot(outlier.size = 0) +
    geom_smooth(method="loess", se=TRUE, fullrange=TRUE, level=0.95, color="black") +
    geom_point(aes(fill=group, group=group), pch = 21, size=4, position = position_jitterdodge(5)) + 
    xlab("Hours Post-Fertilization") + 
    scale_fill_manual(name="Lifestage", values=c("#BA55D3", "#8C510A", "#DFC27D","#80CDC1", "#003C30"))+
    ylab(expression(bold(paste("Symbiont cells individual"^-1))))+
    ylim(0,9000)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); symb_plot

#EGG: #8C510A
#EMBRYO: #DFC27D
#LARVAE: #80CDC1
#RECRUIT: #003C30
#ATTACHED: 

#COLORS TO TRY FROM BRBG: "#543005" "#8C510A" "#BF812D" "#DFC27D" "#F6E8C3" "#F5F5F5" "#C7EAE5" "#80CDC1" "#35978F" "#01665E" "#003C30"

#ggsave("Mcap2020/Figures/Physiology/SymbiontCounts.png", symb_plot, dpi=300, w=8.5, h=5, units="in")
```

Plot data as box plot  
```{r}

symb_plot2<-df %>%
    ggplot(., aes(x = as.factor(hpf), y = cells.ind)) +
    geom_boxplot(aes(color=group), outlier.size = 0, lwd=1) +
    #geom_smooth(method="loess", se=TRUE, fullrange=TRUE, level=0.95, color="black") +
    geom_point(aes(fill=group), pch = 21, size=4, position = position_jitterdodge(0.2)) + 
    xlab("Hours Post-Fertilization)") + 
    scale_fill_manual(name="Lifestage", values=c("#BA55D3", "#8C510A", "#DFC27D","#80CDC1", "#003C30"))+
    scale_color_manual(name="Lifestage", values=c("#8C510A", "#DFC27D","#80CDC1", "#003C30"), guide=FALSE)+
    ylab(expression(bold(paste("Symbiont cells individual"^-1))))+
    ylim(0,9000)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14)
      ); symb_plot2

ggsave("Mcap2020/Figures/Physiology/SymbiontCounts_Box.png", symb_plot2, dpi=300, w=8.5, h=5, units="in")
```

# Statistical analysis  

Run ANOVA on cells per larvae by sampling point, specified by sequence of samples taken (life stage, hpf).    

```{r}
model<-aov(cells.ind~lifestage, data=df)
qqPlot(residuals(model))
leveneTest(residuals(model)~lifestage, data=df)
```

Both normality and homogeneity of variance pass. View summary and p-values and save model output table.  

```{r}
output<-anova(model)
output

output.new <- output[,c("Sum Sq", "Df", "Mean Sq", "F value", "Pr(>F)")]
names(output.new) <- c("SS", "df", "MS", "F", "P")
totals <- colSums(output.new)
output.anova.table <- as.data.frame(rbind(output.new, totals), 
                      row.names = c("Treatment", "Error", "Total"))

output.anova.table$Model.Component<-c("Treatment", "Error", "Total")
output.anova.table <- output.anova.table[,c("Model.Component", "SS", "df", "MS", "F", "P")]

output.anova.table%>%
  write_csv(., "Mcap2020/Output/Physiology/cell_density_model.csv")


output.anova.table%>%
  kbl(caption="ANOVA analysis of Symbiodiniaceae cell densities across ontogeny", format="html", digits = c(3, 0, 0, 2, 11))%>%
  kable_classic(full_width=FALSE, html_font="Arial")


```

There is a significant effect of lifestage on cell densities. View posthoc comparisons for differences between lifestages.  

```{r}
emm = emmeans(model, ~ lifestage)
cld(emm, Letters=c(LETTERS)) #letter display
```

Output data to file.  

```{r}
df %>%
  write_csv(., file = "Mcap2020/Output/Physiology/calculated_densities.csv")
```

