---
title: Montipora capitata 2020 ITS2 analysis
author: "AS Huffmyer"
date: '2021'
output:
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

ITS2 strain variation data analysis based on scripts written by Shayle Matsuda, Shedd Aquarium.   

```{r}
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("readxl")) install.packages("readxl")
if (!require("phyloseq")) install.packages("phyloseq")
if (!require("janitor")) install.packages("janitor")
if (!require("writexl")) install.packages("writexl")

library("tidyverse")
library("readxl")
library("phyloseq")
library("janitor")
library("writexl")

library("ggplot2")
library('cowplot')
library("tidyverse")
library("janitor")
library("car")
library("phyloseq")
library("stringr")
library("ape")
library("DESeq2")
library("microViz")
library("vegan")
```

# **Prepare data**  

Load metadata.  
```{r}
#create metadata and load
Metadata<-read.csv("Mcap2020/Data/ITS2/ITS2_Metadata.csv")
rownames(Metadata) <- Metadata$sample_name
Metadata <- Metadata %>%
  select(code, hpf, group, lifestage)
Metadata <- as.matrix(Metadata)
Metadata <- sample_data(data.frame(Metadata))
```

Load profile level data.  
```{r}
#load data for taxonomy
tax0 <- read_tsv(
  file  = "Mcap2020/Data/ITS2/mcap.profiles.absolute.abund_and_meta.txt",
  n_max = 6) %>%
  dplyr::select(-2) %>% 
  gather(UID, value, -1) %>% 
  spread(1, value) %>%
  clean_names()

tax1 <- as.matrix(tax0[, -1], dimnames = list(tax0$uid, colnames(tax0[-1])))
rownames(tax1) <- tax0$uid
tax <- tax_table(tax1)

#load data for count matrix 
otu0 <- read_tsv(
  file  = "Mcap2020/Data/ITS2/mcap.profiles.absolute.abund_and_meta.txt", col_names = TRUE) 

list<-c("UID", "sample_name", "8", "9")
colnames(otu0)<-list

otu0<-otu0%>%
  select(-1) %>%
  dplyr::slice(7:n()) %>%
  mutate_at(2:ncol(.), as.numeric)

otu1 <- as.matrix(otu0[, -1])
rownames(otu1) <- otu0$sample_name
otu <- otu_table(otu1, taxa_are_rows = FALSE)

#combine to phyloseq object
coral <- phyloseq(otu, tax, Metadata)
coral
```

Note that taxa 8 is CladeC and taxa 9 is CladeD.  

Next load the data for coral strain level information.  

```{r}
#gather the taxonomy names  
taxnames <- read_tsv(
  file  = "Mcap2020/Data/ITS2/mcap.seqs.absolute.abund_only.txt",
  n_max = 0) %>%
  select(-1) %>%
  names(.)

#extract clade letter  
tax0 <- data_frame(
  DIV = taxnames,
  clade = str_extract(DIV, "[A-Z]")
)

tax1 <- as.matrix(tax0)
rownames(tax1) <- tax0$DIV
tax <- tax_table(tax1)

otu0 <- read_tsv(
  file  = "Mcap2020/Data/ITS2/mcap.seqs.absolute.abund_only.txt") %>%
  select(-1, )
otu1 <- as.matrix(otu0[, -1])
rownames(otu1) <- Metadata$sample_name
otu <- otu_table(otu1, taxa_are_rows = FALSE)

coralDIV <- phyloseq(otu, tax, Metadata)
coralDIV
```

Now save the object.  
```{r} 
save(coral, coralDIV, file = "Mcap2020/Data/ITS2/ITS2_phyloseq.RData")
```

# **Explore DIV Data**  

Load the data.  
```{r}
load("Mcap2020/Data/ITS2/ITS2_phyloseq.RData")
```

```{r}
coralDIV
```

View the number of sequence counts in our dataset.   
```{r}
Samps<-read_tsv(file  = "Mcap2020/Data/ITS2/mcap.seqs.absolute.abund_and_meta.txt")

hist(Samps$post_taxa_id_absolute_symbiodiniaceae_seqs, breaks=30)

```
We have >10,000 sequences for all of our samples. This is a high number and we can keep all samples.  

Plot tree.  
```{r}
random_tree = rtree(ntaxa(coralDIV), rooted=TRUE, tip.label=taxa_names(coralDIV))
plot(random_tree)
```

Merge phyloseq data. 
```{r}
phylo_coral_DIV = merge_phyloseq(coralDIV, Metadata, random_tree)
phylo_coral_DIV
```

Filter out samples with low numbers in groups from phyloseq object.  
```{r}
phylo_coral_DIV<-ps_filter(phylo_coral_DIV, lifestage != "Metamorphosed Recruit (183 hpf)")
phylo_coral_DIV<-ps_arrange(phylo_coral_DIV, lifestage)
```

Now build a tree with the combined data.  
```{r}
plot_tree(phylo_coral_DIV, color="group", label.tips="taxa_names", ladderize="left", plot.margin=0.3)
```

Plot a heatmap.  

```{r}
plot_heatmap(phylo_coral_DIV, "NMDS", "bray", "lifestage", low="#66CCFF", high="#000033", na.value="white")
```

Plot ordinations.  

```{r}
GP1 = phylo_coral_DIV
GP1
```

Transform to relative abundance.  
```{r}
GP1 = transform_sample_counts(GP1, function(x) 1E6 * x/sum(x))
```

View ordination by OTU.  
```{r}
GP.ord <- ordinate(GP1, "NMDS", "bray")
p1 = plot_ordination(GP1, GP.ord, type="taxa", title="taxa")
p1
```

Plot with samples now.  
```{r}
p2 = plot_ordination(GP1, GP.ord, type="samples", color="lifestage") 
p2<- p2 + geom_polygon(aes(fill=lifestage)) + geom_point(size=4) + ggtitle("samples")
p2
```

Conduct a permanova in the vegan package.  

```{r}
metadata <- as(sample_data(GP1), "data.frame")

adonis(phyloseq::distance(GP1, method="bray") ~ lifestage,
       data = metadata)
```

Permanova suggests there is separation by lifestage in ITS2 profiles.  

Look at alpha diversity.  

Remove any OTU that has a count of 0 in all samples. (There are none in this dataset).    
```{r}
GP2f <- prune_taxa(taxa_sums(phylo_coral_DIV) > 0, phylo_coral_DIV)
```

Plot the two most commonly used richness metrics - Shannon and Inverse Simpson.  
```{r}
plot_richness(GP2f, measures=c("InvSimpson", "Shannon"), x="lifestage", color="group")
```

Pretty similar across life stages.     

Test for differences in alpha diversity.  
```{r}
metadata <- as(sample_data(GP2f), "data.frame")
metadata$sample<-rownames(metadata)

alpha_full<-estimate_richness(GP2f, measures=c("InvSimpson", "Shannon"))
alpha_full$sample<-rownames(alpha_full)

alpha_full<-full_join(metadata, alpha_full, by="sample")

summary(aov(Shannon~lifestage, data=alpha_full))
summary(aov(InvSimpson~lifestage, data=alpha_full))
```

No difference in ITS2 alpha diversity by lifestage.  

Try DESeq2 to look at drivers of differences.  

Convert phyloseq object to DESeq2 object.  

Construct DESeq2 object looking at differences by lifestage.  
```{r}
diagdds_full = phyloseq_to_deseq2(phylo_coral_DIV, ~ lifestage)

gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}
geoMeans = apply(counts(diagdds_full), 1, gm_mean)
diagdds_full = estimateSizeFactors(diagdds_full, geoMeans = geoMeans)
#diagdds_full = DESeq(diagdds_full, fitType="local")

diagdds_full = DESeq(diagdds_full, test="Wald", fitType="parametric")
```

Look at results table.  
```{r}
res_full = results(diagdds_full, cooksCutoff = FALSE)
alpha = 0.05
sigtab = res_full[which(res_full$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(phylo_coral_DIV)[rownames(sigtab), ], "matrix"))
head(sigtab)
length(sigtab)
```

There are 6 members of the community that shift by lifestage, 5 in clade C and 1 in clade D.  

```{r}
sigtab
sig_taxa<-rownames(sigtab)
```

FIGURE OUT HOW TO CONVERT TO RELATIVE ABUNDANCE? 

Plot the abundance of these taxa over each lifestage using the GP1 object. 

```{r}
sig_coral_DIV = prune_taxa(sig_taxa, GP1)
plot_bar(sig_coral_DIV, fill = "lifestage", facet_grid=~DIV)
```

