---
title: Montipora capitata 2020 ITS2 analysis
author: "AS Huffmyer"
date: '2023'
output:
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

ITS2 strain variation data analysis based on scripts from Shayle Matsuda, Shedd Aquarium.   

```{r}
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("readxl")) install.packages("readxl")
if (!require("phyloseq")) BiocManager::install("phyloseq")
if (!require("janitor")) install.packages("janitor")
if (!require("writexl")) install.packages("writexl")
#if (!require("microViz")) BiocManager::install("microViz")

library("readxl")
library("phyloseq")
library("janitor")
library("writexl")
library("ggplot2")
library('cowplot')
library("janitor")
library("car")
library("phyloseq")
library("stringr")
library("ape")
library("DESeq2")
library("microViz")
library("vegan")
library("ComplexHeatmap")
library("tidyverse")
library("dplyr")
library(emmeans)
```

# **Prepare data**  

Load metadata.  
```{r}
#create metadata and load
Metadata<-read.csv("Mcap2020/Data/ITS2/ITS2_Metadata.csv")
rownames(Metadata) <- Metadata$sample_name
Metadata <- Metadata %>%
  select(code, hpf, group, lifestage)
Metadata <- as.matrix(Metadata)
Metadata <- sample_data(data.frame(Metadata))
```

Load profile level data.  
```{r}
#load data for taxonomy
tax0 <- read_tsv(
  file  = "Mcap2020/Data/ITS2/mcap.profiles.absolute.abund_and_meta.txt",
  n_max = 6) %>%
  dplyr::select(-2) %>% 
  gather(UID, value, -1) %>% 
  spread(1, value) %>%
  clean_names()

tax1 <- as.matrix(tax0[, -1], dimnames = list(tax0$uid, colnames(tax0[-1])))
rownames(tax1) <- tax0$uid
tax <- tax_table(tax1)

#load data for count matrix 
otu0 <- read_tsv(
  file  = "Mcap2020/Data/ITS2/mcap.profiles.absolute.abund_and_meta.txt", col_names = TRUE) 

list<-c("UID", "sample_name", "8", "9")
colnames(otu0)<-list

otu0<-otu0%>%
  select(-1) %>%
  dplyr::slice(7:n()) %>%
  mutate_at(2:ncol(.), as.numeric)

otu1 <- as.matrix(otu0[, -1])
rownames(otu1) <- otu0$sample_name
otu <- otu_table(otu1, taxa_are_rows = FALSE)

#combine to phyloseq object
coral <- phyloseq(otu, tax, Metadata)
coral
```

Note that taxa 8 is Clade C and taxa 9 is Clade D.  

Next load the data for coral strain level information.  

```{r}
#gather the taxonomy names  
taxnames <- read_tsv(
  file  = "Mcap2020/Data/ITS2/mcap.seqs.absolute.abund_only.txt",
  n_max = 0) %>%
  select(-1) %>%
  names(.)

#extract clade letter  
tax0 <- data_frame(
  DIV = taxnames,
  clade = str_extract(DIV, "[A-Z]")
)

tax1 <- as.matrix(tax0)
rownames(tax1) <- tax0$DIV
tax <- tax_table(tax1)

otu0 <- read_tsv(
  file  = "Mcap2020/Data/ITS2/mcap.seqs.absolute.abund_and_meta.txt") %>%
  select(-1, )

otu1 <- as.matrix(otu0[, 37:113])
rownames(otu1) <- otu0$sample_name
otu <- otu_table(otu1, taxa_are_rows = FALSE)

coralDIV <- phyloseq(otu, tax, Metadata)
coralDIV
```

Now save the object.  
```{r} 
save(coral, coralDIV, file = "Mcap2020/Data/ITS2/ITS2_phyloseq.RData")
```

# **Build DIV Data**  

Load the data.  
```{r}
load("Mcap2020/Data/ITS2/ITS2_phyloseq.RData")
```

```{r}
coralDIV
```

View the number of sequence counts in our dataset.   
```{r}
Samps<-read_tsv(file  = "Mcap2020/Data/ITS2/mcap.seqs.absolute.abund_and_meta.txt")

hist(Samps$post_taxa_id_absolute_symbiodiniaceae_seqs, breaks=30)

Samps$post_taxa_id_absolute_symbiodiniaceae_seqs
```
We have >10,000 sequences for all of our samples. This is a high number and we can keep all samples. We will subsample to even sampling depth below. The lowest value is 11,265.

Plot tree.  
```{r}
random_tree = rtree(ntaxa(coralDIV), rooted=TRUE, tip.label=taxa_names(coralDIV))
plot(random_tree)
```

Merge phyloseq data. 
```{r}
phylo_coral_DIV = merge_phyloseq(coralDIV, Metadata, random_tree)
phylo_coral_DIV
```

Filter out samples with low numbers in groups (n=1 per group) from phyloseq object.  
```{r}
phylo_coral_DIV<-ps_filter(phylo_coral_DIV, lifestage != "Metamorphosed Polyp (183 hpf)")
phylo_coral_DIV<-ps_arrange(phylo_coral_DIV, lifestage)
```

Reorder lifestage.  
```{r}
levels(as.factor(sample_data(phylo_coral_DIV)$lifestage))

sample_data(phylo_coral_DIV)$lifestage = factor(sample_data(phylo_coral_DIV)$lifestage, levels = c("Egg (1 hpf)","Embryo (5 hpf)","Embryo (38 hpf)","Embryo (65 hpf)", "Larvae (93 hpf)", "Larvae (163 hpf)", "Larvae (231 hpf)", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf)", "Attached Recruit (231 hpf)"))

levels(as.factor(sample_data(phylo_coral_DIV)$lifestage))
```

Now build a tree with the combined data.  
```{r}
plot_tree(phylo_coral_DIV, color="lifestage", label.tips="taxa_names", ladderize="left", plot.margin=0.3)
```

Plot a heatmap.  

```{r}
plot_heatmap(phylo_coral_DIV, "NMDS", "bray", "lifestage", low="#66CCFF", high="#000033", na.value="white")
```

Assign object to a shorter name. 

```{r}
GP1 = phylo_coral_DIV
GP1
```

Display smaple sequencing depth. 
```{r}
min(sample_sums(GP1))
```

Transform to relative abundance and subsample to even sampling depth to minimum count in dataset (10,551).  
```{r}
GP1_rare<-rarefy_even_depth(GP1, sample.size = min(sample_sums(GP1)),
  rngseed = TRUE, replace = TRUE, trimOTUs = TRUE, verbose = TRUE)
```

View sample sequencing depth again after subsampling. 
```{r}
sample_sums(GP1_rare)
```

All samples are now at 10,551 sequences. No OTU's removed. 

Now transform counts to relative abundance expressed as a proportion.    

```{r}
#GP1_rare = transform_sample_counts(GP1_rare, function(x) 100 * x/sum(x))
GP1_rare = transform_sample_counts(GP1_rare, function(x) x/sum(x))
```

View tree and heatmap after subsampling.   
```{r}
plot_tree(GP1_rare, color="lifestage", label.tips="taxa_names", ladderize="left", plot.margin=0.3)
plot_heatmap(GP1_rare, "NMDS", "bray", "lifestage", low="#66CCFF", high="#000033", na.value="white")
```

# **DIV Ordinations: NMDS** 

View NMDS ordination by OTU.    
```{r}
GP.ord <- ordinate(GP1_rare, "NMDS", "bray")
ordinate(GP1_rare, "NMDS", "bray")
p1 = plot_ordination(GP1_rare, GP.ord, type="taxa", title="taxa")
p1
```

Plot with samples now.  
```{r}
p2 = plot_ordination(GP1_rare, GP.ord, type="samples", color="lifestage") 
p2_geom<- p2 + geom_polygon(aes(fill=lifestage)) + geom_point(size=4) + ggtitle("samples")
p2_geom
```

View PCoA by OTU.  
```{r}
GP.ordb.pcoa <- ordinate(GP1_rare, "PCoA", "bray")
p1b = plot_ordination(GP1_rare, GP.ordb.pcoa, type="taxa", title="taxa")
p1b
```

Plot with samples now.  
```{r}
p2b = plot_ordination(GP1_rare, GP.ord, type="samples", color="lifestage"); p2b 
```

Output NMDS plot for visualization.  
```{r}
p2b_viz<-p2  + 
  geom_point(aes(colour=lifestage, shape=lifestage), size=4) + 
  ggtitle("NMDS")+
  scale_colour_manual(name="Lifestage", values=c("#8C510A", "#DFC27D","#DFC27D", "#DFC27D", "#80CDC1", "#80CDC1", "#80CDC1","#003C30", "#BA55D3", "#BA55D3")) + 
  scale_shape_manual(name="Lifestage", values = c("Egg (1 hpf)"=1, "Attached Recruit (231 hpf)"=13, "Embryo (5 hpf)"=2, "Embryo (38 hpf)"=3, "Embryo (65 hpf)"=4, "Larvae (93 hpf)"=5, "Larvae (163 hpf)"=6, "Larvae (231 hpf)"=8, "Metamorphosed Polyp (231 hpf)"=11, "Attached Recruit (183 hpf)"=12)) +
  theme_classic() + 
  theme(
    panel.background=element_blank()
  ); p2b_viz

ggsave(plot=p2b_viz, "Mcap2020/Figures/ITS2/nmds_its2_div.jpeg", h=5, w=8)
ggsave(plot=p2b_viz, "Mcap2020/Figures/ITS2/nmds_its2_div.pdf", h=5, w=8)
```

Conduct a permanova in the vegan package.  

```{r}
metadata <- as(sample_data(GP1_rare), "data.frame")

perm1<-adonis2(phyloseq::distance(GP1_rare, method="bray") ~ lifestage,
       data = metadata)

perm1
```

Permanova suggests there is significant separation by lifestage in ITS2 profiles (p=0.001).  

Permutation test for adonis under reduced model
Terms added sequentially (first to last)
Permutation: free
Number of permutations: 999

adonis2(formula = phyloseq::distance(GP1_rare, method = "bray") ~ lifestage, data = metadata)
          Df SumOfSqs      R2      F Pr(>F)    
lifestage  9 0.090070 0.48981 2.9868  0.001 ***
Residual  28 0.093818 0.51019                  
Total     37 0.183888 1.00000     

# **Plot heat map of ITS2 DIVs**  

Draw with Phyloseq.  

Filter out taxa that comprise <1% mean relative abundance across samples. 

There were 77 OTU's before filtering and 11 after filtering. This indicates there were many rare taxa.   
```{r}
GP1_rare

GP1_prune = filter_taxa(GP1_rare, function(x) mean(x) > 0.01, TRUE)
GP1_prune 

sample_order<-c("WSH053", "WSH054", "WSH055", "WSH056", "WSH061", "WSH062", "WSH063", "WSH064",  "WSH065", "WSH066", "WSH067", "WSH068",  "WSH069", "WSH070", "WSH071", "WSH072", "WSH073", "WSH074", "WSH075", "WSH076", "WSH077", "WSH078", "WSH079", "WSH080", "WSH081", "WSH082", "WSH083", "WSH084", "WSH057", "WSH058", "WSH059", "WSH060", "WSH046", "WSH047", "WSH048", "WSH049", "WSH050", "WSH051", "WSH052")
#sample_name by lifestage order 

p<-plot_heatmap(GP1_prune, sample.label="lifestage", low="#66CCFF", high="#000033", na.value="white", sample.order=sample_order, taxa.label="DIV")
p$scales$scales[[1]]$name <- "Lifestage"

print(p)

#ggplot2::ggsave("Mcap2020/Figures/ITS2/div_heatmap.jpg", p, width = 6, height = 6)
```

Run the NMDS visualization again with the dataset with rare samples removed. Run a PERMANOVA.  

```{r}
GP.ordb.filt <- ordinate(GP1_prune, "NMDS", "bray")

p3b = plot_ordination(GP1_prune, GP.ordb.filt, type="samples", color="lifestage"); p3b 

p3b_viz_filt<-p3b  + 
  geom_point(aes(colour=lifestage, shape=lifestage), size=4) + 
  ggtitle("NMDS")+
  scale_colour_manual(name="Lifestage", values=c("#8C510A", "#DFC27D","#DFC27D", "#DFC27D", "#80CDC1", "#80CDC1", "#80CDC1","#003C30", "#BA55D3", "#BA55D3")) + 
  scale_shape_manual(name="Lifestage", values = c("Egg (1 hpf)"=1, "Attached Recruit (231 hpf)"=13, "Embryo (5 hpf)"=2, "Embryo (38 hpf)"=3, "Embryo (65 hpf)"=4, "Larvae (93 hpf)"=5, "Larvae (163 hpf)"=6, "Larvae (231 hpf)"=8, "Metamorphosed Polyp (231 hpf)"=11, "Attached Recruit (183 hpf)"=12)) +
  theme_classic() + 
  theme(
    panel.background=element_blank()
  ); p3b_viz_filt

ggsave(plot=p3b_viz_filt, "Mcap2020/Figures/ITS2/nmds_its2_div_filtered.jpeg", h=5, w=8)
ggsave(plot=p3b_viz_filt, "Mcap2020/Figures/ITS2/nmds_its2_div_filtered.pdf", h=5, w=8)
```

Run PERMANOVA on filtered data set. 
```{r}
metadata_filt <- as(sample_data(GP1_prune), "data.frame")

perm2_filt<-adonis2(phyloseq::distance(GP1_prune, method="bray") ~ lifestage,
       data = metadata_filt)

perm2_filt
```

There is significant variation of dominant taxa by life stage.  

adonis2(formula = phyloseq::distance(GP1_prune, method = "bray") ~ lifestage, data = metadata_filt)
          Df SumOfSqs      R2      F Pr(>F)    
lifestage  9 0.042778 0.50967 3.2338  0.001 ***
Residual  28 0.041155 0.49033                  
Total     37 0.083934 1.00000  

Summarize by lifestage calculating means across replicates of each life stage.  
```{r}
mergedGP = merge_samples(GP1_prune, "lifestage")

lifestage_order<-c("Egg (1 hpf)", "Embryo (5 hpf)", "Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf)", "Larvae (163 hpf)", "Larvae (231 hpf)", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf)", "Attached Recruit (231 hpf)")

plot_heatmap(mergedGP)

```

Draw by lifestage with Complex Heatmap.  
```{r}
library(dendsort)
library(WGCNA)
library(circlize)

table<-as.matrix(t(otu_table(mergedGP)))
                 
row_dend = dendsort(hclust(dist(table)))
col_dend = dendsort(hclust(dist(t(table))))

#col_fun = colorRamp2(c(0, 50, 100), c("white", "darkgray", "black"))
col_fun = colorRamp2(c(0, .25, .50, .75, 1), hcl_palette = "Blue-Red 3")

pdf(file = "Mcap2020/Figures/ITS2/DIV-ComplexHeatmap.pdf", height = 7, width = 7)
ComplexHeatmap::Heatmap(table, name = "Relative \nAbundance", row_title = "DIV", column_title = "ITS2 DIV Relative Abundance (>1%)", 
        col = col_fun, 
        row_names_side = "left", 
        row_dend_side = "left",
        width = unit(4, "in"), 
        height = unit(4, "in"), 
        column_dend_height = unit(.5, "in"),
        row_dend_width = unit(.5, "in"),
        column_dend_reorder = FALSE,
        row_dend_reorder = TRUE,
        row_gap = unit(2.5, "mm"),
        clustering_distance_rows = "euclidean",
        clustering_method_rows = "complete",
        border = TRUE, 
        column_names_gp =  gpar(fontsize = 12, border=FALSE),
        column_names_rot = 35,
        cluster_rows = row_dend, 
        #cluster_columns = col_dend,
        column_order = lifestage_order, 
        row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE))
dev.off()

```

Plot heatmap for all taxa, including those <1%.  
```{r}
mergedGP_all = merge_samples(GP1_rare, "lifestage")

lifestage_order<-c("Egg (1 hpf)", "Embryo (5 hpf)", "Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf)", "Larvae (163 hpf)", "Larvae (231 hpf)", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf)", "Attached Recruit (231 hpf)")

plot_heatmap(mergedGP_all)

```

Draw by lifestage with Complex Heatmap for all taxa.  
```{r}
table<-as.matrix(t(otu_table(mergedGP_all)))
                 
row_dend = dendsort(hclust(dist(table)))
col_dend = dendsort(hclust(dist(t(table))))

#col_fun = colorRamp2(c(0, 50, 100), c("white", "darkgray", "black"))
col_fun = colorRamp2(c(0, .25, .50, .75, 1), hcl_palette = "Blue-Red 3")

pdf(file = "Mcap2020/Figures/ITS2/DIV-ComplexHeatmap_all_taxa.pdf", height = 14, width = 7)
ComplexHeatmap::Heatmap(table, name = "Relative \nAbundance", row_title = "DIV", column_title = "ITS2 DIV Relative Abundance", 
        col = col_fun, 
        row_names_side = "left", 
        row_dend_side = "left",
        width = unit(4, "in"), 
        height = unit(12, "in"), 
        column_dend_height = unit(.5, "in"),
        row_dend_width = unit(.5, "in"),
        column_dend_reorder = FALSE,
        row_dend_reorder = TRUE,
        row_gap = unit(2.5, "mm"),
        clustering_distance_rows = "euclidean",
        clustering_method_rows = "complete",
        border = TRUE, 
        column_names_gp =  gpar(fontsize = 12, border=FALSE),
        column_names_rot = 35,
        cluster_rows = row_dend, 
        #cluster_columns = col_dend,
        column_order = lifestage_order, 
        row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE))
dev.off()

```

# **Alpha Diversity**  

Calculate diversity of those that have more than 1 sequence (no singletons).
```{r}
GP1_rare_untransformed<-rarefy_even_depth(GP1, sample.size = min(sample_sums(GP1)),
  rngseed = TRUE, replace = TRUE, trimOTUs = TRUE, verbose = TRUE)

GP2 <- prune_taxa(taxa_sums(GP1_rare_untransformed) > 1, GP1_rare_untransformed)
```

Plot all richness metrics.  
```{r}
plot_richness(GP2)
```

Plot the two most commonly used - Shannon and Inverse Simpson.  
```{r}
alpha1<-plot_richness(GP2, measures=c("InvSimpson", "Shannon"), x="lifestage", color="lifestage", nrow=2)

#reorder factors 
alpha1$data$lifestage <- as.character(alpha1$data$lifestage)
alpha1$data$lifestage <- factor(alpha1$data$lifestage, levels=lifestage_order)

alpha1_plot <- alpha1 + geom_boxplot() + geom_point(size=3) + theme_classic() + scale_color_manual() + scale_color_manual(name="Lifestage", 
                      values = c("Egg (1 hpf)"="#8C510A", "Embryo (5 hpf)"="#DFC27D", "Attached Recruit (231 hpf)"="#BA55D3", "Embryo (38 hpf)"="#DFC27D", "Embryo (65 hpf)"="#DFC27D", "Larvae (93 hpf)"="#80CDC1", "Larvae (163 hpf)"="#80CDC1", "Larvae (231 hpf)"="#80CDC1", "Metamorphosed Polyp (231 hpf)"="#003C30", "Attached Recruit (183 hpf)"="#BA55D3"),
                      breaks = c("Egg (1 hpf)", "Embryo (5 hpf)", "Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)"), 
                     labels = c("Egg (1 hpf)", "Embryo (5 hpf)", "Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)")) + 
  theme(axis.text.x=element_text(angle=45, hjust=1, color="black")) + 
  theme(axis.title.x=element_blank())+
  theme(legend.position="none");alpha1_plot

ggsave(plot=alpha1_plot, file="Mcap2020/Figures/ITS2/alpha_rare.png", width=4, height=7)
ggsave(plot=alpha1_plot, file="Mcap2020/Figures/ITS2/alpha_rare.pdf", width=4, height=7)
```

Test for differences in alpha diversity.  
```{r}
metadata <- as(sample_data(GP2), "data.frame")
metadata$sample<-rownames(metadata)

alpha_rare<-estimate_richness(GP2, measures=c("InvSimpson", "Shannon", "Chao1"))
alpha_rare$sample<-rownames(alpha_rare)

alpha_rare<-full_join(metadata, alpha_rare, by="sample")

summary(aov(Shannon~lifestage, data=alpha_rare))
summary(aov(InvSimpson~lifestage, data=alpha_rare))
summary(aov(Chao1~lifestage, data=alpha_rare))
```

There are no changes in alpha diversity of ITS2 in development measured by Shannon, Inverse Simpson, or Chao1.  

# **Plot profile level information** 

View the profile level phyloseq object.   
```{r}
coral
```

Note that taxa 8 is Clade C and taxa 9 is Clade D. 

Create tree.  
```{r}
random_tree_profile = rtree(ntaxa(coral), rooted=TRUE, tip.label=taxa_names(coral))
```

Merge phyloseq data. 
```{r}
phylo_coral = merge_phyloseq(coral, Metadata, random_tree_profile)
phylo_coral
```

Filter out samples with low numbers in groups from phyloseq object.  
```{r}
phylo_coral<-ps_filter(phylo_coral, lifestage != "Metamorphosed Polyp (183 hpf)")
phylo_coral<-ps_arrange(phylo_coral, lifestage)
```

Reorder lifestage.  
```{r}
levels(as.factor(sample_data(phylo_coral)$lifestage))

sample_data(phylo_coral)$lifestage = factor(sample_data(phylo_coral)$lifestage, levels = c("Egg (1 hpf)","Embryo (5 hpf)","Embryo (38 hpf)","Embryo (65 hpf)", "Larvae (93 hpf)", "Larvae (163 hpf)", "Larvae (231 hpf)", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf)", "Attached Recruit (231 hpf)"))

levels(as.factor(sample_data(phylo_coral)$lifestage))
```

Now build a tree with the combined data.  
```{r}
plot_tree(phylo_coral, color="lifestage", label.tips="taxa_names", ladderize="left", plot.margin=0.3)
```

Plot a heatmap.  

```{r}
plot_heatmap(phylo_coral, "NMDS", "bray", "lifestage", low="#66CCFF", high="#000033", na.value="white")
```

Plot ordinations.  

```{r}
GP1_profile = phylo_coral
GP1_profile
```

Subsample to minimum sampling depth of 8,657.  
```{r}
min(sample_sums(GP1_profile))

GP1_rare_profile<-rarefy_even_depth(GP1_profile, sample.size = min(sample_sums(GP1_profile)),
  rngseed = TRUE, replace = TRUE, trimOTUs = TRUE, verbose = TRUE)

sample_sums(GP1_rare_profile)
```
All samples now have 8657 sequences. 

Transform to relative abundance as a proportion. 
```{r}
GP1_rare_profile = transform_sample_counts(GP1_rare_profile, function(x) x/sum(x))
```

View tree and heatmap after subsampling  
```{r}
plot_tree(GP1_rare_profile, color="lifestage", label.tips="taxa_names", ladderize="left", plot.margin=0.3)
plot_heatmap(GP1_rare_profile, "NMDS", "bray", "lifestage", low="#66CCFF", high="#000033", na.value="white")
```

Plot abundance of each profile across lifestages.  
```{r}
t<-plot_bar(GP1_rare_profile, fill="its2_type_profile")
t$data$Sample <- factor(t$data$Sample, levels = sample_order)
t_plot<-t + scale_fill_manual(name="ITS2 Type Profile", values=c("darkorange", "brown4")) + geom_hline(yintercept=0.5, linetype="dashed", color = "black", size=1) + facet_grid(~lifestage, scales = "free", space = "free", labeller = label_wrap_gen(width=10)) + ylab("Relative Abundance (%)") + theme(strip.text.x = element_text(size = 5, face="bold"))

t_plot

rel_abund_data<-t$data

ggsave("Mcap2020/Figures/ITS2/profile_plot.png", t_plot, width=11, height=6)
ggsave("Mcap2020/Figures/ITS2/profile_plot.pdf", t_plot, width=11, height=6)
```

Calculate the average percent abundance for each taxa in each lifestage.  
```{r}
means <- rel_abund_data %>%
  group_by(its2_type_profile, lifestage) %>%
  # Add abundances within each profile and lifestage
  summarize_at("Abundance", mean)%>%
  spread(its2_type_profile, Abundance);means
```

Calculate the average percent abundance for each taxa across all stages.  
```{r}
means_all <- rel_abund_data %>%
  group_by(its2_type_profile) %>%
  # Add abundances within each profile and lifestage
  summarise(mean=mean(Abundance), sd=sd(Abundance), n=length(Abundance), se=sd/sqrt(n));means_all
```

Use ANOVA to test for differences in relative abundance of C and D across lifestages.  
```{r}
model<-rel_abund_data %>% 
  select(its2_type_profile, Abundance, lifestage, Sample)%>%
  spread(its2_type_profile, Abundance)%>%
  aov(`C31-C17d-C21-C21ac-C31a`~lifestage, data=.);summary(model)
```

There is a significant change in relative abundance of major profile (C) over development.  

            Df  Sum Sq   Mean Sq F value  Pr(>F)   
lifestage    9 0.01336 0.0014841   3.563 0.00464 **
Residuals   28 0.01166 0.0004166   

Conduct posthoc testing by lifestage. 
```{r}
library(multcomp)
emm<-emmeans(model, ~lifestage)
pairs(emm)
cld(emm, Letters=letters)
```

The only differences are between eggs and 231 hpf larvae (p=0.0173) and eggs and 93 hpf larvae (p=0.0420).  

## Plot relative abundance mean values  

```{r}
rel_abund_plot<-rel_abund_data%>%
  dplyr::select(its2_type_profile, Abundance, lifestage, Sample)%>%
  group_by(its2_type_profile, lifestage)%>%
  summarise(mean=mean(Abundance, na.rm=TRUE), se=sd(Abundance, na.rm=TRUE)/sqrt(length(Abundance)))%>%
  
  ggplot(aes(x=lifestage, y=mean, color=its2_type_profile))+
  geom_point()+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0.5)+
  ylab("Mean Relative Abundance")+
  ylim(0,1)+
  xlab("Life Stage")+
  scale_color_manual(name="ITS2 Type Profile", values=c("darkorange", "brown4"))+
  theme_classic()+
  theme(
    legend.position="right",
    text=element_text(color="black", size=12), 
    axis.text=element_text(color="black", size=11), 
    axis.text.x=element_text(angle=45, hjust=1, vjust=1), 
    axis.title=element_text(face="bold")
  );rel_abund_plot

ggsave("Mcap2020/Figures/ITS2/profile_plot_means.png", plot=rel_abund_plot, width=6, height=6)
ggsave("Mcap2020/Figures/ITS2/profile_plot_means.pdf", plot=rel_abund_plot, width=6, height=6)
```

## Plot ratio of C:D 

Posthoc results added to plot are analyzed below.  

```{r}
ratio_plot<-rel_abund_data%>%
  dplyr::select(its2_type_profile, Abundance, lifestage, Sample)%>%
  pivot_wider(values_from = Abundance, names_from = its2_type_profile)%>%
  mutate(ratio=`C31-C17d-C21-C21ac-C31a`/`D1-D4-D6-D1ab-D17d`)%>%
  group_by(lifestage)%>%
  summarise(mean=mean(ratio, na.rm=TRUE), se=sd(ratio, na.rm=TRUE)/sqrt(length(ratio)))%>%
  
  ggplot(aes(x=lifestage, y=mean))+
  geom_point()+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0.2)+
  ylab("C:D Ratio")+
  ylim(0,2)+
  xlab("Lifestage")+
  geom_text(x=1, y=1.25, label="a", color="red")+
  geom_text(x=2, y=1.35, label="ab")+
  geom_text(x=3, y=1.3, label="ab")+
  geom_text(x=4, y=1.5, label="ab")+
  geom_text(x=5, y=1.55, label="ab")+
  geom_text(x=6, y=1.58, label="ab")+
  geom_text(x=7, y=1.68, label="b", color="red")+
  geom_text(x=8, y=1.5, label="ab")+
  geom_text(x=9, y=1.35, label="ab")+
  geom_text(x=10, y=1.45, label="ab")+
  theme_classic()+
  theme(
    text=element_text(color="black", size=12), 
    axis.text=element_text(color="black", size=11), 
    axis.text.x=element_text(angle=45, hjust=1, vjust=1), 
    axis.title=element_text(face="bold")
  );ratio_plot

ggsave("Mcap2020/Figures/ITS2/CD_ratio_means.png", plot=ratio_plot, width=5, height=5)
ggsave("Mcap2020/Figures/ITS2/CD_ratio_means.pdf", plot=ratio_plot, width=5, height=5)
```

Run ANOVA on ratios. 
```{r}
model<-rel_abund_data%>%
  dplyr::select(its2_type_profile, Abundance, lifestage, Sample)%>%
  pivot_wider(values_from = Abundance, names_from = its2_type_profile)%>%
  mutate(ratio=`C31-C17d-C21-C21ac-C31a`/`D1-D4-D6-D1ab-D17d`)

qqPlot(model$ratio)

model<-rel_abund_data%>%
  dplyr::select(its2_type_profile, Abundance, lifestage, Sample)%>%
  pivot_wider(values_from = Abundance, names_from = its2_type_profile)%>%
  mutate(ratio=`C31-C17d-C21-C21ac-C31a`/`D1-D4-D6-D1ab-D17d`)%>%
  aov(ratio~lifestage, data=.);summary(model)
```

There is a significant change in C:D ratio across development. 

            Df Sum Sq Mean Sq F value Pr(>F)  
lifestage    9 0.3829 0.04254   3.056 0.0112 *
Residuals   28 0.3897 0.01392                 
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1  

Conduct posthoc testing by lifestage. 
```{r}
emm<-emmeans(model, ~lifestage)
pairs(emm)
cld(emm, Letters=letters)
```

Egg (1 hpf) - Larvae (231 hpf)                              -0.3104 0.0834 28  -3.721  0.0254
