---
title: "WGCNA - GoSeq Functional Annotation"
author: "Ariana S Huffmyer"
date: "5/2/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

Functional enrichment of WGCNA modules produced in `WGCNA_Mcap.Rmd` script for the *Montipora capitata* 2020 developmental time series.  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# The following setting is important, do not omit.
options(stringsAsFactors = FALSE) #Set Strings to character
```

Load required libraries.  

```{r}
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("genefilter" %in% rownames(installed.packages()) == 'FALSE') install.packages('genefilter') 
if ("DESeq2" %in% rownames(installed.packages()) == 'FALSE') install.packages('DESeq2') 
if ("RColorBrewer" %in% rownames(installed.packages()) == 'FALSE') install.packages('RColorBrewer') 
if ("WGCNA" %in% rownames(installed.packages()) == 'FALSE') install.packages('WGCNA') 
if ("flashClust" %in% rownames(installed.packages()) == 'FALSE') install.packages('flashClust') 
if ("gridExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('gridExtra') 
if ("ComplexHeatmap" %in% rownames(installed.packages()) == 'FALSE') install.packages('ComplexHeatmap') 
if ("goseq" %in% rownames(installed.packages()) == 'FALSE') install.packages('goseq') 
if ("dplyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('dplyr') 
if ("clusterProfiler" %in% rownames(installed.packages()) == 'FALSE') install.packages('clusterProfiler') 
if ("pheatmap" %in% rownames(installed.packages()) == 'FALSE') install.packages('pheatmap') 
if ("magrittr" %in% rownames(installed.packages()) == 'FALSE') install.packages('magrittr') 

library("tidyverse")
library("genefilter")
library("DESeq2")
library("RColorBrewer")
library("WGCNA")
library("flashClust")
library("gridExtra")
library("ComplexHeatmap")
library("goseq")
library("dplyr")
library("clusterProfiler")
library("pheatmap")
library("magrittr")
```

# **Read in data**  

Load in Gene Info data frame generated in WGCNA_Mcap.Rmd script and subset list of genes and module assignment for each gene.   

```{r}
geneInfo<-read.csv("Mcap2020/Output/TagSeq/geneInfo.csv")

geneInfo<-geneInfo%>%
  select(gene_id, moduleColor)%>%
  rename(Module=moduleColor)%>%
  mutate(Module=as.factor(Module))
```

Add in length and GO term information from annotation file from Erin Chille's genome annotation.    

```{r}
annot<-read_tsv("Mcap2020/Data/TagSeq/200824_Mcap_Blast_GO_KO.tsv", col_names=TRUE)

geneInfo$Length<-annot$length[match(geneInfo$gene_id, annot$gene_id)]
geneInfo$GO_IDs<-annot$GO_IDs[match(geneInfo$gene_id, annot$gene_id)]
```

Generate a file with all genes 

```{r}
all<-geneInfo%>%
  select(gene_id, Length, GO_IDs)
```

Generate a file for genes in modules list  

```{r}
module_list<-geneInfo%>%
  select(gene_id, Length, GO_IDs, Module)

levels(module_list$Module)

module1<-module_list%>%filter(Module=="1")
module2<-module_list%>%filter(Module=="2")
module4<-module_list%>%filter(Module=="4")
module6<-module_list%>%filter(Module=="6")
module8<-module_list%>%filter(Module=="8")
module11<-module_list%>%filter(Module=="11")
module12<-module_list%>%filter(Module=="12")
module14<-module_list%>%filter(Module=="14")
module15<-module_list%>%filter(Module=="15")
module16<-module_list%>%filter(Module=="16")
module17<-module_list%>%filter(Module=="17")
module18<-module_list%>%filter(Module=="18")
```

Read in GO Slim file.  

```{r}
goslim<-read.csv("Mcap2020/Data/TagSeq/GO-GOslim.csv") 

goslim <- goslim %>% select(-term)
```

# **Run GoSeq by Module**  

## Module 1  

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(module1$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- module1%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(module1$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
module1.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(module1.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_module1.csv")
```


Plot results by Biological Process level.    

```{r}
# Plotting by BP function
module1.GO_plot <- module1.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

module1.GO_plot$GOSlim_bin <- as.factor(module1.GO_plot$GOSlim_bin)

#plot significantly enriched GO terms by Slim Category
GO_module1.plot <-  ggplot(module1.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_Module1.png", plot=GO_module1.plot, dpi=300, width=12, height=12, units="in")
```

## Module 2 

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(module2$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- module2%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(module2$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
module2.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(module2.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_module2.csv")
```

Plot results by Biological Process level.    

```{r}
# Plotting by BP function
module2.GO_plot <- module2.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

module2.GO_plot$GOSlim_bin <- as.factor(module2.GO_plot$GOSlim_bin)

#plot significantly enriched GO terms by Slim Category
GO_module2.plot <-  ggplot(module2.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_Module2.png", plot=GO_module2.plot, dpi=300, width=12, height=12, units="in")
```

## Module 4 

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(module4$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- module4%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(module4$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
module4.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(module4.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_module4.csv")
```

Plot results by Biological Process level.    

```{r}
# Plotting by BP function
module4.GO_plot <- module4.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

module4.GO_plot$GOSlim_bin <- as.factor(module4.GO_plot$GOSlim_bin)

#plot significantly enriched GO terms by Slim Category
GO_module4.plot <-  ggplot(module4.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_Module4.png", plot=GO_module4.plot, dpi=300, width=12, height=12, units="in")
```















NEXT STEP: 
Generate a file for each lifestage 






Bring in goseq files 

Add length for each gene 

Edit function to do by - function and - lifestage  


## Module 6 

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(module6$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- module6%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(module6$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
module6.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(module6.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_module6.csv")
```

Plot results by Biological Process level.    

```{r}
# Plotting by BP function
module6.GO_plot <- module6.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

module6.GO_plot$GOSlim_bin <- as.factor(module6.GO_plot$GOSlim_bin)

#plot significantly enriched GO terms by Slim Category
GO_module6.plot <-  ggplot(module6.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_Module6.png", plot=GO_module6.plot, dpi=300, width=12, height=12, units="in")
```
## Module 8 

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(module8$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- module8%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(module8$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
module8.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(module8.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_module8.csv")
```

Plot results by Biological Process level.    

```{r}
# Plotting by BP function
module8.GO_plot <- module8.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

module8.GO_plot$GOSlim_bin <- as.factor(module8.GO_plot$GOSlim_bin)

#plot significantly enriched GO terms by Slim Category
GO_module8.plot <-  ggplot(module8.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_Module8.png", plot=GO_module8.plot, dpi=300, width=12, height=12, units="in")
```
## Module 11 

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(module11$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- module11%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(module11$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
module11.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(module11.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_module11.csv")
```

Plot results by Biological Process level.    

```{r}
# Plotting by BP function
module11.GO_plot <- module11.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

module11.GO_plot$GOSlim_bin <- as.factor(module11.GO_plot$GOSlim_bin)

#plot significantly enriched GO terms by Slim Category
GO_module11.plot <-  ggplot(module11.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_Module11.png", plot=GO_module11.plot, dpi=300, width=12, height=12, units="in")
```

## Module 12 

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(module12$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- module12%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(module12$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
module12.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(module12.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_module12.csv")
```

Plot results by Biological Process level.    

```{r}
# Plotting by BP function
module12.GO_plot <- module12.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

module12.GO_plot$GOSlim_bin <- as.factor(module12.GO_plot$GOSlim_bin)

#plot significantly enriched GO terms by Slim Category
GO_module12.plot <-  ggplot(module12.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_Module12.png", plot=GO_module12.plot, dpi=300, width=12, height=12, units="in")
```
## Module 14 

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(module14$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- module14%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(module14$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
module14.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(module14.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_module14.csv")
```

Plot results by Biological Process level.    

```{r}
# Plotting by BP function
module14.GO_plot <- module14.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

module14.GO_plot$GOSlim_bin <- as.factor(module14.GO_plot$GOSlim_bin)

#plot significantly enriched GO terms by Slim Category
GO_module14.plot <-  ggplot(module14.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_Module14.png", plot=GO_module14.plot, dpi=300, width=12, height=12, units="in")
```
## Module 15 

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(module15$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- module15%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(module15$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
module15.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(module15.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_module15.csv")
```

Plot results by Biological Process level.    

```{r}
# Plotting by BP function
module15.GO_plot <- module15.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

module15.GO_plot$GOSlim_bin <- as.factor(module15.GO_plot$GOSlim_bin)

#plot significantly enriched GO terms by Slim Category
GO_module15.plot <-  ggplot(module15.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_Module15.png", plot=GO_module15.plot, dpi=300, width=12, height=12, units="in")
```
## Module 16 

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(module16$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- module16%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(module16$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
module16.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(module16.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_module16.csv")
```

Plot results by Biological Process level.    

```{r}
# Plotting by BP function
module16.GO_plot <- module16.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

module16.GO_plot$GOSlim_bin <- as.factor(module16.GO_plot$GOSlim_bin)

#plot significantly enriched GO terms by Slim Category
GO_module16.plot <-  ggplot(module16.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_Module16.png", plot=GO_module16.plot, dpi=300, width=12, height=12, units="in")
```
## Module 17 

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(module17$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- module17%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(module17$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
module17.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(module17.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_module17.csv")
```

Plot results by Biological Process level.    

```{r}
# Plotting by BP function
module17.GO_plot <- module17.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

module17.GO_plot$GOSlim_bin <- as.factor(module17.GO_plot$GOSlim_bin)

#plot significantly enriched GO terms by Slim Category
GO_module17.plot <-  ggplot(module17.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_Module17.png", plot=GO_module17.plot, dpi=300, width=12, height=12, units="in")
```
## Module 18 

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(module18$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- module18%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(module18$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
module18.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(module18.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_module18.csv")
```

Plot results by Biological Process level.    

```{r}
# Plotting by BP function
module18.GO_plot <- module18.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

module18.GO_plot$GOSlim_bin <- as.factor(module18.GO_plot$GOSlim_bin)

#plot significantly enriched GO terms by Slim Category
GO_module18.plot <-  ggplot(module18.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_Module18.png", plot=GO_module18.plot, dpi=300, width=12, height=12, units="in")
```

# **Run GoSeq by Lifestage**  

Assemble a list of genes in the modules that are correlated with each respective lifestage and run GOSeq with those lifestage groupings.  

```{r}
egg_list<-c("1")
embryo_list<-c("1")
larvae1_list<-c("6", "11", "17", "14")
larvae2_list<-c("16", "18")
larvae3_list<-c("16", "4", "2")
larvae4_list<-c("4", "2")
larvae6_list<-c("12")
polyp_list<-c("12")
recruit1_list<-c("15", "8", "14")
recruit2_list<-c("15", "8", "14")

eggs<-module_list%>%filter(Module %in% egg_list)
embryos<-module_list%>%filter(Module %in% embryo_list)
larvae1<-module_list%>%filter(Module %in% larvae1_list)
larvae2<-module_list%>%filter(Module %in% larvae2_list)
larvae3<-module_list%>%filter(Module %in% larvae3_list)
larvae4<-module_list%>%filter(Module %in% larvae4_list)
larvae6<-module_list%>%filter(Module %in% larvae6_list)
polyps<-module_list%>%filter(Module %in% polyp_list)
recruit1<-module_list%>%filter(Module %in% recruit1_list)
recruit2<-module_list%>%filter(Module %in% recruit2_list)
```


## Eggs  

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(eggs$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- eggs%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(eggs$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
eggs.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(eggs.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_eggs.csv")
```


Plot results by Biological Process level.    

```{r}
# Plotting by BP function
eggs.GO_plot <- eggs.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

eggs.GO_plot$GOSlim_bin <- as.factor(eggs.GO_plot$GOSlim_bin)
eggs.GO_plot$Lifestage <- c("Fertilized Egg (1 hpf)")

#plot significantly enriched GO terms by Slim Category
GO_eggs.plot <-  ggplot(eggs.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_eggs_BP.png", plot=GO_eggs.plot, dpi=300, width=12, height=12, units="in")
```

Plot results by Molecular Function level.    

```{r}
# Plotting by BP function
eggs.GO_plot.MF <- eggs.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "MF") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

eggs.GO_plot.MF$GOSlim_bin <- as.factor(eggs.GO_plot.MF$GOSlim_bin)
eggs.GO_plot.MF$Lifestage <- c("Fertilized Egg (1 hpf)")

#plot significantly enriched GO terms by Slim Category
GO_eggs.plot.MF <-  ggplot(eggs.GO_plot.MF, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_eggs_MF.png", plot=GO_eggs.plot.MF, dpi=300, width=12, height=12, units="in")
```

## Embryos  

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(embryos$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- embryos%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(embryos$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
embryos.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(embryos.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_embryos.csv")
```


Plot results by Biological Process level.    

```{r}
# Plotting by BP function
embryos.GO_plot <- embryos.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

embryos.GO_plot$GOSlim_bin <- as.factor(embryos.GO_plot$GOSlim_bin)
embryos.GO_plot$Lifestage <- c("Embryo (5 hpf)")

#plot significantly enriched GO terms by Slim Category
GO_embryos.plot <-  ggplot(embryos.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_embryos_BP.png", plot=GO_embryos.plot, dpi=300, width=12, height=12, units="in")
```

Plot results by Molecular level.    

```{r}
# Plotting by BP function
embryos.GO_plot.MF <- embryos.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "MF") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

embryos.GO_plot.MF$GOSlim_bin <- as.factor(embryos.GO_plot.MF$GOSlim_bin)
embryos.GO_plot.MF$Lifestage <- c("Embryo (5 hpf)")

#plot significantly enriched GO terms by Slim Category
GO_embryos.plot.MF <-  ggplot(embryos.GO_plot.MF, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_embryos_MF.png", plot=GO_embryos.plot.MF, dpi=300, width=12, height=24, units="in")
```


## Larvae 1  

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(larvae1$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- larvae1%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(larvae1$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
larvae1.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(larvae1.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_larvae1.csv")
```

Plot results by Biological Process level.    

```{r}
# Plotting by BP function
larvae1.GO_plot <- larvae1.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

larvae1.GO_plot$GOSlim_bin <- as.factor(larvae1.GO_plot$GOSlim_bin)
larvae1.GO_plot$Lifestage <- c("Embryo (38 hpf)")

#plot significantly enriched GO terms by Slim Category
GO_larvae1.plot <-  ggplot(larvae1.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_larvae1_BP.png", plot=GO_larvae1.plot, dpi=300, width=12, height=12, units="in")
```

Plot by Molecular Function level.  

```{r}
# Plotting by BP function
larvae1.GO_plot.MF <- larvae1.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "MF") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

larvae1.GO_plot.MF$GOSlim_bin <- as.factor(larvae1.GO_plot.MF$GOSlim_bin)
larvae1.GO_plot.MF$Lifestage <- c("Embryo (38 hpf)")

#plot significantly enriched GO terms by Slim Category
GO_larvae1.plot.MF <-  ggplot(larvae1.GO_plot.MF, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_larvae1_MF.png", plot=GO_larvae1.plot.MF, dpi=300, width=12, height=12, units="in")
```


## Larvae 2  

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(larvae2$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- larvae2%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(larvae2$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
larvae2.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(larvae2.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_larvae2.csv")
```

Plot results by Biological Process level.    

```{r}
# Plotting by BP function
larvae2.GO_plot <- larvae2.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

larvae2.GO_plot$GOSlim_bin <- as.factor(larvae2.GO_plot$GOSlim_bin)
larvae2.GO_plot$Lifestage <- c("Embryo (65 hpf)")

#plot significantly enriched GO terms by Slim Category
GO_larvae2.plot <-  ggplot(larvae2.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_larvae2_BP.png", plot=GO_larvae2.plot, dpi=300, width=12, height=12, units="in")
```

Plot results by Molecular Function level.    

```{r}
# Plotting by BP function
larvae2.GO_plot.MF <- larvae2.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "MF") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

larvae2.GO_plot.MF$GOSlim_bin <- as.factor(larvae2.GO_plot.MF$GOSlim_bin)
larvae2.GO_plot.MF$Lifestage <- c("Embryo (65 hpf)")

#plot significantly enriched GO terms by Slim Category
GO_larvae2.plot.MF <-  ggplot(larvae2.GO_plot.MF, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_larvae2_MF.png", plot=GO_larvae2.plot.MF, dpi=300, width=12, height=12, units="in")
```


## Larvae 3  

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(larvae3$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- larvae3%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(larvae3$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
larvae3.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(larvae3.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_larvae3.csv")
```

Plot results by Biological Process level.    

```{r}
# Plotting by BP function
larvae3.GO_plot <- larvae3.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

larvae3.GO_plot$GOSlim_bin <- as.factor(larvae3.GO_plot$GOSlim_bin)
larvae3.GO_plot$Lifestage <- c("Larvae (93 hpf)")

#plot significantly enriched GO terms by Slim Category
GO_larvae3.plot <-  ggplot(larvae3.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_larvae3_BP.png", plot=GO_larvae3.plot, dpi=300, width=12, height=12, units="in")
```

Plot results by Molecular Function level.    

```{r}
# Plotting by BP function
larvae3.GO_plot.MF <- larvae3.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "MF") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

larvae3.GO_plot.MF$GOSlim_bin <- as.factor(larvae3.GO_plot.MF$GOSlim_bin)
larvae3.GO_plot.MF$Lifestage <- c("Larvae (93 hpf)")

#plot significantly enriched GO terms by Slim Category
GO_larvae3.plot.MF <-  ggplot(larvae3.GO_plot.MF, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_larvae3_MF.png", plot=GO_larvae3.plot.MF, dpi=300, width=12, height=24, units="in")
```


## Larvae 4  

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(larvae4$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- larvae4%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(larvae4$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
larvae4.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(larvae4.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_larvae4.csv")
```

Plot results by Biological Process level.    

```{r}
# Plotting by BP function
larvae4.GO_plot <- larvae4.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

larvae4.GO_plot$GOSlim_bin <- as.factor(larvae4.GO_plot$GOSlim_bin)
larvae4.GO_plot$Lifestage <- c("Larvae (163 hpf)")

#plot significantly enriched GO terms by Slim Category
GO_larvae4.plot <-  ggplot(larvae4.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_larvae4_BP.png", plot=GO_larvae4.plot, dpi=300, width=12, height=12, units="in")
```

Plot results by Molecular Function level.    

```{r}
# Plotting by BP function
larvae4.GO_plot.MF <- larvae4.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "MF") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

larvae4.GO_plot.MF$GOSlim_bin <- as.factor(larvae4.GO_plot.MF$GOSlim_bin)
larvae4.GO_plot.MF$Lifestage <- c("Larvae (163 hpf)")

#plot significantly enriched GO terms by Slim Category
GO_larvae4.plot.MF <-  ggplot(larvae4.GO_plot.MF, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_larvae4_MF.png", plot=GO_larvae4.plot.MF, dpi=300, width=12, height=24, units="in")
```


## Larvae 6  

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(larvae6$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- larvae6%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(larvae6$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
larvae6.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(larvae6.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_larvae6.csv")
```

Plot results by Biological Process level.    

```{r}
# Plotting by BP function
larvae6.GO_plot <- larvae6.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

larvae6.GO_plot$GOSlim_bin <- as.factor(larvae6.GO_plot$GOSlim_bin)
larvae6.GO_plot$Lifestage <- c("Larvae (231 hpf)")

#plot significantly enriched GO terms by Slim Category
GO_larvae6.plot <-  ggplot(larvae6.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_larvae6_BP.png", plot=GO_larvae6.plot, dpi=300, width=12, height=12, units="in")
```

Plot results by Molecular Function level.    

```{r}
# Plotting by BP function
larvae6.GO_plot.MF <- larvae6.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "MF") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

larvae6.GO_plot.MF$GOSlim_bin <- as.factor(larvae6.GO_plot.MF$GOSlim_bin)
larvae6.GO_plot.MF$Lifestage <- c("Larvae (231 hpf)")

#plot significantly enriched GO terms by Slim Category
GO_larvae6.plot.MF <-  ggplot(larvae6.GO_plot.MF, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_larvae6_MF.png", plot=GO_larvae6.plot.MF, dpi=300, width=12, height=24, units="in")
```


## Polyps  

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(polyps$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- polyps%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(polyps$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
polyps.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(polyps.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_polyps.csv")
```

Plot results by Biological Process level.    

```{r}
# Plotting by BP function
polyps.GO_plot <- polyps.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

polyps.GO_plot$GOSlim_bin <- as.factor(polyps.GO_plot$GOSlim_bin)
polyps.GO_plot$Lifestage <- c("Metamorphosed Polyp (231 hpf)")

#plot significantly enriched GO terms by Slim Category
GO_polyps.plot <-  ggplot(polyps.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_polyps_BP.png", plot=GO_polyps.plot, dpi=300, width=12, height=12, units="in")
```

Plot results by Molecular Function level.    

```{r}
# Plotting by BP function
polyps.GO_plot.MF <- polyps.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "MF") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

polyps.GO_plot.MF$GOSlim_bin <- as.factor(polyps.GO_plot.MF$GOSlim_bin)
polyps.GO_plot.MF$Lifestage <- c("Metamorphosed Polyp (231 hpf)")

#plot significantly enriched GO terms by Slim Category
GO_polyps.plot.MF <-  ggplot(polyps.GO_plot.MF, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_polyps_MF.png", plot=GO_polyps.plot.MF, dpi=300, width=12, height=24, units="in")
```


## Recruit 1  

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(recruit1$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- recruit1%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(recruit1$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
recruit1.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(recruit1.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_recruit1.csv")
```

Plot results by Biological Process level.    

```{r}
# Plotting by BP function
recruit1.GO_plot <- recruit1.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

recruit1.GO_plot$GOSlim_bin <- as.factor(recruit1.GO_plot$GOSlim_bin)
recruit1.GO_plot$Lifestage <- c("Attached Recruit (183 hpf)")

#plot significantly enriched GO terms by Slim Category
GO_recruit1.plot <-  ggplot(recruit1.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_recruit1_BP.png", plot=GO_recruit1.plot, dpi=300, width=12, height=12, units="in")
```

Plot results by Molecular Function level.    

```{r}
# Plotting by BP function
recruit1.GO_plot.MF <- recruit1.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "MF") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

recruit1.GO_plot.MF$GOSlim_bin <- as.factor(recruit1.GO_plot.MF$GOSlim_bin)
recruit1.GO_plot.MF$Lifestage <- c("Attached Recruit (183 hpf)")

#plot significantly enriched GO terms by Slim Category
GO_recruit1.plot.MF <-  ggplot(recruit1.GO_plot.MF, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_recruit1_MF.png", plot=GO_recruit1.plot.MF, dpi=300, width=12, height=12, units="in")
```

## Recruit 2  

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(recruit2$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- recruit2%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(recruit2$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
recruit2.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(recruit2.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_recruit2.csv")
```

Plot results by Biological Process level.    

```{r}
# Plotting by BP function
recruit2.GO_plot <- recruit2.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

recruit2.GO_plot$GOSlim_bin <- as.factor(recruit2.GO_plot$GOSlim_bin)
recruit2.GO_plot$Lifestage <- c("Attached Recruit (231 hpf)")

#plot significantly enriched GO terms by Slim Category
GO_recruit2.plot <-  ggplot(recruit2.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_recruit2_BP.png", plot=GO_recruit2.plot, dpi=300, width=12, height=12, units="in")
```

Plot results by Molecular Function level.    

```{r}
# Plotting by BP function
recruit2.GO_plot.MF <- recruit2.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "MF") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

recruit2.GO_plot.MF$GOSlim_bin <- as.factor(recruit2.GO_plot.MF$GOSlim_bin)
recruit2.GO_plot.MF$Lifestage <- c("Attached Recruit (231 hpf)")

#plot significantly enriched GO terms by Slim Category
GO_recruit2.plot.MF <-  ggplot(recruit2.GO_plot.MF, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_recruit2_MF.png", plot=GO_recruit2.plot.MF, dpi=300, width=12, height=24, units="in")
```

# **Plot all lifestages together on one plot**  

Combine data frames together in order to generate one GoSeq plot with all terms across lifestages.  

Plot at Biological Process level.  

```{r}
go_all<-rbind(eggs.GO_plot, embryos.GO_plot, larvae1.GO_plot, larvae2.GO_plot, larvae3.GO_plot, larvae4.GO_plot, larvae6.GO_plot, polyps.GO_plot, recruit1.GO_plot, recruit2.GO_plot)

head(go_all)
dim(go_all)

levels(as.factor(go_all$Lifestage))
```

Files are merged.  

Now plot.  

```{r}
#reorder x axis lifestages 
list_lifestages<-c("Fertilized Egg (1 hpf)", "Embryo (5 hpf)", "Embryo (38 hpf)", "Larvae (93 hpf)", "Larvae (163 hpf)", "Larvae (231 hpf)", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf)", "Attached Recruit (231 hpf)")
go_all$Lifestage<-factor(as.character(go_all$Lifestage), levels=list_lifestages)

#add group to dataframe for scale colour manual 
go_all<-go_all%>%
  mutate(group=if_else(grepl("Egg", Lifestage), "Fertilized Egg", as.character(Lifestage)))%>%
  mutate(group=if_else(grepl("Embryo", Lifestage), "Embryo", as.character(group)))%>%
  mutate(group=if_else(grepl("Larvae", Lifestage), "Larvae", as.character(group)))%>%
  mutate(group=if_else(grepl("Polyp", Lifestage), "Metamorphosed Polyp", as.character(group)))%>%
  mutate(group=if_else(grepl("Recruit", Lifestage), "Attached Recruit", as.character(group)))

list_groups<-c("Fertilized Egg", "Embryo", "Larvae", "Metamorphosed Polyp", "Attached Recruit")
go_all$group<-factor(as.character(go_all$group), levels=list_groups)
  
#plot significantly enriched GO terms by Slim Category across lifestages
GO_all.BP <-  ggplot(go_all, aes(x = Lifestage, y = term)) + 
  #geom_tile(aes(fill =over_represented_pvalue)) + 
  geom_point(aes(colour=group, size=over_represented_pvalue)) +
  #facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(multi_line = TRUE))+
  facet_grid(GOSlim_bin ~ ., scales = "free", space= "free", shrink=TRUE, labeller = label_wrap_gen(width=30, multi_line = TRUE))+
  scale_colour_manual(name="Lifestage", values = c("Fertilized Egg"="#8C510A", "Embryo"="#DFC27D", "Larvae"="#80CDC1", "Metamorphosed Polyp"="#003C30", "Attached Recruit"="#BA55D3")) +
  scale_size(name="Over rep. p-value", range = c(10, 25))+
  guides(colour = guide_legend(override.aes = list(size=14)))+
  xlab("Lifestage")+
  ylab("Biological Process GO Term")+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 30, face="bold"),
  strip.background = element_rect(fill="white"), 
  axis.text.y = element_text(size = 30),
  axis.text.x = element_text(angle=35, size=30, hjust=1),
  axis.title.x = element_text(size = 40, face="bold"),
  axis.title.y = element_text(size = 40, face="bold"), 
  legend.text = element_text(size=30), 
  legend.title = element_text(size=40, face="bold"), 
  legend.key.size = unit(2, 'cm'))

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_all_BP.png", plot=GO_all.BP, dpi=80, width=35, height=35, units="in")
```
Note that there were no enriched categories at 65 hpf so it is not displayed in this plot.  


Plot at Molecular Function level.  

```{r}
go_all.MF<-rbind(eggs.GO_plot.MF, embryos.GO_plot.MF, larvae1.GO_plot.MF, larvae2.GO_plot.MF, larvae3.GO_plot.MF, larvae4.GO_plot.MF, larvae6.GO_plot.MF, polyps.GO_plot.MF, recruit1.GO_plot.MF, recruit2.GO_plot.MF)

head(go_all.MF)
dim(go_all.MF)

levels(as.factor(go_all$Lifestage))
```

Files are merged.  

Now plot.  

```{r}
#reorder x axis lifestages 
list_lifestages<-c("Fertilized Egg (1 hpf)", "Embryo (5 hpf)", "Embryo (38 hpf)", "Larvae (93 hpf)", "Larvae (163 hpf)", "Larvae (231 hpf)", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf)", "Attached Recruit (231 hpf)")
go_all.MF$Lifestage<-factor(as.character(go_all.MF$Lifestage), levels=list_lifestages)

#add group to dataframe for scale colour manual 
go_all.MF<-go_all.MF%>%
  mutate(group=if_else(grepl("Egg", Lifestage), "Fertilized Egg", as.character(Lifestage)))%>%
  mutate(group=if_else(grepl("Embryo", Lifestage), "Embryo", as.character(group)))%>%
  mutate(group=if_else(grepl("Larvae", Lifestage), "Larvae", as.character(group)))%>%
  mutate(group=if_else(grepl("Polyp", Lifestage), "Metamorphosed Polyp", as.character(group)))%>%
  mutate(group=if_else(grepl("Recruit", Lifestage), "Attached Recruit", as.character(group)))

list_groups<-c("Fertilized Egg", "Embryo", "Larvae", "Metamorphosed Polyp", "Attached Recruit")
go_all.MF$group<-factor(as.character(go_all.MF$group), levels=list_groups)
  
#plot significantly enriched GO terms by Slim Category across lifestages
GO_all.MF <-  ggplot(go_all.MF, aes(x = Lifestage, y = term)) + 
  #geom_tile(aes(fill =over_represented_pvalue)) + 
  geom_point(aes(colour=group, size=over_represented_pvalue)) +
  #facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(multi_line = TRUE))+
  facet_grid(GOSlim_bin ~ ., scales = "free", space= "free", shrink=TRUE, labeller = label_wrap_gen(width=30, multi_line = TRUE))+
  scale_colour_manual(name="Lifestage", values = c("Fertilized Egg"="#8C510A", "Embryo"="#DFC27D", "Larvae"="#80CDC1", "Metamorphosed Polyp"="#003C30", "Attached Recruit"="#BA55D3")) +
  scale_size(name="Over rep. p-value", range = c(10, 25))+
  guides(colour = guide_legend(override.aes = list(size=14)))+
  xlab("Lifestage")+
  ylab("Molecular function GO Term")+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 30, face="bold"),
  strip.background = element_rect(fill="white"), 
  axis.text.y = element_text(size = 30),
  axis.text.x = element_text(angle=35, size=30, hjust=1),
  axis.title.x = element_text(size = 40, face="bold"),
  axis.title.y = element_text(size = 40, face="bold"), 
  legend.text = element_text(size=30), 
  legend.title = element_text(size=40, face="bold"), 
  legend.key.size = unit(2, 'cm'))

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/GOenrich_all_MF.png", plot=GO_all.MF, dpi=50, width=49, height=49, units="in")
```
Note that there were no enriched categories at 65 hpf so it is not displayed in this plot.  



# **Run GoSeq by Pattern**  

Assemble a list of genes in the modules that display the following patterns from the `expression_eigengene.pdf` plot:  

- Higher early (i.e., eggs, embryos) in development (ME1, ME11, ME6)
- Higher later (i.e., larvae, recruits) in development (ME12, ME15, ME17, ME8)
- Peak in mid development (i.e., larvae) (ME16, ME2, ME4)
- Lower in mid development (i.e., larvae) (ME15, ME14)

are correlated with each respective lifestage and run GOSeq with those lifestage groupings.  

```{r}
high_early_list<-c("1", "11", "6")
high_late_list<-c("12", "15", "17", "8")
high_mid_list<-c("16", "2", "4")

high_early<-module_list%>%filter(Module %in% high_early_list)
high_late<-module_list%>%filter(Module %in% high_late_list)
high_mid<-module_list%>%filter(Module %in% high_mid_list)
```


## Genes higher in early development (embryos, eggs)  

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(high_early$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- high_early%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(high_early$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
high_early.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(high_early.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_high_early.csv")
```

Plot results by Biological Process level.    

```{r}
# Plotting by BP function
high_early.GO_plot <- high_early.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

high_early.GO_plot$GOSlim_bin <- as.factor(high_early.GO_plot$GOSlim_bin)
high_early.GO_plot$Pattern <- c("High Early in Development")

#plot significantly enriched GO terms by Slim Category
GO_high_early.plot <-  ggplot(high_early.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/Patterns/GOenrich_pattern_high_early_BP.png", plot=GO_high_early.plot, dpi=300, width=12, height=12, units="in")
```

Plot results by Molecular Function level.    

```{r}
# Plotting by BP function
high_early.GO.MF <- high_early.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "MF") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

high_early.GO.MF$GOSlim_bin <- as.factor(high_early.GO.MF$GOSlim_bin)
high_early.GO.MF$Pattern <- c("High Early in Development")

#plot significantly enriched GO terms by Slim Category
GO_high_early.plot.MF <-  ggplot(high_early.GO.MF, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/Patterns/GOenrich_pattern_high_early_MF.png", plot=GO_high_early.plot.MF, dpi=300, width=12, height=12, units="in")
```

## Genes higher in late development (late larvae, polyps, recruits)  

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(high_late$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- high_late%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(high_late$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
high_late.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(high_late.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_high_late.csv")
```

Plot results by Biological Process level.    

```{r}
# Plotting by BP function
high_late.GO_plot <- high_late.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

high_late.GO_plot$GOSlim_bin <- as.factor(high_late.GO_plot$GOSlim_bin)
high_late.GO_plot$Pattern <- c("High Late in Development")

#plot significantly enriched GO terms by Slim Category
GO_high_late.plot <-  ggplot(high_late.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/Patterns/GOenrich_pattern_high_late_BP.png", plot=GO_high_late.plot, dpi=300, width=12, height=12, units="in")
```

Plot results by Molecular Function level.    

```{r}
# Plotting by BP function
high_late.GO.MF <- high_late.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "MF") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

high_late.GO.MF$GOSlim_bin <- as.factor(high_late.GO.MF$GOSlim_bin)
high_late.GO.MF$Pattern <- c("High Late in Development")

#plot significantly enriched GO terms by Slim Category
GO_high_late.plot.MF <-  ggplot(high_late.GO.MF, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/Patterns/GOenrich_pattern_high_late_MF.png", plot=GO_high_late.plot.MF, dpi=300, width=12, height=12, units="in")
```


## Genes higher in mid development (larvae)  

Assign vectors  

```{r}
# Vector with all genes after filtering
ALL.vector <- c(all$gene_id)
ID.vector <- c(high_mid$gene_id)

#Length vector for all genes
LENGTH.vector <- c(all$Length)

GO.terms <- high_mid%>%
  select(gene_id, GO_IDs)

str(GO.terms) 

#Formatting sig gene file with a goterm per row
split <- strsplit(as.character(GO.terms$GO_IDs), "; ") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO_IDs")

GO.terms<-split2

str(GO.terms)
```

Run GoSeq  

```{r}
DEG <- as.character(high_mid$gene_id) #set the enrichment test list
DEG.vector <-c(DEG) #change to vectors
  
gene.vector=as.integer(ALL.vector%in%DEG.vector) #Construct new vector with 1 for DEG and 0 for others
names(gene.vector)=ALL.vector #set names

DEG.pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) #weight vector by length of gene

#Find enriched GO terms
GO.wall<-goseq(DEG.pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:CC", "GO:BP", "GO:MF"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GO_ID"
  
#GOslim 
GO.slim <- left_join(GO, goslim, by = "GO_ID")
GO.slim <- GO.slim[!duplicated(GO.slim$GO_ID), ]
  
#Filtering for p > 0.05
high_mid.GO <- GO.slim %>%
  dplyr::filter(over_represented_pvalue<0.05) %>%
  arrange(., ontology, term, over_represented_pvalue)

write.csv(high_mid.GO , file = "Mcap2020/Output/TagSeq/GOSeq/goseq_high_mid.csv")
```

Plot results by Biological Process level.    

```{r}
# Plotting by BP function
high_mid.GO_plot <- high_mid.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "BP") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

high_mid.GO_plot$GOSlim_bin <- as.factor(high_mid.GO_plot$GOSlim_bin)
high_mid.GO_plot$Pattern <- c("High in Mid Development")

#plot significantly enriched GO terms by Slim Category
GO_high_mid.plot <-  ggplot(high_mid.GO_plot, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/Patterns/GOenrich_pattern_high_mid_BP.png", plot=GO_high_mid.plot, dpi=300, width=12, height=12, units="in")
```

Plot results by Molecular Function level.    

```{r}
# Plotting by BP function
high_mid.GO.MF <- high_mid.GO %>%
  filter(over_represented_pvalue != "NA") %>%
  filter(numInCat >5) %>%
  filter(ontology == "MF") %>%
  arrange(., ontology, GOSlim_bin, over_represented_pvalue)

high_mid.GO.MF$GOSlim_bin <- as.factor(high_mid.GO.MF$GOSlim_bin)
high_mid.GO.MF$Pattern <- c("High in Mid Development")

#plot significantly enriched GO terms by Slim Category
GO_high_mid.plot.MF <-  ggplot(high_mid.GO.MF, aes(x = ontology, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 10),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 8),
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/Patterns/GOenrich_pattern_high_mid_MF.png", plot=GO_high_mid.plot.MF, dpi=300, width=12, height=12, units="in")
```





# **Plot all patterns together on one plot**  

Combine data frames together in order to generate one GoSeq plot with all terms across patterns  

Plot at Biological Process level.  

```{r}
go_all<-rbind(high_early.GO_plot, high_mid.GO_plot, high_late.GO_plot)

head(go_all)
dim(go_all)

levels(as.factor(go_all$Pattern))
```

Files are merged.  

Now plot.  

```{r}
#reorder x axis lifestages 
list_patterns<-c("High Early in Development", "High in Mid Development", "High Late in Development")
go_all$Pattern<-factor(as.character(go_all$Pattern), levels=list_patterns)

#plot significantly enriched GO terms by Slim Category across lifestages
GO_all.BP <-  ggplot(go_all, aes(x = Pattern, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  #geom_point(aes(fill=Pattern,size=over_represented_pvalue)) +
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 20),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 20),
  axis.text.x = element_text(angle=90, size=20),
  axis.title.x = element_blank(),
  axis.title.y = element_blank(), 
  legend.text = element_text(size=20), 
  legend.title = element_text(size=20), 
  legend.key.size = unit(1, 'cm'))

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/Patterns/GOenrich_all_patterns_BP.png", plot=GO_all.BP, dpi=300, width=20, height=30, units="in")
```

Plot at Molecular Function level.  

```{r}
go_all.MF<-rbind(high_early.GO.MF, high_mid.GO.MF, high_late.GO.MF)

head(go_all.MF)
dim(go_all.MF)

levels(as.factor(go_all$Pattern))
```

Files are merged.  

Now plot.  

```{r}
#reorder x axis lifestages 
list_patterns<-c("High Early in Development", "High in Mid Development", "High Late in Development")
go_all.MF$Pattern<-factor(as.character(go_all.MF$Pattern), levels=list_patterns)

#plot significantly enriched GO terms by Slim Category across lifestages
GO_all.MF <-  ggplot(go_all.MF, aes(x = Pattern, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 14),
  strip.text.x = element_text(size = 14),
  axis.text = element_text(size = 16),
  axis.text.x = element_text(angle=90, size=14),
  axis.title.x = element_blank(),
  axis.title.y = element_blank(), 
  legend.text = element_text(size=14), 
  legend.title = element_text(size=14), 
  legend.key.size = unit(1, 'cm'))

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/Patterns/GOenrich_all_patterns_MF.png", plot=GO_all.MF, dpi=300, width=30, height=49, units="in")

#plot "other functions for MF"
GO_all.MF.other <-  go_all.MF%>%
  filter(GOSlim_bin=="other molecular function")%>%
  
  ggplot(., aes(x = Pattern, y = term)) + 
  geom_tile(aes(fill =over_represented_pvalue)) + 
  facet_grid(GOSlim_bin ~ ., scales = "free_y", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
  strip.text.y = element_text(angle=0, size = 20),
  strip.text.x = element_text(size = 20),
  axis.text = element_text(size = 20),
  axis.text.x = element_text(angle=90, size=20),
  axis.title.x = element_blank(),
  axis.title.y = element_blank(), 
  legend.text = element_text(size=20), 
  legend.title = element_text(size=20), 
  legend.key.size = unit(1, 'cm')); GO_all.MF.other

ggsave(filename="Mcap2020/Figures/TagSeq/GOSeq/Patterns/GOenrich_all_patterns_MF_other.png", plot=GO_all.MF.other, dpi=300, width=30, height=30, units="in")

```










