---
title: "Analysis of 16S data output from mothur bioniformatic pipeline"
author: "Ariana S Huffmyer"
date: '2022'
output:
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("janitor" %in% rownames(installed.packages()) == 'FALSE') install.packages('janitor') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("phyloseq" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install('phyloseq') 
if ("stringr" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install('stringr')
if ("ape" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install('ape') 
if ("DESeq2" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install('DESeq2') 
if ("microViz" %in% rownames(installed.packages()) == 'FALSE') devtools::install_github("david-barnett/microViz@0.9.0") 
if ("vegan" %in% rownames(installed.packages()) == 'FALSE') install.packages('vegan') 

#load packages
library("ggplot2")
library('cowplot')
library("tidyverse")
library("janitor")
library("car")
library("phyloseq")
library("stringr")
library("ape")
library("DESeq2")
library("microViz")
library("vegan")
library(dendsort)
library(WGCNA)
library(circlize)
library(ComplexHeatmap)
```

# **Load data**  

Load data for use in analysis with PhyloSeq. Here we will import an OTU table and a taxonomy table.  

We will load "rare" files as those that were subsampled/rarefied.  

```{r}
otu.rare<-read.table("Mcap2020/Output/16S/mothur/TextFormat/mcap.opti_mcc.0.03.subsample.shared.txt", sep='\t', header=TRUE)

taxonomy<-read.table("Mcap2020/Output/16S/mothur/TextFormat/mcap.taxonomy.txt", sep='\t', header=TRUE)

rarefaction<-read.table("Mcap2020/Output/16S/mothur/TextFormat/mcap.opti_mcc.groups.rarefaction.txt", sep='\t', header=TRUE)
```

Load sample information.  
```{r}
metadata<-read.table("Mcap2020/Data/16S/mcap_metadata.txt", sep='\t', header=TRUE)

rownames(metadata)<-metadata$sample

#Replace metamorphosed "recruit" with "polyp"  
metadata<-metadata%>%
  select(!sample)%>%
  mutate(lifestage=if_else(lifestage=="Metamorphosed Recruit (183 hpf)", "Metamorphosed Polyp (183 hpf)", lifestage))%>%
  mutate(lifestage=if_else(lifestage=="Metamorphosed Recruit (231 hpf)", "Metamorphosed Polyp (231 hpf)", lifestage))


```

Set default plot theme.  

```{r}
theme_set(theme_bw())
```

Format columns of our taxonomy file. We need to separate by ; and remove the quotes. We also need to remove the confidence values.  
```{r}
head(taxonomy)

taxonomy<-taxonomy%>%
  mutate(Taxonomy=str_replace_all(string=Taxonomy, pattern="\\(\\d*\\)", replacement="")) %>% #remove parentheses with numbers
	mutate(Taxonomy=str_replace_all(string=Taxonomy, pattern=";$", replacement="")) %>% #remove semicolons
 # mutate(Taxonomy=str_replace_all(string=Taxonomy, pattern="*", replacement="")) %>% #remove quotes
  #mutate(Taxonomy=str_replace_all(string=Taxonomy, pattern="""$", replacement="")) %>% #remove quotes
	separate(Taxonomy, into=c("Domain", "Phylum", "Class", "Order", "Family", "Genus"), sep=";") #separate columns

head(taxonomy)

rownames(taxonomy)<-taxonomy$OTU
taxonomy<-as.matrix(taxonomy)

```

# **Heatmaps and beta diversity**   

## Prepare data  

Prepare the matrix to be all numeric with taxa in rows.  
```{r}
otu.rare<-otu.rare%>%
  select(!label)%>%
  select(!numOtus)

rownames(otu.rare)<-otu.rare$Group

otu.rare<-otu.rare%>%
  select(!Group)

otu.rare<-t(otu.rare)
```

Combine the OTU and taxonomic information.  
```{r}
OTU_rare = otu_table(otu.rare, taxa_are_rows = TRUE)
TAX = tax_table(taxonomy)
OTU_rare
```

Load into phyloseq.  
```{r}
physeq_rare = phyloseq(OTU_rare, TAX)
physeq_rare
```

Plot an abundance table at the different taxonomic levels.     
```{r}
plot_bar(physeq_rare, fill = "Domain")+theme(legend.position="bottom")
plot_bar(physeq_rare, fill = "Phylum")+theme(legend.position="bottom")
plot_bar(physeq_rare, fill = "Class")+theme(legend.position="bottom")
#plot_bar(physeq_rare, fill = "Order")+theme(legend.position="bottom")
#plot_bar(physeq_rare, fill = "Family")+theme(legend.position="bottom")
#plot_bar(physeq_rare, fill = "Genus")+theme(legend.position="bottom")

```

Plot a tree of OTU's.  
```{r}
random_tree_rare = rtree(ntaxa(physeq_rare), rooted=TRUE, tip.label=taxa_names(physeq_rare))
plot(random_tree_rare)
```

Add metdata 
```{r}
list<-sample_names(physeq_rare)

metadata_rare<-read.table("Mcap2020/Data/16S/mcap_metadata.txt", sep='\t', header=TRUE)

rownames(metadata_rare)<-metadata_rare$sample

metadata_rare<-metadata_rare%>%
  filter(sample %in% list)

metadata_rare<-metadata_rare%>%
  select(!sample)%>%
  mutate(lifestage=if_else(lifestage=="Metamorphosed Recruit (183 hpf)", "Metamorphosed Polyp (183 hpf)", lifestage))%>%
  mutate(lifestage=if_else(lifestage=="Metamorphosed Recruit (231 hpf)", "Metamorphosed Polyp (231 hpf)", lifestage))
```

Add sample data 
```{r}
sampledata = sample_data(data.frame(
  lifestage = metadata_rare$lifestage,
  group = metadata_rare$group, 
  hpf = metadata_rare$hpf,
  row.names=sample_names(physeq_rare),
  stringsAsFactors=FALSE
))
sampledata
```

Merge the tree, sample information, and OTU table.  
```{r}
physeq1_rare = merge_phyloseq(physeq_rare, sampledata, random_tree_rare)
physeq1_rare
```

Now rebuild phyloseq.  
```{r}
physeq2_rare = phyloseq(OTU_rare, TAX, sampledata, random_tree_rare)
physeq2_rare
```

Check that they are identical. Should return TRUE.  
```{r}
identical(physeq1_rare, physeq2_rare)
```

Filter out samples with low numbers in groups from phyloseq object.   
```{r}
physeq1_rare<-ps_filter(physeq1_rare, lifestage != "Metamorphosed Polyp (183 hpf)")
physeq1_rare<-ps_filter(physeq1_rare, lifestage != "Egg (1 hpf)")
physeq1_rare<-ps_filter(physeq1_rare, lifestage != "Embryo (5 hpf)")
#physeq1_rare<-ps_filter(physeq1_rare, lifestage != "Embryo (38 hpf)")
```

Now build a tree with the combined data.  
```{r}
plot_tree(physeq1_rare, color="lifestage", label.tips="taxa_names", ladderize="left", plot.margin=0.3)
```

Plot relative abundance now with group information.  

```{r}
bar1<-plot_bar(physeq1_rare, fill = "Class")+theme(legend.position="bottom")+facet_grid(~lifestage, scales="free_x"); bar1
#ggsave(plot=bar1, file="Mcap2020/Figures/16S/relabund_class_rare.png", h=6, w=18)

bar2<-plot_bar(physeq1_rare, fill = "Phylum")+theme(legend.position="bottom")+facet_grid(~lifestage, scales="free_x"); bar2
#ggsave(plot=bar2, file="Mcap2020/Figures/16S/relabund_phylum_rare.png", h=6, w=18)

bar3<-plot_bar(physeq1_rare, fill = "Order")+theme(legend.position="bottom")+facet_grid(~lifestage, scales="free_x"); bar3
#ggsave(plot=bar3, file="Mcap2020/Figures/16S/relabund_order_rare.png", h=9, w=18)

bar4<-plot_bar(physeq1_rare, fill = "Family")+theme(legend.position="bottom")+facet_grid(~lifestage, scales="free_x"); bar4
#ggsave(plot=bar4, file="Mcap2020/Figures/16S/relabund_family_rare.png", h=12, w=18)

bar5<-plot_bar(physeq1_rare, fill = "Genus")+theme(legend.position="bottom")+facet_grid(~lifestage, scales="free_x"); bar5
#ggsave(plot=bar5, file="Mcap2020/Figures/16S/relabund_genus_rare.png", h=25, w=18)
```

Plot relative abundance at the phylum level.  
```{r}
dat <- physeq1_rare %>% tax_glom(taxrank = "Phylum") %>% psmelt()
head(dat)
```

```{r}
df<-dat %>% 
  group_by(lifestage, Phylum) %>% 
  summarise(TotalAbundance=sum(Abundance))%>% #calculate total abundance of each phylum
  group_by(lifestage) %>%
  mutate(PercRelAbundance=TotalAbundance/sum(TotalAbundance)*100)%>% #calculate relative abundance
  select(!TotalAbundance)

df$Phylum <- gsub('[^[:alnum:] ]','',df$Phylum)

#change to a matrix with phylum in rows and lifestage in columns
df<-df%>%
  spread(key=Phylum, value=PercRelAbundance)

#replace na with 0, since 0 abundance was calculated as na in step above b/c dividing by 0
df[is.na(df)] <- 0

rownames(df)<-df$lifestage
lifestage_list<-df$lifestage

dim(df)

df<-as.matrix(df)

df<-df[,-1]

taxa_list<-colnames(df)

df_mat <- matrix(as.numeric(df),    # Convert to numeric matrix
                  ncol = ncol(df))                                   # Print numeric matrix

colnames(df_mat)<-taxa_list
rownames(df_mat)<-lifestage_list

df_mat<-t(df_mat)
```

Plot heatmap of phylum relative abundance.  
```{r}

lifestage_order<-c("Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)")

row_dend = dendsort(hclust(dist(df_mat)))
col_dend = dendsort(hclust(dist(t(df_mat))))

#col_fun = colorRamp2(c(0, 25, 50, 75, 100), c("purple", "white", "cyan", "white", "darkgreen"))
col_fun = colorRamp2(c(0, 25, 50, 75, 100), hcl_palette = "Blue-Red 3")

pdf(file = "Mcap2020/Figures/16S/16S-Phylum-Heatmap.pdf", height = 8, width = 8)
ComplexHeatmap::Heatmap(df_mat, name = "Rel. Abun. (%)", row_title = "", column_title = "16S Relative Abundance (Phylum)", 
        col = col_fun, 
        row_names_side = "left", 
        row_dend_side = "right",
        width = unit(4, "in"), 
        height = unit(4, "in"), 
        column_dend_height = unit(.5, "in"),
        row_dend_width = unit(.5, "in"),
        column_dend_reorder = FALSE,
        row_dend_reorder = TRUE,
        row_gap = unit(2.5, "mm"),
        clustering_distance_rows = "euclidean",
        clustering_method_rows = "complete",
        border = TRUE, 
        column_names_gp =  gpar(fontsize = 12, border=FALSE),
        column_names_rot = 35,
        cluster_rows = row_dend, 
        #cluster_columns = col_dend,
        column_order = lifestage_order, 
        row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE))
dev.off()

```

Display as % abundance.    

```{r}
physeq1_rare_f = transform_sample_counts(physeq1_rare, function(x) 100 * x/sum(x))
```


Plot relative abundance at the family level.  
```{r}
dat_fam <- physeq1_rare %>% tax_glom(taxrank = "Family") %>% psmelt()
head(dat_fam)
```

```{r}
df_fam<-dat_fam %>% 
  group_by(lifestage, Family) %>% 
  summarise(TotalAbundance=sum(Abundance))%>% #calculate total abundance of each phylum
  group_by(lifestage) %>%
  mutate(PercRelAbundance=TotalAbundance/sum(TotalAbundance)*100)%>% #calculate relative abundance
  select(!TotalAbundance)%>%
  filter(PercRelAbundance>5)

df_fam$Family <- gsub('[^[:alnum:] ]','',df_fam$Family)

#change to a matrix with phylum in rows and lifestage in columns
df_fam<-df_fam%>%
  spread(key=Family, value=PercRelAbundance)

#replace na with 0, since 0 abundance was calculated as na in step above b/c dividing by 0
df_fam[is.na(df_fam)] <- 0

rownames(df_fam)<-df_fam$lifestage
lifestage_list<-df_fam$lifestage

dim(df_fam)

df_fam<-as.matrix(df_fam)

df_fam<-df_fam[,-1]

taxa_list<-colnames(df_fam)

df_mat_fam <- matrix(as.numeric(df_fam),    # Convert to numeric matrix
                  ncol = ncol(df_fam))                                   # Print numeric matrix

colnames(df_mat_fam)<-taxa_list
rownames(df_mat_fam)<-lifestage_list

df_mat_fam<-t(df_mat_fam)
```

Plot heatmap of family relative abundance.  
```{r}

lifestage_order<-c("Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)")

row_dend = dendsort(hclust(dist(df_mat_fam)))
col_dend = dendsort(hclust(dist(t(df_mat_fam))))

#col_fun = colorRamp2(c(0, 50, 100), c("blue", "white", "red"))
col_fun = colorRamp2(c(0, 25, 50), hcl_palette = "Blue-Red 3")

pdf(file = "Mcap2020/Figures/16S/16S-Family-Heatmap.pdf", height = 10, width = 10)
ComplexHeatmap::Heatmap(df_mat_fam, name = "Rel. Abun. (%)", row_title = "", column_title = "16S Relative Abundance (Family)", 
        col = col_fun, 
        row_names_side = "left", 
        row_dend_side = "right",
        width = unit(5, "in"), 
        height = unit(5, "in"), 
        column_dend_height = unit(.5, "in"),
        row_dend_width = unit(.5, "in"),
        column_dend_reorder = FALSE,
        row_dend_reorder = TRUE,
        row_gap = unit(2.5, "mm"),
        clustering_distance_rows = "euclidean",
        clustering_method_rows = "complete",
        border = TRUE, 
        column_names_gp =  gpar(fontsize = 12, border=FALSE),
        column_names_rot = 35,
        cluster_rows = row_dend, 
        #cluster_columns = col_dend,
        column_order = lifestage_order, 
        row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE))
dev.off()

```


## Plot PCoA and conduct PERMANOVA  

Prepare PCoA ordination.  
```{r}
ps.ordb <- ordinate(physeq1_rare, "PCoA", "bray")
p2b = plot_ordination(physeq1_rare, ps.ordb, type="samples", color="lifestage"); p2b 
```

Display levels of lifestages.  
```{r}
levels(as.factor(metadata_rare$lifestage))
#Egg: "#8C510A"
#Embryo: "#DFC27D"
#Larvae: "#80CDC1"
#Polyp Metamorphosed: "#003C30"
#Recruit Attached: "#BA55D3"
```

Output PCoA plot for visualization.  
```{r}
pcoa_viz<-p2b  + 
  geom_point(aes(colour=lifestage, shape=lifestage), size=4) + 
  ggtitle("PCoA")+
  
  scale_colour_manual(name="Lifestage", 
                      values = c("Attached Recruit (231 hpf)"="#BA55D3", "Embryo (38 hpf)"="#DFC27D", "Embryo (65 hpf)"="#DFC27D", "Larvae (93 hpf) "="#80CDC1", "Larvae (163 hpf)"="#80CDC1", "Larvae (231 hpf) "="#80CDC1", "Metamorphosed Polyp (231 hpf)"="#003C30", "Attached Recruit (183 hpf) "="#BA55D3"),
                      breaks = c("Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)"), 
                     labels = c( "Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)")) + 
  
  scale_shape_manual(name="Lifestage", 
                     values = c("Attached Recruit (231 hpf)"=13, "Embryo (38 hpf)"=3, "Embryo (65 hpf)"=4, "Larvae (93 hpf) "=5, "Larvae (163 hpf)"=6, "Larvae (231 hpf) "=8, "Metamorphosed Polyp (231 hpf)"=11, "Attached Recruit (183 hpf) "=12), 
                     breaks = c( "Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)"), 
                     labels = c( "Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)")) +
  
  theme_classic() + 
  theme(
    panel.background=element_blank()
  ); pcoa_viz

ggsave(plot=pcoa_viz, "Mcap2020/Figures/16S/pcoa_16S.jpeg", h=6, w=8)
```

Conduct a permanova in the vegan package.  

```{r}
samples <- as(sample_data(physeq1_rare), "data.frame")

perm1<-adonis2(phyloseq::distance(physeq1_rare, method="bray") ~ lifestage,
       data = samples)

perm1
```

Permanova suggests there is significant separation by lifestage in 16S profiles (p=0.001). 

Permutation test for adonis under reduced model
Terms added sequentially (first to last)
Permutation: free
Number of permutations: 999

adonis2(formula = phyloseq::distance(physeq1_rare, method = "bray") ~ lifestage, data = samples)
          Df SumOfSqs      R2      F Pr(>F)    
lifestage  7   4.8793 0.64321 4.3781  0.001 ***
Residual  17   2.7066 0.35679                  
Total     24   7.5859 1.00000     


## Plot NMDS 

Prepare PCoA ordination.  
```{r}
ps.ordb.nmds <- ordinate(physeq1_rare, "NMDS", "bray")
p2b.nmds = plot_ordination(physeq1_rare, ps.ordb.nmds, type="samples", color="lifestage")
```

Display levels of lifestages.  
```{r}
levels(as.factor(metadata_rare$lifestage))
#Egg: "#8C510A"
#Embryo: "#DFC27D"
#Larvae: "#80CDC1"
#Polyp Metamorphosed: "#003C30"
#Recruit Attached: "#BA55D3"
```

Output PCoA plot for visualization.  
```{r}
pcoa_viz.nmds<-p2b.nmds  + 
  geom_point(aes(colour=lifestage, shape=lifestage), size=4) + 
  ggtitle("NMDS")+
  
  scale_colour_manual(name="Lifestage", 
                      values = c("Attached Recruit (231 hpf)"="#BA55D3", "Embryo (38 hpf)"="#DFC27D", "Embryo (65 hpf)"="#DFC27D", "Larvae (93 hpf) "="#80CDC1", "Larvae (163 hpf)"="#80CDC1", "Larvae (231 hpf) "="#80CDC1", "Metamorphosed Polyp (231 hpf)"="#003C30", "Attached Recruit (183 hpf) "="#BA55D3"),
                      breaks = c("Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)"), 
                     labels = c( "Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)")) + 
  
  scale_shape_manual(name="Lifestage", 
                     values = c("Attached Recruit (231 hpf)"=13, "Embryo (38 hpf)"=3, "Embryo (65 hpf)"=4, "Larvae (93 hpf) "=5, "Larvae (163 hpf)"=6, "Larvae (231 hpf) "=8, "Metamorphosed Polyp (231 hpf)"=11, "Attached Recruit (183 hpf) "=12), 
                     breaks = c( "Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)"), 
                     labels = c( "Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)")) +
  
  theme_classic() + 
  theme(
    panel.background=element_blank()
  ); pcoa_viz.nmds

ggsave(plot=pcoa_viz.nmds, "Mcap2020/Figures/16S/nmds_16S.jpeg", h=6, w=8)
ggsave(plot=pcoa_viz.nmds, "Mcap2020/Figures/16S/nmds_16S.pdf", h=6, w=8)
```

Conduct a permanova in the vegan package (same as test in above PCoA section).  

```{r}
samples <- as(sample_data(physeq1_rare), "data.frame")

perm1<-adonis2(phyloseq::distance(physeq1_rare, method="bray") ~ lifestage,
       data = samples)

perm1
```

## Plot Heatmap of relative abundance at the family/genus level   

Filter out low frequency taxa <1% relative abundance.  
```{r}
physeq1_rare_f

physeq1_rare_prune = filter_taxa(physeq1_rare_f, function(x) mean(x) > 0.5, TRUE)

taxa_names(physeq1_rare_prune) 
```

There were 1364 taxa before filtering out taxa <1% and 36 taxa after filtering.  

Summarize by lifestage calculating means across replicates of each life stage.  
```{r}
mergedGP = merge_samples(physeq1_rare_prune, "lifestage")

lifestage_order<-c("Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)")
```

Replace OTU number with taxa names.  

View current data frame.  
```{r}
table<-as.matrix(t(otu_table(mergedGP)))
table<-as.matrix(as.data.frame(table))
colnames(table)
rownames(table)
```

Extract taxa names at the level of Family.  
```{r}
tax_table<-as.matrix(t(tax_table(mergedGP)))
tax_table<-as.matrix(t(as.data.frame(tax_table)))
colnames(tax_table)
rownames(tax_table)
tax_table<-as.data.frame(tax_table)
```
 
```{r}
#replace OTU with name from taxonomy table  
table_names<-as.data.frame(table)
table_names$OTU<-rownames(table)

table_names$Family<-tax_table$Family[match(table_names$OTU, tax_table$OTU)]
table_names$Genus<-tax_table$Genus[match(table_names$OTU, tax_table$OTU)]

#remove non alphabetic characters from Family and Genus columns 
table_names$Family <- gsub('[^[:alnum:] ]','',table_names$Family)

table_names$Taxa<-paste(table_names$OTU, "-", table_names$Family, ":", table_names$Genus)

#add to data frame 
rownames(table_names)<-table_names$Taxa

table_names<-table_names%>%
  select(!Family)%>%
  select(!Taxa)%>%
  select(!Genus)%>%
  select(!OTU)

table_names<-as.matrix(table_names)
```

Draw by lifestage with Complex Heatmap. 
```{r}
row_dend = dendsort(hclust(dist(table)))
col_dend = dendsort(hclust(dist(t(table))))

#col_fun = colorRamp2(c(0, 50, 100), c("blue", "white", "red"))
col_fun = colorRamp2(c(0, 25, 50, 100), hcl_palette = "Blue-Red 3")

pdf(file = "Mcap2020/Figures/16S/16S-ComplexHeatmap.pdf", height = 10, width = 18)
ComplexHeatmap::Heatmap(table_names, name = "Rel. Abun. (%)", row_title = "", column_title = "16S Relative Abundance (OTU - Family:Genus) Taxa >0.5%", 
        col = col_fun, 
        row_names_side = "left", 
        row_dend_side = "right",
        width = unit(6, "in"), 
        height = unit(8, "in"), 
        column_dend_height = unit(.5, "in"),
        row_dend_width = unit(.5, "in"),
        column_dend_reorder = FALSE,
        row_dend_reorder = TRUE,
        row_gap = unit(2.5, "mm"),
        clustering_distance_rows = "euclidean",
        clustering_method_rows = "complete",
        border = TRUE, 
        column_names_gp =  gpar(fontsize = 12, border=FALSE),
        column_names_rot = 35,
        cluster_rows = row_dend, 
        #cluster_columns = col_dend,
        column_order = lifestage_order, 
        row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE))
dev.off()

```

# **Alpha Diversity**  

```{r}
GP2 <- prune_taxa(taxa_sums(physeq1_rare) > 0, physeq1_rare)
```

Plot all richness metrics.  
```{r}
plot_richness(GP2)
```

Plot the two most commonly used - Shannon and Inverse Simpson.  
```{r}
alpha1<-plot_richness(GP2, measures=c("InvSimpson", "Shannon"), x="lifestage", color="lifestage", nrow=2)

#reorder factors 
alpha1$data$lifestage <- as.character(alpha1$data$lifestage)
alpha1$data$lifestage <- factor(alpha1$data$lifestage, levels=lifestage_order)

alpha1_plot <- alpha1 + geom_point(size=3, position=position_jitterdodge(0.6)) + theme_classic() + scale_color_manual() + scale_color_manual(name="Lifestage", 
                      values = c("Attached Recruit (231 hpf)"="#BA55D3", "Embryo (38 hpf)"="#DFC27D", "Embryo (65 hpf)"="#DFC27D", "Larvae (93 hpf) "="#80CDC1", "Larvae (163 hpf)"="#80CDC1", "Larvae (231 hpf) "="#80CDC1", "Metamorphosed Polyp (231 hpf)"="#003C30", "Attached Recruit (183 hpf) "="#BA55D3"),
                      breaks = c("Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)"), 
                     labels = c( "Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)")) + 
  theme(axis.text.x=element_text(angle=45, hjust=1, color="black")) + 
  theme(axis.title.x=element_blank())+
  theme(legend.position="none");alpha1_plot

ggsave(plot=alpha1_plot, file="Mcap2020/Figures/16S/alpha_rare.png", width=4, height=7)
ggsave(plot=alpha1_plot, file="Mcap2020/Figures/16S/alpha_rare.pdf", width=4, height=7)
```

Similar to the full dataset, we see increases in alpha diversity across development with higher values in attached recruits.  

Test for differences in alpha diversity.  
```{r}
metadata <- as(sample_data(GP2), "data.frame")
metadata$sample<-rownames(metadata)

alpha_rare<-estimate_richness(GP2, measures=c("InvSimpson", "Shannon", "Chao1"))
alpha_rare$sample<-rownames(alpha_rare)

alpha_rare<-full_join(metadata, alpha_rare, by="sample")

summary(aov(Shannon~lifestage, data=alpha_rare))
summary(aov(InvSimpson~lifestage, data=alpha_rare))
summary(aov(Chao1~lifestage, data=alpha_rare))
```

There are significant differences in alpha diversity between lifestages for all metrics.    

Export relative change in alpha diversity for integrative analyses. Calculate relative change for all lifestages compared to the first time point in this dataset (embryos at 38 hpf). We will use the Shannon Index for this calculation.    

```{r}

alpha_rel_change<-alpha_rare%>%
  select(lifestage, sample, Shannon)%>%
  group_by(lifestage)%>%
  summarise(mean_alpha=mean(Shannon)) #calculate means

alpha_rel_change$lifestage<-str_trim(alpha_rel_change$lifestage, side = c("right")) #remove whitespace at end of strings
alpha_rel_change$lifestage<-factor(alpha_rel_change$lifestage, levels=c("Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf)", "Larvae (163 hpf)", "Larvae (231 hpf)", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf)", "Attached Recruit (231 hpf)"))

#calculate relative change to the first time point  
alpha_rel_change<-alpha_rel_change%>%
  arrange(lifestage)%>%
  mutate(bacteria_alpha_diversity=((mean_alpha-dplyr::first(.$mean_alpha))/dplyr::first(.$mean_alpha)))%>%
  select(lifestage, bacteria_alpha_diversity)%>%
  write_csv("Mcap2020/Data/Integration/RelativeChange/16S_alpha_relchange.csv")
```


# **Plot unknowns** 

Plot data at the level of phylum in the rarefied dataset. Bacteria_unclassified are our unknowns.      
```{r}
plot_bar(physeq1_rare, fill="Phylum", facet_grid=~Phylum)
```

We have some unclassified bacteria, but a higher proportion of proteobacteria and bacteriodes.  

# **Export matrix for multi-omic analyses**  +

Export a matrix of the relative abundance of taxa >0.5% relative abundance for multiomic analysis at the level of sample.    

Filter out low frequency taxa <0.5% relative abundance.  
```{r}
physeq1_rare_f

physeq1_rare_prune = filter_taxa(physeq1_rare_f, function(x) mean(x) > 0.5, TRUE)

taxa_names(physeq1_rare_prune) 
```

There were 1364 taxa before filtering out taxa <0.5% and 36 taxa after filtering.  

Replace OTU number with taxa names.  

View current data frame.  
```{r}
table<-as.matrix(otu_table(physeq1_rare_prune))
table<-as.matrix(as.data.frame(table))
colnames(table)
rownames(table)
```

Extract taxa names at the level of Family.  
```{r}
tax_table<-as.matrix(t(tax_table(physeq1_rare_prune)))
tax_table<-as.matrix(t(as.data.frame(tax_table)))
colnames(tax_table)
rownames(tax_table)
tax_table<-as.data.frame(tax_table)
```
 
```{r}
#replace OTU with name from taxonomy table  
table_names<-as.data.frame(table)
table_names$OTU<-rownames(table)

table_names$Family<-tax_table$Family[match(table_names$OTU, tax_table$OTU)]
table_names$Genus<-tax_table$Genus[match(table_names$OTU, tax_table$OTU)]

#remove non alphabetic characters from Family and Genus columns 
table_names$Family <- gsub('[^[:alnum:] ]','',table_names$Family)

table_names$Taxa<-paste(table_names$OTU, "-", table_names$Family, ":", table_names$Genus)

#add to data frame 
rownames(table_names)<-table_names$Taxa

table_names<-table_names%>%
  select(!Family)%>%
  select(!Taxa)%>%
  select(!Genus)%>%
  select(!OTU)

table_names<-as.matrix(table_names)

rownames(table_names)
colnames(table_names)

table_16s<-as.data.frame(table_names)
table_16s$taxa<-rownames(table_16s)
```

The dataframe now has OTU with Genus and Family information in rownames and sample name in columns.  

Export for integration analysis.  

```{r}
write_csv(table_16s, "Mcap2020/Data/Integration/16s_matrix.csv")
```




