---
title: "Analysis of 16S data output from mothur bioniformatic pipeline"
author: "Ariana S Huffmyer"
date: '2022'
output:
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("janitor" %in% rownames(installed.packages()) == 'FALSE') install.packages('janitor') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("phyloseq" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install('phyloseq') 
if ("stringr" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install('stringr')
if ("ape" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install('ape') 
if ("DESeq2" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install('DESeq2') 
if ("microViz" %in% rownames(installed.packages()) == 'FALSE') devtools::install_github("david-barnett/microViz@0.9.0") 
if ("vegan" %in% rownames(installed.packages()) == 'FALSE') install.packages('vegan') 

#load packages
library("ggplot2")
library('cowplot')
library("tidyverse")
library("janitor")
library("car")
library("phyloseq")
library("stringr")
library("ape")
library("DESeq2")
library("microViz")
library("vegan")
library(dendsort)
library(WGCNA)
library(circlize)
library(ComplexHeatmap)
```

# **Load data**  

Load data for use in analysis with PhyloSeq. Here we will import an OTU table and a taxonomy table.  

We will load "rare" files as those that were subsampled/rarefied.  

```{r}
otu.rare<-read.table("Mcap2020/Output/16S/mothur/TextFormat/mcap.opti_mcc.0.03.subsample.shared.txt", sep='\t', header=TRUE)
otu.full<-read.table("Mcap2020/Output/16S/mothur/TextFormat/mcap.opti_mcc.shared.txt", sep='\t', header=TRUE)

taxonomy<-read.table("Mcap2020/Output/16S/mothur/TextFormat/mcap.taxonomy.txt", sep='\t', header=TRUE)

rarefaction<-read.table("Mcap2020/Output/16S/mothur/TextFormat/mcap.opti_mcc.groups.rarefaction.txt", sep='\t', header=TRUE)
```

Load sample information.  
```{r}
metadata<-read.table("Mcap2020/Data/16S/mcap_metadata.txt", sep='\t', header=TRUE)

rownames(metadata)<-metadata$sample

metadata<-metadata%>%
  select(!sample)
  
```

Set default plot theme.  

```{r}
theme_set(theme_bw())
```

Format columns of our taxonomy file. We need to separate by ; and remove the quotes. We also need to remove the confidence values.  
```{r}
head(taxonomy)

taxonomy<-taxonomy%>%
  mutate(Taxonomy=str_replace_all(string=Taxonomy, pattern="\\(\\d*\\)", replacement="")) %>% #remove parentheses with numbers
	mutate(Taxonomy=str_replace_all(string=Taxonomy, pattern=";$", replacement="")) %>% #remove semicolons
 # mutate(Taxonomy=str_replace_all(string=Taxonomy, pattern="*", replacement="")) %>% #remove quotes
  #mutate(Taxonomy=str_replace_all(string=Taxonomy, pattern="""$", replacement="")) %>% #remove quotes
	separate(Taxonomy, into=c("Domain", "Phylum", "Class", "Order", "Family", "Genus"), sep=";") #separate columns

head(taxonomy)

rownames(taxonomy)<-taxonomy$OTU
taxonomy<-as.matrix(taxonomy)

```


# **Tree and heatmap for "full" dataset**  

Prepare the matrix to be all numeric with taxa in rows.  
```{r}
otu.full<-otu.full%>%
  select(!label)%>%
  select(!numOtus)

rownames(otu.full)<-otu.full$Group

otu.full<-otu.full%>%
  select(!Group)

otu.full<-t(otu.full)
```

Prepare the taxonomy table.  
```{r}
rownames(taxonomy)<-taxonomy$OTU

taxonomy<-taxonomy%>%
  select(!Size)%>%
  select(!OTU)

taxonomy<-as.matrix(taxonomy)
```

Combine the OTU and taxonomic information.  
```{r}
OTU_full = otu_table(otu.full, taxa_are_rows = TRUE)
TAX = tax_table(taxonomy)
OTU_full
```

Load into phyloseq.  
```{r}
physeq = phyloseq(OTU_full, TAX)
physeq
```

Plot an abundance table at the different taxonomic levels.    
```{r}
plot_bar(physeq, fill = "Domain")+theme(legend.position="bottom")
plot_bar(physeq, fill = "Phylum")+theme(legend.position="bottom")
#plot_bar(physeq, fill = "Class")+theme(legend.position="bottom")
#plot_bar(physeq, fill = "Order")+theme(legend.position="bottom")
#plot_bar(physeq, fill = "Family")+theme(legend.position="bottom")
```

Plot a tree of OTU's.  
```{r}
random_tree = rtree(ntaxa(physeq), rooted=TRUE, tip.label=taxa_names(physeq))
plot(random_tree)
```

Generate sample information.  
```{r}
sampledata = sample_data(data.frame(
  lifestage = metadata$lifestage,
  group = metadata$group, 
  hpf = metadata$hpf,
  row.names=sample_names(physeq),
  stringsAsFactors=FALSE
))
sampledata
```

Merge the tree, sample information, and OTU table.  
```{r}
physeq1 = merge_phyloseq(physeq, sampledata, random_tree)
physeq1
```

Now rebuild phyloseq.  
```{r}
physeq2 = phyloseq(OTU_full, TAX, sampledata, random_tree)
physeq2
```

Check that they are identical. Should return TRUE.  
```{r}
identical(physeq1, physeq2)
```

Filter out samples with low numbers in groups from phyloseq object.  
```{r}
physeq1<-ps_filter(physeq1, lifestage != "Metamorphosed Recruit (183 hpf)")
physeq1<-ps_arrange(physeq1, hpf)
```

Now build a tree with the combined data.  
```{r}
plot_tree(physeq1, color="group", label.tips="taxa_names", ladderize="left", plot.margin=0.3)
```

Plot a heatmap.  

```{r}
plot_heatmap(physeq1, "NMDS", "bray", "lifestage", "Family", low="#66CCFF", high="#000033", na.value="white")
```

# **Analyze "rarefied" dataset**   

## Prepare data  

Prepare the matrix to be all numeric with taxa in rows.  
```{r}
otu.rare<-otu.rare%>%
  select(!label)%>%
  select(!numOtus)

rownames(otu.rare)<-otu.rare$Group

otu.rare<-otu.rare%>%
  select(!Group)

otu.rare<-t(otu.rare)
```

Combine the OTU and taxonomic information.  
```{r}
OTU_rare = otu_table(otu.rare, taxa_are_rows = TRUE)
TAX = tax_table(taxonomy)
OTU_rare
```

Load into phyloseq.  
```{r}
physeq_rare = phyloseq(OTU_rare, TAX)
physeq_rare
```

Plot an abundance table at the different taxonomic levels.    
```{r}
plot_bar(physeq_rare, fill = "Domain")+theme(legend.position="bottom")
plot_bar(physeq_rare, fill = "Phylum")+theme(legend.position="bottom")
plot_bar(physeq_rare, fill = "Class")+theme(legend.position="bottom")
plot_bar(physeq_rare, fill = "Order")+theme(legend.position="bottom")
plot_bar(physeq_rare, fill = "Family")+theme(legend.position="bottom")
plot_bar(physeq_rare, fill = "Genus")+theme(legend.position="bottom")
```

Plot a tree of OTU's.  
```{r}
random_tree_rare = rtree(ntaxa(physeq_rare), rooted=TRUE, tip.label=taxa_names(physeq_rare))
plot(random_tree_rare)
```

Add metdata 
```{r}
list<-sample_names(physeq_rare)

metadata_rare<-read.table("Mcap2020/Data/16S/mcap_metadata.txt", sep='\t', header=TRUE)

rownames(metadata_rare)<-metadata_rare$sample

metadata_rare<-metadata_rare%>%
  filter(sample %in% list)

metadata_rare<-metadata_rare%>%
  select(!sample)
```

Add sample data 
```{r}
sampledata = sample_data(data.frame(
  lifestage = metadata_rare$lifestage,
  group = metadata_rare$group, 
  hpf = metadata_rare$hpf,
  row.names=sample_names(physeq_rare),
  stringsAsFactors=FALSE
))
sampledata
```

Merge the tree, sample information, and OTU table.  
```{r}
physeq1_rare = merge_phyloseq(physeq_rare, sampledata, random_tree_rare)
physeq1_rare
```

Now rebuild phyloseq.  
```{r}
physeq2_rare = phyloseq(OTU_rare, TAX, sampledata, random_tree_rare)
physeq2_rare
```

Check that they are identical. Should return TRUE.  
```{r}
identical(physeq1_rare, physeq2_rare)
```

Filter out samples with low numbers in groups from phyloseq object.   
```{r}
physeq1_rare<-ps_filter(physeq1_rare, lifestage != "Metamorphosed Recruit (183 hpf)")
physeq1_rare<-ps_filter(physeq1_rare, lifestage != "Egg (1 hpf)")
physeq1_rare<-ps_filter(physeq1_rare, lifestage != "Embryo (5 hpf)")
#physeq1_rare<-ps_filter(physeq1_rare, lifestage != "Embryo (38 hpf)")
```

Now build a tree with the combined data.  
```{r}
plot_tree(physeq1_rare, color="lifestage", label.tips="taxa_names", ladderize="left", plot.margin=0.3)
```

Plot relative abundance now with group information.  

```{r}
bar1<-plot_bar(physeq1_rare, fill = "Class")+theme(legend.position="bottom")+facet_grid(~lifestage, scales="free_x"); bar1
#ggsave(plot=bar1, file="Mcap2020/Figures/16S/relabund_class_rare.png", h=6, w=18)

bar2<-plot_bar(physeq1_rare, fill = "Phylum")+theme(legend.position="bottom")+facet_grid(~lifestage, scales="free_x"); bar2
#ggsave(plot=bar2, file="Mcap2020/Figures/16S/relabund_phylum_rare.png", h=6, w=18)

bar3<-plot_bar(physeq1_rare, fill = "Order")+theme(legend.position="bottom")+facet_grid(~lifestage, scales="free_x"); bar3
#ggsave(plot=bar3, file="Mcap2020/Figures/16S/relabund_order_rare.png", h=9, w=18)

bar4<-plot_bar(physeq1_rare, fill = "Family")+theme(legend.position="bottom")+facet_grid(~lifestage, scales="free_x"); bar4
#ggsave(plot=bar4, file="Mcap2020/Figures/16S/relabund_family_rare.png", h=12, w=18)

bar5<-plot_bar(physeq1_rare, fill = "Genus")+theme(legend.position="bottom")+facet_grid(~lifestage, scales="free_x"); bar5
#ggsave(plot=bar5, file="Mcap2020/Figures/16S/relabund_genus_rare.png", h=25, w=18)
```

Display as % abundance and filter out taxa <1% abundance.    

```{r}
physeq1_rare_f = transform_sample_counts(physeq1_rare, function(x) 100 * x/sum(x))
```

## Plot PCoA and conduct PERMANOVA  

Prepare PCoA ordination.  
```{r}
ps.ordb <- ordinate(physeq1_rare_f, "PCoA", "bray")
p2b = plot_ordination(physeq1_rare_f, ps.ordb, type="samples", color="lifestage"); p2b 
```

Display levels of lifestages.  
```{r}
levels(as.factor(metadata_rare$lifestage))
#Egg: "#8C510A"
#Embryo: "#DFC27D"
#Larvae: "#80CDC1"
#Polyp Metamorphosed: "#003C30"
#Recruit Attached: "#BA55D3"
```

Output PCoA plot for visualization.  
```{r}
pcoa_viz<-p2b  + 
  geom_point(aes(colour=lifestage, shape=lifestage), size=4) + 
  ggtitle("PCoA")+
  
  scale_colour_manual(name="Lifestage", 
                      values = c("Attached Recruit (231 hpf)"="#BA55D3", "Embryo (38 hpf)"="#DFC27D", "Embryo (65 hpf)"="#DFC27D", "Larvae (93 hpf) "="#80CDC1", "Larvae (163 hpf)"="#80CDC1", "Larvae (231 hpf) "="#80CDC1", "Metamorphosed Recruit (231 hpf)"="#003C30", "Attached Recruit (183 hpf) "="#BA55D3"),
                      breaks = c("Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Recruit (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)"), 
                     labels = c( "Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)")) + 
  
  scale_shape_manual(name="Lifestage", 
                     values = c("Attached Recruit (231 hpf)"=13, "Embryo (38 hpf)"=3, "Embryo (65 hpf)"=4, "Larvae (93 hpf) "=5, "Larvae (163 hpf)"=6, "Larvae (231 hpf) "=8, "Metamorphosed Recruit (231 hpf)"=11, "Attached Recruit (183 hpf) "=12), 
                     breaks = c( "Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Recruit (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)"), 
                     labels = c( "Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)")) +
  
  theme_classic() + 
  theme(
    panel.background=element_blank()
  ); pcoa_viz

ggsave(plot=pcoa_viz, "Mcap2020/Figures/16S/pcoa_16S.jpeg", h=6, w=8)
```

Conduct a permanova in the vegan package.  

```{r}
samples <- as(sample_data(physeq1_rare_f), "data.frame")

perm1<-adonis2(phyloseq::distance(physeq1_rare_f, method="bray") ~ lifestage,
       data = samples)

perm1
```

Permanova suggests there is significant separation by lifestage in 16S profiles (p=0.001). 


## Plot Heatmap of relative abundance  

Filter out low frequency taxa <1% relative abundance.  
```{r}
physeq1_rare_f

physeq1_rare_prune = filter_taxa(physeq1_rare_f, function(x) mean(x) > 1, TRUE)

taxa_names(physeq1_rare_prune) 
```

There were 1364 taxa before filtering out taxa <1% and 16 taxa after filtering.  

Summarize by lifestage calculating means across replicates of each life stage.  
```{r}
mergedGP = merge_samples(physeq1_rare_prune, "lifestage")

#lifestage_order<-c("Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)")

lifestage_order<-c("Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Recruit (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)")
```

Replace OTU number with taxa names.  

View current data frame.  
```{r}
table<-as.matrix(t(otu_table(mergedGP)))
table<-as.matrix(as.data.frame(table))
colnames(table)
rownames(table)
```

Extract taxa names at the level of Family.  
```{r}
tax_table<-as.matrix(t(tax_table(mergedGP)))
tax_table<-as.matrix(t(as.data.frame(tax_table)))
colnames(tax_table)
rownames(tax_table)
tax_table<-as.data.frame(tax_table)
```
 
```{r}
#replace OTU with name from taxonomy table  
table_names<-as.data.frame(table)
table_names$OTU<-rownames(table)

table_names$Family<-tax_table$Family[match(table_names$OTU, tax_table$OTU)]
table_names$Genus<-tax_table$Genus[match(table_names$OTU, tax_table$OTU)]

#remove non alphabetic characters from Family and Genus columns 
table_names$Family <- gsub('[^[:alnum:] ]','',table_names$Family)

table_names$Taxa<-paste(table_names$OTU, "-", table_names$Family, ":", table_names$Genus)

#add to data frame 
rownames(table_names)<-table_names$Taxa

table_names<-table_names%>%
  select(!Family)%>%
  select(!Taxa)%>%
  select(!Genus)%>%
  select(!OTU)

table_names<-as.matrix(table_names)
```

Draw by lifestage with Complex Heatmap. 
```{r}
row_dend = dendsort(hclust(dist(table)))
col_dend = dendsort(hclust(dist(t(table))))

col_fun = colorRamp2(c(0, 50, 100), c("white", "darkgray", "black"))

pdf(file = "Mcap2020/Figures/16S/16S-ComplexHeatmap.pdf", height = 8, width = 18)
ComplexHeatmap::Heatmap(table_names, name = "Rel. Abun. (%)", row_title = "", column_title = "16S Relative Abundance (OTU - Family:Genus)", 
        col = col_fun, 
        row_names_side = "left", 
        row_dend_side = "right",
        width = unit(4, "in"), 
        height = unit(4, "in"), 
        column_dend_height = unit(.5, "in"),
        row_dend_width = unit(.5, "in"),
        column_dend_reorder = FALSE,
        row_dend_reorder = TRUE,
        row_gap = unit(2.5, "mm"),
        clustering_distance_rows = "euclidean",
        clustering_method_rows = "complete",
        border = TRUE, 
        column_names_gp =  gpar(fontsize = 12, border=FALSE),
        column_names_rot = 35,
        cluster_rows = row_dend, 
        #cluster_columns = col_dend,
        column_order = lifestage_order, 
        row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE))
dev.off()

```

Plot a heatmap.  

```{r}
plot_heatmap(physeq1_rare, "NMDS", "bray", "lifestage", "Family", low="#66CCFF", high="#000033", na.value="white")
```  








# **NMDS of rarefied dataset**  

```{r}
GP1 = physeq1_rare
GP1
```

Transform to even sampling depth.  
```{r}
#GP1 = transform_sample_counts(GP1, function(x) 1E6 * x/sum(x))
```

View ordinations: PCoA + unifrac.    
```{r}
GP.ord <- ordinate(GP1, "PCoA", "unifrac")
p1 = plot_ordination(GP1, GP.ord, type="taxa", color="Class", title="taxa")
```

View faceted by Phylum.  
```{r}
p1 + facet_wrap(~Class, 3) + theme(legend.position="none")
```

Plot with samples now.  
```{r}
p2 = plot_ordination(GP1, GP.ord, type="samples", color="lifestage") 
p2<-p2 + geom_polygon(aes(fill=lifestage)) + geom_point(size=5) + ggtitle("subsampled dataset PCoA Unifrac")
#p2<-p2 + geom_point(size=5) + ggtitle("subsampled dataset")
p2

ggsave(plot=p2, filename="Mcap2020/Figures/16S/pcoa_unifrac_rare.png")
```


View ordinations: PCoA + bray curtis    
```{r}
GP.ord <- ordinate(GP1, "PCoA", "bray")
p1 = plot_ordination(GP1, GP.ord, type="taxa", color="Class", title="taxa")
```

View faceted by Phylum.  
```{r}
p1 + facet_wrap(~Class, 3) + theme(legend.position="none")
```

Plot with samples now.  
```{r}
p2 = plot_ordination(GP1, GP.ord, type="samples", color="lifestage") 
p2<-p2 + geom_polygon(aes(fill=lifestage)) + geom_point(size=5) + ggtitle("subsampled dataset PCoA Bray Curtis")
#p2<-p2 + geom_point(size=5) + ggtitle("subsampled dataset")
p2

ggsave(plot=p2, filename="Mcap2020/Figures/16S/pcoa_bray_rare.png")
```


View ordinations: NMDS + bray curtis    
```{r}
GP.ord <- ordinate(GP1, "NMDS", "bray")
p1 = plot_ordination(GP1, GP.ord, type="taxa", color="Class", title="taxa")
```

View faceted by Phylum.  
```{r}
p1 + facet_wrap(~Class, 3) + theme(legend.position="none")
```

Plot with samples now.  
```{r}
p2 = plot_ordination(GP1, GP.ord, type="samples", color="lifestage") 
p2<-p2 + geom_polygon(aes(fill=lifestage)) + geom_point(size=5) + ggtitle("subsampled dataset NMDS Bray Curtis")
#p2<-p2 + geom_point(size=5) + ggtitle("subsampled dataset")
p2

ggsave(plot=p2, filename="Mcap2020/Figures/16S/NMDS_bray_rare.png")
```


View ordinations: NMDS + unifrac    
```{r}
GP.ord <- ordinate(GP1, "NMDS", "unifrac")
p1 = plot_ordination(GP1, GP.ord, type="taxa", color="Class", title="taxa")
```

View faceted by Phylum.  
```{r}
p1 + facet_wrap(~Class, 3) + theme(legend.position="none")
```

Plot with samples now.  
```{r}
p2 = plot_ordination(GP1, GP.ord, type="samples", color="lifestage") 
p2<-p2 + geom_polygon(aes(fill=lifestage)) + geom_point(size=5) + ggtitle("subsampled dataset NMDS unifrac")
#p2<-p2 + geom_point(size=5) + ggtitle("subsampled dataset")
p2

ggsave(plot=p2, filename="Mcap2020/Figures/16S/NMDS_unifrac_rare.png")
```


We do see separation between lifestages.    

Conduct a permanova in the vegan package.  

```{r}

metadata <- as(sample_data(GP1), "data.frame")

adonis(phyloseq::distance(GP1, method="bray") ~ lifestage,
       data = metadata)

adonis(phyloseq::distance(GP1, method="unifrac") ~ lifestage,
       data = metadata)
```

There is significant difference in the centroids between life stages in the rarefied dataset for both unifrac and bray-curtis distance measures.  

# **NMDS of full dataset**  

```{r}
GP1f = physeq1
GP1f
```

Transform to even sampling depth.  
```{r}
GP1f = transform_sample_counts(GP1f, function(x) 1E6 * x/sum(x))
```

View ordination by OTU.  
```{r}
GP.ord <- ordinate(GP1f, "NMDS", "bray")
p1 = plot_ordination(GP1f, GP.ord, type="taxa", color="Class", title="taxa")
```

View faceted by Phylum.  
```{r}
p1 + facet_wrap(~Class, 3) + theme(legend.position="none")
```

Plot with samples now.  
```{r}
p2 = plot_ordination(GP1f, GP.ord, type="samples", color="lifestage") 
p2<- p2 + geom_polygon(aes(fill=lifestage)) + geom_point(size=5) + ggtitle("full dataset")
p2
```

Conduct a permanova in the vegan package.  

```{r}
metadata <- as(sample_data(GP1f), "data.frame")

adonis(phyloseq::distance(GP1f, method="bray") ~ lifestage,
       data = metadata)
```

There are also significant differences between centroids in life stages in the full dataset.  

# **Alpha Diversity - Full dataset**  

Remove any OTU that has a count of 0 in all samples. (There are none in this dataset).    
```{r}
GP2f <- prune_taxa(taxa_sums(physeq1) > 0, physeq1)
```

Plot all richness metrics.  
```{r}
plot_richness(GP2f)
```

Plot the two most commonly used - Shannon and Inverse Simpson.  
```{r}
plot_richness(GP2f, measures=c("InvSimpson", "Shannon"), x="lifestage", color="group")
```

We see increase in alpha diversity across development with higher values in attached recruits.    

Test for differences in alpha diversity.  
```{r}
metadata <- as(sample_data(GP2f), "data.frame")
metadata$sample<-rownames(metadata)

alpha_full<-estimate_richness(GP2f, measures=c("InvSimpson", "Shannon"))
alpha_full$sample<-rownames(alpha_full)

alpha_full<-full_join(metadata, alpha_full, by="sample")

summary(aov(Shannon~lifestage, data=alpha_full))
summary(aov(InvSimpson~lifestage, data=alpha_full))
```

There are significant differences in alpha diversity between lifestages.  

# **Alpha Diversity - Rare dataset**  

```{r}
GP2 <- prune_taxa(taxa_sums(physeq1_rare) > 0, physeq1_rare)
```

Plot all richness metrics.  
```{r}
plot_richness(GP2)
```

Plot the two most commonly used - Shannon and Inverse Simpson.  
```{r}
alpha1<-plot_richness(GP2, measures=c("InvSimpson", "Shannon", "Chao1"), x="lifestage", color="group")
alpha1

ggsave(plot=alpha1, file="Mcap2020/Figures/16S/alpha_rare.png")
```

Similar to the full dataset, we see increases in alpha diversity across development with higher values in attached recruits.  

Test for differences in alpha diversity.  
```{r}
metadata <- as(sample_data(GP2), "data.frame")
metadata$sample<-rownames(metadata)

alpha_rare<-estimate_richness(GP2, measures=c("InvSimpson", "Shannon", "Chao1"))
alpha_rare$sample<-rownames(alpha_rare)

alpha_rare<-full_join(metadata, alpha_rare, by="sample")

summary(aov(Shannon~lifestage, data=alpha_rare))
summary(aov(InvSimpson~lifestage, data=alpha_rare))
summary(aov(Chao1~lifestage, data=alpha_rare))
```

There are significant differences in alpha diversity between lifestages for all metrics.    

# **Networks - Full dataset**  

Remove NA's. 
```{r}
GP3f = physeq1
```

Plot network.  
```{r}
plot_net(GP3f, maxdist = 1, point_label = "lifestage", color = "group")
```

Plot network.  
```{r}
igF <- make_network(GP3f, max.dist=1)
plot_network(igF, GP3f, color="group", label=NULL)
```

The networks are not particularly informative for this dataset.  

# **Networks - Rare dataset**  

Remove NA's. 
```{r}
GP3 = physeq1_rare
```

Plot network.  
```{r}
plot_net(GP3, maxdist = 1, point_label = "lifestage", color = "group")
```

Plot network.  
```{r}
ig <- make_network(GP3, max.dist=1)
plot_network(ig, GP3, color="group", label=NULL)
```

The networks are not particularly informative for this dataset.  \

# **Plot unknowns - Full dataset** 

Plot data at the level of phylum. Bacteria_unclassified are our unknowns.      
```{r}
plot_bar(physeq1, fill="Phylum", facet_grid=~Phylum)

p5<-plot_bar(physeq1, fill="Order", facet_grid=~Order)

pdf(file="Mcap2020/Figures/16S/order_abund.pdf", height = 8, width = 100)
p5 + theme(legend.position="none")
dev.off()
```

We have some unclassified bacteria, but a higher proportion of proteobacteria and bacteriodes.  

# **Plot unknowns - rare dataset** 

Plot data at the level of phylum in the rarefied dataset. Bacteria_unclassified are our unknowns.      
```{r}
plot_bar(physeq1_rare, fill="Phylum", facet_grid=~Phylum)
```

We have some unclassified bacteria, but a higher proportion of proteobacteria and bacteriodes.  


# **DESeq2 - Full dataset**  

THERE IS SOMETHING GOING ON WITH THE PVALUE ADJUSTMENT HERE - all the values are close to 1 - high number of comparisons? But there must be significant drivers? 

Convert phyloseq object to DESeq2 object.  

Construct DESeq2 object looking at differences by lifestage.  
```{r}
diagdds_full = phyloseq_to_deseq2(physeq1, ~ lifestage)

gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}
geoMeans = apply(counts(diagdds_full), 1, gm_mean)
diagdds_full = estimateSizeFactors(diagdds_full, geoMeans = geoMeans)
#diagdds_full = DESeq(diagdds_full, fitType="local")

diagdds_full = DESeq(diagdds_full, test="Wald", fitType="parametric")
```

Look at results table.  
```{r}
res_full = results(diagdds_full, cooksCutoff = FALSE)
alpha = 0.05
sigtab = res_full[which(res_full$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(physeq1)[rownames(sigtab), ], "matrix"))
head(sigtab)
length(sigtab)
```


# **DESeq2 - Rare dataset**  

Convert phyloseq object to DESeq2 object.  

Construct DESeq2 object looking at differences by lifestage.  
```{r}
diagdds_rare = phyloseq_to_deseq2(physeq1_rare, ~ group)

gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}
geoMeans = apply(counts(diagdds_rare), 1, gm_mean)
diagdds_rare = estimateSizeFactors(diagdds_rare, geoMeans = geoMeans)
#diagdds_rare = DESeq(diagdds_rare, fitType="local")

diagdds_rare = DESeq(diagdds_rare, test="Wald", fitType="parametric")
```

Look at results table.  
```{r}
res_rare = results(diagdds_rare, cooksCutoff = FALSE)
alpha = 0.05
sigtab = res_rare[which(res_rare$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(physeq1_rare)[rownames(sigtab), ], "matrix"))
head(sigtab)
```

We do not get any members that are differential - why is this when we have group separation?   