---
title: "Analysis of 16S data output from mothur bioniformatic pipeline"
author: "Ariana S Huffmyer"
date: '2022'
output:
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("janitor" %in% rownames(installed.packages()) == 'FALSE') install.packages('janitor') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("phyloseq" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install('phyloseq') 
if ("stringr" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install('stringr')
if ("ape" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install('ape') 
if ("DESeq2" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install('DESeq2') 
if ("microViz" %in% rownames(installed.packages()) == 'FALSE') devtools::install_github("david-barnett/microViz@0.9.0") 
if ("vegan" %in% rownames(installed.packages()) == 'FALSE') install.packages('vegan') 

#load packages
library("ggplot2")
library('cowplot')
library("tidyverse")
library("janitor")
library("car")
library("phyloseq")
library("stringr")
library("ape")
library("DESeq2")
library("microViz")
library("vegan")
library(dendsort)
library(WGCNA)
library(circlize)
library(ComplexHeatmap)
```

# **Load data**  

Load data for use in analysis with PhyloSeq. Here we will import an OTU table and a taxonomy table.  

We will load "rare" files as those that were subsampled/rarefied.  

```{r}
otu.rare<-read.table("Mcap2020/Output/16S/mothur/TextFormat/mcap.opti_mcc.0.03.subsample.shared.txt", sep='\t', header=TRUE)

taxonomy<-read.table("Mcap2020/Output/16S/mothur/TextFormat/mcap.taxonomy.txt", sep='\t', header=TRUE)

rarefaction<-read.table("Mcap2020/Output/16S/mothur/TextFormat/mcap.opti_mcc.groups.rarefaction.txt", sep='\t', header=TRUE)
```

Load sample information.  
```{r}
metadata<-read.table("Mcap2020/Data/16S/mcap_metadata.txt", sep='\t', header=TRUE)

rownames(metadata)<-metadata$sample

metadata<-metadata%>%
  select(!sample)%>%
  mutate(lifestage=if_else(lifestage=="Metamorphosed Recruit (183 hpf)", "Metamorphosed Polyp (183 hpf)", lifestage))%>%
  mutate(lifestage=if_else(lifestage=="Metamorphosed Recruit (231 hpf)", "Metamorphosed Polyp (231 hpf)", lifestage))

#Replace metamorphosed "recruit" with "polyp"  

  
```

Set default plot theme.  

```{r}
theme_set(theme_bw())
```

Format columns of our taxonomy file. We need to separate by ; and remove the quotes. We also need to remove the confidence values.  
```{r}
head(taxonomy)

taxonomy<-taxonomy%>%
  mutate(Taxonomy=str_replace_all(string=Taxonomy, pattern="\\(\\d*\\)", replacement="")) %>% #remove parentheses with numbers
	mutate(Taxonomy=str_replace_all(string=Taxonomy, pattern=";$", replacement="")) %>% #remove semicolons
 # mutate(Taxonomy=str_replace_all(string=Taxonomy, pattern="*", replacement="")) %>% #remove quotes
  #mutate(Taxonomy=str_replace_all(string=Taxonomy, pattern="""$", replacement="")) %>% #remove quotes
	separate(Taxonomy, into=c("Domain", "Phylum", "Class", "Order", "Family", "Genus"), sep=";") #separate columns

head(taxonomy)

rownames(taxonomy)<-taxonomy$OTU
taxonomy<-as.matrix(taxonomy)

```

# **Heatmaps and PCoA of rarefied dataset**   

## Prepare data  

Prepare the matrix to be all numeric with taxa in rows.  
```{r}
otu.rare<-otu.rare%>%
  select(!label)%>%
  select(!numOtus)

rownames(otu.rare)<-otu.rare$Group

otu.rare<-otu.rare%>%
  select(!Group)

otu.rare<-t(otu.rare)
```

Combine the OTU and taxonomic information.  
```{r}
OTU_rare = otu_table(otu.rare, taxa_are_rows = TRUE)
TAX = tax_table(taxonomy)
OTU_rare
```

Load into phyloseq.  
```{r}
physeq_rare = phyloseq(OTU_rare, TAX)
physeq_rare
```

Plot an abundance table at the different taxonomic levels.     
```{r}
plot_bar(physeq_rare, fill = "Domain")+theme(legend.position="bottom")
plot_bar(physeq_rare, fill = "Phylum")+theme(legend.position="bottom")
plot_bar(physeq_rare, fill = "Class")+theme(legend.position="bottom")
#plot_bar(physeq_rare, fill = "Order")+theme(legend.position="bottom")
#plot_bar(physeq_rare, fill = "Family")+theme(legend.position="bottom")
#plot_bar(physeq_rare, fill = "Genus")+theme(legend.position="bottom")

```

Plot a tree of OTU's.  
```{r}
random_tree_rare = rtree(ntaxa(physeq_rare), rooted=TRUE, tip.label=taxa_names(physeq_rare))
plot(random_tree_rare)
```

Add metdata 
```{r}
list<-sample_names(physeq_rare)

metadata_rare<-read.table("Mcap2020/Data/16S/mcap_metadata.txt", sep='\t', header=TRUE)

rownames(metadata_rare)<-metadata_rare$sample

metadata_rare<-metadata_rare%>%
  filter(sample %in% list)

metadata_rare<-metadata_rare%>%
  select(!sample)%>%
  mutate(lifestage=if_else(lifestage=="Metamorphosed Recruit (183 hpf)", "Metamorphosed Polyp (183 hpf)", lifestage))%>%
  mutate(lifestage=if_else(lifestage=="Metamorphosed Recruit (231 hpf)", "Metamorphosed Polyp (231 hpf)", lifestage))
```

Add sample data 
```{r}
sampledata = sample_data(data.frame(
  lifestage = metadata_rare$lifestage,
  group = metadata_rare$group, 
  hpf = metadata_rare$hpf,
  row.names=sample_names(physeq_rare),
  stringsAsFactors=FALSE
))
sampledata
```

Merge the tree, sample information, and OTU table.  
```{r}
physeq1_rare = merge_phyloseq(physeq_rare, sampledata, random_tree_rare)
physeq1_rare
```

Now rebuild phyloseq.  
```{r}
physeq2_rare = phyloseq(OTU_rare, TAX, sampledata, random_tree_rare)
physeq2_rare
```

Check that they are identical. Should return TRUE.  
```{r}
identical(physeq1_rare, physeq2_rare)
```

Filter out samples with low numbers in groups from phyloseq object.   
```{r}
physeq1_rare<-ps_filter(physeq1_rare, lifestage != "Metamorphosed Polyp (183 hpf)")
physeq1_rare<-ps_filter(physeq1_rare, lifestage != "Egg (1 hpf)")
physeq1_rare<-ps_filter(physeq1_rare, lifestage != "Embryo (5 hpf)")
#physeq1_rare<-ps_filter(physeq1_rare, lifestage != "Embryo (38 hpf)")
```

Now build a tree with the combined data.  
```{r}
plot_tree(physeq1_rare, color="lifestage", label.tips="taxa_names", ladderize="left", plot.margin=0.3)
```

Plot relative abundance now with group information.  

```{r}
bar1<-plot_bar(physeq1_rare, fill = "Class")+theme(legend.position="bottom")+facet_grid(~lifestage, scales="free_x"); bar1
#ggsave(plot=bar1, file="Mcap2020/Figures/16S/relabund_class_rare.png", h=6, w=18)

bar2<-plot_bar(physeq1_rare, fill = "Phylum")+theme(legend.position="bottom")+facet_grid(~lifestage, scales="free_x"); bar2
#ggsave(plot=bar2, file="Mcap2020/Figures/16S/relabund_phylum_rare.png", h=6, w=18)

bar3<-plot_bar(physeq1_rare, fill = "Order")+theme(legend.position="bottom")+facet_grid(~lifestage, scales="free_x"); bar3
#ggsave(plot=bar3, file="Mcap2020/Figures/16S/relabund_order_rare.png", h=9, w=18)

bar4<-plot_bar(physeq1_rare, fill = "Family")+theme(legend.position="bottom")+facet_grid(~lifestage, scales="free_x"); bar4
#ggsave(plot=bar4, file="Mcap2020/Figures/16S/relabund_family_rare.png", h=12, w=18)

bar5<-plot_bar(physeq1_rare, fill = "Genus")+theme(legend.position="bottom")+facet_grid(~lifestage, scales="free_x"); bar5
#ggsave(plot=bar5, file="Mcap2020/Figures/16S/relabund_genus_rare.png", h=25, w=18)
```

Plot relative abundance at the phylum level.  
```{r}
dat <- physeq1_rare %>% tax_glom(taxrank = "Phylum") %>% psmelt()
head(dat)
```

```{r}
df<-dat %>% 
  group_by(lifestage, Phylum) %>% 
  summarise(TotalAbundance=sum(Abundance))%>% #calculate total abundance of each phylum
  group_by(lifestage) %>%
  mutate(PercRelAbundance=TotalAbundance/sum(TotalAbundance)*100)%>% #calculate relative abundance
  select(!TotalAbundance)

df$Phylum <- gsub('[^[:alnum:] ]','',df$Phylum)

#change to a matrix with phylum in rows and lifestage in columns
df<-df%>%
  spread(key=Phylum, value=PercRelAbundance)

#replace na with 0, since 0 abundance was calculated as na in step above b/c dividing by 0
df[is.na(df)] <- 0

rownames(df)<-df$lifestage
lifestage_list<-df$lifestage

dim(df)

df<-as.matrix(df)

df<-df[,-1]

taxa_list<-colnames(df)

df_mat <- matrix(as.numeric(df),    # Convert to numeric matrix
                  ncol = ncol(df))                                   # Print numeric matrix

colnames(df_mat)<-taxa_list
rownames(df_mat)<-lifestage_list

df_mat<-t(df_mat)
```

Plot heatmap of phylum relative abundance.  
```{r}
row_dend = dendsort(hclust(dist(df_mat)))
col_dend = dendsort(hclust(dist(t(df_mat))))

col_fun = colorRamp2(c(0, 50, 100), c("white", "darkgray", "black"))

pdf(file = "Mcap2020/Figures/16S/16S-Phylum-Heatmap.pdf", height = 8, width = 18)
ComplexHeatmap::Heatmap(df_mat, name = "Rel. Abun. (%)", row_title = "", column_title = "16S Relative Abundance (Phylum)", 
        col = col_fun, 
        row_names_side = "left", 
        row_dend_side = "right",
        width = unit(4, "in"), 
        height = unit(4, "in"), 
        column_dend_height = unit(.5, "in"),
        row_dend_width = unit(.5, "in"),
        column_dend_reorder = FALSE,
        row_dend_reorder = TRUE,
        row_gap = unit(2.5, "mm"),
        clustering_distance_rows = "euclidean",
        clustering_method_rows = "complete",
        border = TRUE, 
        column_names_gp =  gpar(fontsize = 12, border=FALSE),
        column_names_rot = 35,
        cluster_rows = row_dend, 
        #cluster_columns = col_dend,
        column_order = lifestage_order, 
        row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE))
dev.off()

```

Display as % abundance and filter out taxa <1% abundance.    

```{r}
physeq1_rare_f = transform_sample_counts(physeq1_rare, function(x) 100 * x/sum(x))
```

## Plot PCoA and conduct PERMANOVA  

Prepare PCoA ordination.  
```{r}
ps.ordb <- ordinate(physeq1_rare, "PCoA", "bray")
p2b = plot_ordination(physeq1_rare, ps.ordb, type="samples", color="lifestage"); p2b 
```

Display levels of lifestages.  
```{r}
levels(as.factor(metadata_rare$lifestage))
#Egg: "#8C510A"
#Embryo: "#DFC27D"
#Larvae: "#80CDC1"
#Polyp Metamorphosed: "#003C30"
#Recruit Attached: "#BA55D3"
```

Output PCoA plot for visualization.  
```{r}
pcoa_viz<-p2b  + 
  geom_point(aes(colour=lifestage, shape=lifestage), size=4) + 
  ggtitle("PCoA")+
  
  scale_colour_manual(name="Lifestage", 
                      values = c("Attached Recruit (231 hpf)"="#BA55D3", "Embryo (38 hpf)"="#DFC27D", "Embryo (65 hpf)"="#DFC27D", "Larvae (93 hpf) "="#80CDC1", "Larvae (163 hpf)"="#80CDC1", "Larvae (231 hpf) "="#80CDC1", "Metamorphosed Polyp (231 hpf)"="#003C30", "Attached Recruit (183 hpf) "="#BA55D3"),
                      breaks = c("Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)"), 
                     labels = c( "Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)")) + 
  
  scale_shape_manual(name="Lifestage", 
                     values = c("Attached Recruit (231 hpf)"=13, "Embryo (38 hpf)"=3, "Embryo (65 hpf)"=4, "Larvae (93 hpf) "=5, "Larvae (163 hpf)"=6, "Larvae (231 hpf) "=8, "Metamorphosed Polyp (231 hpf)"=11, "Attached Recruit (183 hpf) "=12), 
                     breaks = c( "Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)"), 
                     labels = c( "Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)")) +
  
  theme_classic() + 
  theme(
    panel.background=element_blank()
  ); pcoa_viz

ggsave(plot=pcoa_viz, "Mcap2020/Figures/16S/pcoa_16S.jpeg", h=6, w=8)
```

Conduct a permanova in the vegan package.  

```{r}
samples <- as(sample_data(physeq1_rare), "data.frame")

perm1<-adonis2(phyloseq::distance(physeq1_rare, method="bray") ~ lifestage,
       data = samples)

perm1
```

Permanova suggests there is significant separation by lifestage in 16S profiles (p=0.001). 

## Plot Heatmap of relative abundance at the family/genus level   

Filter out low frequency taxa <1% relative abundance.  
```{r}
physeq1_rare_f

physeq1_rare_prune = filter_taxa(physeq1_rare_f, function(x) mean(x) > 1, TRUE)

taxa_names(physeq1_rare_prune) 
```

There were 1364 taxa before filtering out taxa <1% and 16 taxa after filtering.  

Summarize by lifestage calculating means across replicates of each life stage.  
```{r}
mergedGP = merge_samples(physeq1_rare_prune, "lifestage")

lifestage_order<-c("Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)")
```

Replace OTU number with taxa names.  

View current data frame.  
```{r}
table<-as.matrix(t(otu_table(mergedGP)))
table<-as.matrix(as.data.frame(table))
colnames(table)
rownames(table)
```

Extract taxa names at the level of Family.  
```{r}
tax_table<-as.matrix(t(tax_table(mergedGP)))
tax_table<-as.matrix(t(as.data.frame(tax_table)))
colnames(tax_table)
rownames(tax_table)
tax_table<-as.data.frame(tax_table)
```
 
```{r}
#replace OTU with name from taxonomy table  
table_names<-as.data.frame(table)
table_names$OTU<-rownames(table)

table_names$Family<-tax_table$Family[match(table_names$OTU, tax_table$OTU)]
table_names$Genus<-tax_table$Genus[match(table_names$OTU, tax_table$OTU)]

#remove non alphabetic characters from Family and Genus columns 
table_names$Family <- gsub('[^[:alnum:] ]','',table_names$Family)

table_names$Taxa<-paste(table_names$OTU, "-", table_names$Family, ":", table_names$Genus)

#add to data frame 
rownames(table_names)<-table_names$Taxa

table_names<-table_names%>%
  select(!Family)%>%
  select(!Taxa)%>%
  select(!Genus)%>%
  select(!OTU)

table_names<-as.matrix(table_names)
```

Draw by lifestage with Complex Heatmap. 
```{r}
row_dend = dendsort(hclust(dist(table)))
col_dend = dendsort(hclust(dist(t(table))))

col_fun = colorRamp2(c(0, 50, 100), c("white", "darkgray", "black"))

pdf(file = "Mcap2020/Figures/16S/16S-ComplexHeatmap.pdf", height = 8, width = 18)
ComplexHeatmap::Heatmap(table_names, name = "Rel. Abun. (%)", row_title = "", column_title = "16S Relative Abundance (OTU - Family:Genus) Taxa >1%", 
        col = col_fun, 
        row_names_side = "left", 
        row_dend_side = "right",
        width = unit(4, "in"), 
        height = unit(4, "in"), 
        column_dend_height = unit(.5, "in"),
        row_dend_width = unit(.5, "in"),
        column_dend_reorder = FALSE,
        row_dend_reorder = TRUE,
        row_gap = unit(2.5, "mm"),
        clustering_distance_rows = "euclidean",
        clustering_method_rows = "complete",
        border = TRUE, 
        column_names_gp =  gpar(fontsize = 12, border=FALSE),
        column_names_rot = 35,
        cluster_rows = row_dend, 
        #cluster_columns = col_dend,
        column_order = lifestage_order, 
        row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE))
dev.off()

```


# **NMDS of rarefied dataset**  

```{r}
GP1 = physeq1_rare
GP1
```

View ordinations: PCoA + unifrac.    
```{r}
GP.ord <- ordinate(GP1, "PCoA", "unifrac")
p1 = plot_ordination(GP1, GP.ord, type="taxa", color="Class", title="taxa")
```

View faceted by Phylum.  
```{r}
p1 + facet_wrap(~Class, 3) + theme(legend.position="none")
```

Plot with samples now.  
```{r}
p2 = plot_ordination(GP1, GP.ord, type="samples", color="lifestage") 
p2<-p2 + geom_polygon(aes(fill=lifestage)) + geom_point(size=5) + ggtitle("subsampled dataset PCoA Unifrac")
#p2<-p2 + geom_point(size=5) + ggtitle("subsampled dataset")
p2

#ggsave(plot=p2, filename="Mcap2020/Figures/16S/pcoa_unifrac_rare.png")
```


View ordinations: PCoA + bray curtis    
```{r}
GP.ord <- ordinate(GP1, "PCoA", "bray")
p1 = plot_ordination(GP1, GP.ord, type="taxa", color="Class", title="taxa")
```

View faceted by Phylum.  
```{r}
p1 + facet_wrap(~Class, 3) + theme(legend.position="none")
```

Plot with samples now.  
```{r}
p2 = plot_ordination(GP1, GP.ord, type="samples", color="lifestage") 
p2<-p2 + geom_polygon(aes(fill=lifestage)) + geom_point(size=5) + ggtitle("subsampled dataset PCoA Bray Curtis")
#p2<-p2 + geom_point(size=5) + ggtitle("subsampled dataset")
p2

#ggsave(plot=p2, filename="Mcap2020/Figures/16S/pcoa_bray_rare.png")
```


View ordinations: NMDS + bray curtis    
```{r}
GP.ord <- ordinate(GP1, "NMDS", "bray")
p1 = plot_ordination(GP1, GP.ord, type="taxa", color="Class", title="taxa")
```

View faceted by Phylum.  
```{r}
p1 + facet_wrap(~Class, 3) + theme(legend.position="none")
```

Plot with samples now.  
```{r}
p2 = plot_ordination(GP1, GP.ord, type="samples", color="lifestage") 
p2<-p2 + geom_polygon(aes(fill=lifestage)) + geom_point(size=5) + ggtitle("subsampled dataset NMDS Bray Curtis")
#p2<-p2 + geom_point(size=5) + ggtitle("subsampled dataset")
p2

ggsave(plot=p2, filename="Mcap2020/Figures/16S/NMDS_bray_rare.png")
```


View ordinations: NMDS + unifrac    
```{r}
GP.ord <- ordinate(GP1, "NMDS", "unifrac")
p1 = plot_ordination(GP1, GP.ord, type="taxa", color="Class", title="taxa")
```

View faceted by Phylum.  
```{r}
p1 + facet_wrap(~Class, 3) + theme(legend.position="none")
```

Plot with samples now.  
```{r}
p2 = plot_ordination(GP1, GP.ord, type="samples", color="lifestage") 
p2<-p2 + geom_polygon(aes(fill=lifestage)) + geom_point(size=5) + ggtitle("subsampled dataset NMDS unifrac")
#p2<-p2 + geom_point(size=5) + ggtitle("subsampled dataset")
p2

#ggsave(plot=p2, filename="Mcap2020/Figures/16S/NMDS_unifrac_rare.png")
```

Conduct a permanova in the vegan package.  

```{r}

metadata <- as(sample_data(GP1), "data.frame")

adonis2(phyloseq::distance(GP1, method="bray") ~ lifestage,
       data = metadata)

adonis2(phyloseq::distance(GP1, method="unifrac") ~ lifestage,
       data = metadata)
```

Separation by lifestage p=0.001, as in PCoA from above.  

# **Alpha Diversity - Rare dataset**  

```{r}
GP2 <- prune_taxa(taxa_sums(physeq1_rare) > 0, physeq1_rare)
```

Plot all richness metrics.  
```{r}
plot_richness(GP2)
```

```{r}

```

Plot the two most commonly used - Shannon and Inverse Simpson.  
```{r}
alpha1<-plot_richness(GP2, measures=c("InvSimpson", "Shannon"), x="lifestage", color="lifestage", nrow=2)

#reorder factors 
alpha1$data$lifestage <- as.character(alpha1$data$lifestage)
alpha1$data$lifestage <- factor(alpha1$data$lifestage, levels=lifestage_order)

alpha1 <- alpha1 + geom_point(size=3) + geom_boxplot() + theme_classic() + scale_color_manual() + scale_color_manual(name="Lifestage", 
                      values = c("Attached Recruit (231 hpf)"="#BA55D3", "Embryo (38 hpf)"="#DFC27D", "Embryo (65 hpf)"="#DFC27D", "Larvae (93 hpf) "="#80CDC1", "Larvae (163 hpf)"="#80CDC1", "Larvae (231 hpf) "="#80CDC1", "Metamorphosed Polyp (231 hpf)"="#003C30", "Attached Recruit (183 hpf) "="#BA55D3"),
                      breaks = c("Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)"), 
                     labels = c( "Embryo (38 hpf)", "Embryo (65 hpf)", "Larvae (93 hpf) ", "Larvae (163 hpf)", "Larvae (231 hpf) ", "Metamorphosed Polyp (231 hpf)", "Attached Recruit (183 hpf) ", "Attached Recruit (231 hpf)")) + 
  theme(axis.text.x=element_text(angle=45, hjust=1, color="black")) + 
  theme(axis.title.x=element_blank())+
  theme(legend.position="none");alpha1

ggsave(plot=alpha1, file="Mcap2020/Figures/16S/alpha_rare.png", width=4, height=7)
```

Similar to the full dataset, we see increases in alpha diversity across development with higher values in attached recruits.  

Test for differences in alpha diversity.  
```{r}
metadata <- as(sample_data(GP2), "data.frame")
metadata$sample<-rownames(metadata)

alpha_rare<-estimate_richness(GP2, measures=c("InvSimpson", "Shannon", "Chao1"))
alpha_rare$sample<-rownames(alpha_rare)

alpha_rare<-full_join(metadata, alpha_rare, by="sample")

summary(aov(Shannon~lifestage, data=alpha_rare))
summary(aov(InvSimpson~lifestage, data=alpha_rare))
summary(aov(Chao1~lifestage, data=alpha_rare))
```

There are significant differences in alpha diversity between lifestages for all metrics.    


# **Plot unknowns - rare dataset** 

Plot data at the level of phylum in the rarefied dataset. Bacteria_unclassified are our unknowns.      
```{r}
plot_bar(physeq1_rare, fill="Phylum", facet_grid=~Phylum)
```

We have some unclassified bacteria, but a higher proportion of proteobacteria and bacteriodes.  

# **DESeq2 - Rare dataset**  

Convert phyloseq object to DESeq2 object.  

Construct DESeq2 object looking at differences by lifestage.  
```{r}
diagdds_rare = phyloseq_to_deseq2(physeq1_rare, ~ lifestage)

gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}
geoMeans = apply(counts(diagdds_rare), 1, gm_mean)
diagdds_rare = estimateSizeFactors(diagdds_rare, geoMeans = geoMeans)
#diagdds_rare = DESeq(diagdds_rare, fitType="local")

diagdds_rare = DESeq(diagdds_rare, test="Wald", fitType="parametric")
```

Look at results table.  
```{r}
res_rare = results(diagdds_rare, cooksCutoff = FALSE)
alpha = 0.05
sigtab = res_rare[which(res_rare$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(physeq1_rare)[rownames(sigtab), ], "matrix"))
head(sigtab)
```

DESeq2 analysis is not particularly informative for this dataset, no differential members are identified.  


