---
title: Early life history respirometry (photosynthesis & respiration) with LoLinR
author: "AS Huffmyer"
date: '2020'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
---
## Setup  

Set up workspace, set options, and load required packages.    
```{r, echo=TRUE, show=FALSE}
rm(list=ls(all=TRUE)) 
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("devtools" %in% rownames(installed.packages()) == 'FALSE') install.packages('devtools') 
library(devtools)
if ("segmented" %in% rownames(installed.packages()) == 'FALSE') install.packages('segmented') 
if ("plotrix" %in% rownames(installed.packages()) == 'FALSE') install.packages('plotrix') 
if ("gridExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('gridExtra') 
if ("LoLinR" %in% rownames(installed.packages()) == 'FALSE') install_github('colin-olito/LoLinR') 
if ("lubridate" %in% rownames(installed.packages()) == 'FALSE') install.packages('lubridate') 
if ("chron" %in% rownames(installed.packages()) == 'FALSE') install.packages('chron') 
if ("plyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('plyr') 
if ("dplyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('dplyr') 
if ("stringr" %in% rownames(installed.packages()) == 'FALSE') install.packages('stringr') 
if ("Rmisc" %in% rownames(installed.packages()) == 'FALSE') install.packages('Rmisc') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 

#load packages
library("ggplot2")
library("segmented")
library("plotrix")
library("gridExtra")
library("LoLinR")
library("lubridate")
library("chron")
library('plyr')
library('dplyr')
library('stringr')
library('Rmisc')
library('car')
library('lme4')
library('lmerTest')
```


## Respirometry in LoLinR

Modified from K. Wong, S. Gurr, H. Putnam scripts  

Set the path of all respirometry files within the R project.   

```{r, warning=FALSE, message=FALSE}
path.p<-"../Data/Respiration/Mcap2020/Runs" #location of files
```

Bring in the file names.
```{r, warning=FALSE, message=FALSE}
# bring in the respiration file names
file.names<-basename(list.files(path = path.p, pattern = "csv$", recursive = TRUE)) #list all csv file names in the folder and subfolders
```

Generate respiration data frames.

```{r, warning=FALSE, message=FALSE}
#generate a 6 column dataframe with specific column names
#respiration
Resp.R <- data.frame(matrix(NA, ncol=6))
colnames(Resp.R) <- c("Date", "Run", "Sample.ID","Chamber.ID","Intercept", "umol.L.sec")

Resp.Rb <- data.frame(matrix(NA, ncol=6))
colnames(Resp.Rb) <- c("Date", "Run", "Sample.ID","Chamber.ID","Intercept", "umol.L.sec")

#photosynthesis 
Photo.R <- data.frame(matrix(NA, ncol=6))
colnames(Photo.R) <- c("Date", "Run", "Sample.ID","Chamber.ID","Intercept", "umol.L.sec")

Photo.Rb <- data.frame(matrix(NA, ncol=6))
colnames(Photo.Rb) <- c("Date", "Run", "Sample.ID","Chamber.ID","Intercept", "umol.L.sec")
```

Load in the sample information file. It is important to have chambers in the order "A1, B1, C1, D1", rather than "A1, A2, A3, A4".  

```{r, warning=FALSE, message=FALSE}
Sample.Info<-read.csv("../Data/Respiration/Mcap2020/Resp_Sample_Info.csv") #information on life stage, number of individuals, runs, etc.
rename <- Sample.Info$Chamber.ID
samp <- Sample.Info$Sample.ID
run <- str_sub(file.names, 10, 13) #grab run from file name
date <- str_sub(file.names, 1, str_length(file.names)-16) #grab date from file name

#load in respiration start times as a list the same length as the number of files
starttimes<-read.csv("../Data/Respiration/Mcap2020/starttimes.csv") #read in file of starting R and P phase times
rtime<-starttimes$Rtime #list respiration start times. For respiration measurements, filter as > than this time
ptime<-starttimes$Ptime #for photosynthesis, filter as < than this time
```

Run loop to extract slopes from respiration data. Subset for respiration data by filtering time >= respiration start time. 

```{r, results=FALSE, warning=FALSE, message=FALSE}
for(i in 1:length(file.names)) { # for every file in list start at the first and run this following function
  Resp.Data <-read.table(file.path(path.p,file.names[i]), skip = 72, header=T, sep=",", na.string="NA", fill = TRUE, as.is=TRUE, fileEncoding="latin1") #reads in the data files
  Resp.Data$Temp <- Resp.Data[,31] #assigns temp column
  Resp.Data$Time.Min <- seq.int(0.017, (nrow(Resp.Data))*0.25, by = 0.25) #set time in min
  Resp.Data <- Resp.Data %>% #filters data by phase (respiration only)
     filter(Time.Min > rtime[i])
  Resp.Data.N <- Resp.Data[,3:26] #subset desired columns
  
  for(j in 1:(ncol(Resp.Data.N))){
    model <- rankLocReg(
      xall=Resp.Data$Time.Min, yall=as.numeric(Resp.Data.N[, j]), 
      alpha=0.4, method="pc", verbose=TRUE) #extract slopes, percentile rank method with minimum window size of 0.4. This means that in order to fit a slope, it has to encompass at least 40% of available datapoints. 
    
    pdf(paste0("../Output/Mcap2020/Respiration/",date[i], "_",run[i],"_",rename[j],"_regression_trunc.pdf")) #generate output file names
    plot(model)
    dev.off()
    
    Resp.Rb[j,1] <- as.character(date[i]) #stores the date
    Resp.Rb[j,2] <- as.character(run[i]) #stores the run number
    Resp.Rb[j,3] <- as.character(samp[j+(i-1)*ncol(Resp.Data.N)]) #stores the sample ID
    Resp.Rb[j,4] <- as.character(rename[j]) #stores the chamber ID
    Resp.Rb[j,5:6] <- model$allRegs[i,c(4,5)] #inserts slope and intercept in the dataframe
    
  }
  Resp.R <- rbind(Resp.R, Resp.Rb) #bind final data frame
}

```

Save respiration data frames.  

```{r, results=FALSE, warning=FALSE, message=FALSE}
Resp.R <- Resp.R[-1,] #remove empty column

write.csv(Resp.R, paste0("../Output/Mcap2020/Respiration_rates_Mcap2020.csv", sep="")) #save respiration rate data

Resp.Rates <- read.csv(file = "../Output/Mcap2020/Respiration_rates_Mcap2020.csv") #read file back in so slopes don't have to be generated every time
Resp.Rates = subset(Resp.Rates, select = -c(X) ) #remove empty column

#format "run" column
Resp.Rates<-Resp.Rates %>%
  mutate(Run = str_sub(Run, 4, -1))
Resp.Rates$Run<-as.integer(Resp.Rates$Run) #format as # rather than run #, set as integer
```

Run loop to extract slopes from photosynthesis data ( time < respiration start time). 

```{r, warning=FALSE, message=FALSE}
for(i in 1:length(file.names)) { # for every file in list start at the first and run this following function
  Photo.Data <-read.table(file.path(path.p,file.names[i]), skip = 72, header=T, sep=",", na.string="NA", fill = TRUE, as.is=TRUE, fileEncoding="latin1") #reads in the data files
  Photo.Data$Temp <- Photo.Data[,31] #assign temperature column
  Photo.Data$Time.Min <- seq.int(0.017, (nrow(Photo.Data))*0.25, by = 0.25) #set time in min
  Photo.Data <- Photo.Data %>% #filters data by phase (photosynthesis only)
     filter(Time.Min <= ptime[i])
  Photo.Data.N <- Photo.Data[,3:26] #select desired columns
  
  for(j in 1:(ncol(Photo.Data.N))){
    model <- rankLocReg(
      xall=Photo.Data$Time.Min, yall=as.numeric(Photo.Data.N[, j]),
      alpha=0.4, method="pc", verbose=TRUE) #extract slopes with percentile method and required minimum window of 40% of data to extract slope
    
    pdf(paste0("../Output/Mcap2020/Photosynthesis/",date[i], "_",run[i],"_",rename[j],"_regression_trunc.pdf")) #generate output files
    plot(model)
    dev.off()
    
    Photo.Rb[j,1] <- as.character(date[i]) #stores the date
    Photo.Rb[j,2] <- as.character(run[i]) #stores the run number
    Photo.Rb[j,3] <- as.character(samp[j+(i-1)*ncol(Photo.Data.N)]) #stores the sample ID
    Photo.Rb[j,4] <- as.character(rename[j]) #stores the chamber ID
    Photo.Rb[j,5:6] <- model$allRegs[i,c(4,5)] #inserts slope and intercept in the dataframe
    
  }
  Photo.R <- rbind(Photo.R, Photo.Rb) #bind final data frame
}
```

Save photosynthesis data frames.  

```{r, warning=FALSE, message=FALSE}
Photo.R <- Photo.R[-1,] #remove empty row

write.csv(Photo.R, paste0("../Output/Mcap2020/Photosynthesis_rates_Mcap2020.csv", sep="")) #save P rates

Photo.Rates <- read.csv(file = "../Output/Mcap2020/Photosynthesis_rates_Mcap2020.csv") #import P rates
Photo.Rates = subset(Photo.Rates, select = -c(X) ) #remove empty column

#format "run" column
Photo.Rates<-Photo.Rates %>% #Format as "#" rather than "Run#"
  mutate(Run = str_sub(Run, 4, -1))
Photo.Rates$Run<-as.integer(Photo.Rates$Run) #format as integer
```

Merge P and R files with sample info for testing and manipulation.  

```{r, warning=FALSE, message=FALSE}
oxygen<-full_join(Sample.Info, Resp.Rates) #add respiration data
colnames(oxygen)[colnames(oxygen) == 'Intercept'] <- 'Resp.Intercept' #rename to specify R
colnames(oxygen)[colnames(oxygen) == 'umol.L.sec'] <- 'R.umol.L.sec' #rename to specify R

oxygen<-full_join(oxygen, Photo.Rates) #add photosynthesis data
colnames(oxygen)[colnames(oxygen) == 'Intercept'] <- 'Photo.Intercept' #rename to specify P
colnames(oxygen)[colnames(oxygen) == 'umol.L.sec'] <- 'P.umol.L.sec' #rename to specify P
```

Account for volume and substract blank values. Average blank calculated for each run for P and R phases. Display mean blank value in order of R and P phases.    
```{r, results=TRUE, warning=FALSE, message=FALSE}
#Account for chamber volume to convert from umol L-1 s-1 to umol s-1. This standardizes across water volumes (different because of coral size) and removes per Liter
oxygen$R.umol.sec <- oxygen$R.umol.L.sec * oxygen$Volume #calculate

oxygen$P.umol.sec <- oxygen$P.umol.L.sec * oxygen$Volume #calculate

blank_data <- subset(oxygen, Type == "Blank") #subset to blank data only
plot(blank_data$R.umol.sec , ylab="umol O2 sec-1") #blanks during dark phase
plot(blank_data$P.umol.sec , ylab="umol O2 sec-1") #blanks during light phase

#display mean blankvalues
mean(blank_data$R.umol.sec) #mean R phase blanks
mean(blank_data$P.umol.sec) #mean P phase blanks

resp.blnk <- aggregate(R.umol.sec ~ Date*Run, data=blank_data, mean) #calculate average blank during light for each run
colnames(resp.blnk)[colnames(resp.blnk) == 'R.umol.sec'] <- 'R.Blank.umol.sec' #rename to specify blank for R

photo.blnk <- aggregate(P.umol.sec ~ Date*Run, data=blank_data, mean) #calculate average blank during dark for each run
colnames(photo.blnk)[colnames(photo.blnk) == 'P.umol.sec'] <- 'P.Blank.umol.sec' #rename to specify blank for P

oxygen <- full_join(oxygen, resp.blnk) #add R blanks to master
oxygen <- full_join(oxygen, photo.blnk) #add P blanks to master
```

Subtract blank values to generate a "corrected" value for umol O2 sec-1.  

```{r, warning=FALSE, message=FALSE}
oxygen$R.umol.sec.corr<-oxygen$R.umol.sec-oxygen$R.Blank.umol.sec #subtract R blanks
oxygen$P.umol.sec.corr<-oxygen$P.umol.sec-oxygen$P.Blank.umol.sec #subtract P blanks
```

Normalize to biologically relevant measure. Here, normalize to number of larvae. This can be substituted or changed for larval size/volume as well.  

```{r, warning=FALSE, message=FALSE}
oxygen.bio <- oxygen %>% filter(Type == "Sample") #isolate only biological samples and drop unused factor levels 
oxygen.bio <- droplevels(oxygen.bio) #drop unused factor levels

#respiration
oxygen.bio$R.umol.org.sec <- oxygen.bio$R.umol.sec.corr/oxygen.bio$Org.Number #calculate oxygen per organism (larvae or recruit)
oxygen.bio$R.umol.org.min <- oxygen.bio$R.umol.org.sec/60 #calculate oxygen per organism per minute
oxygen.bio$R.nmol.org.min <- oxygen.bio$R.umol.org.min*1000 #calculate nanmoles 

#photosynthesis
oxygen.bio$P.umol.org.sec <- oxygen.bio$P.umol.sec.corr/oxygen.bio$Org.Number #calculate oxygen per organism (larvae or recruit)
oxygen.bio$P.umol.org.min <- oxygen.bio$P.umol.org.sec/60 #calculate oxygen per organism per minute
oxygen.bio$P.nmol.org.min <- oxygen.bio$P.umol.org.min*1000 #calculate nanmoles 
```

Save as .csv file.  
```{r, warning=FALSE, message=FALSE}
write.csv(oxygen.bio, paste0("../Output/Mcap2020/oxygen_P_R_calc.csv", sep="")) #save final file
```

This point ends the general use script for P and R rates. Below are project-specific scripts.  

## Data visualization and manipulation  

Load data from LoLinR.   
```{r, warning=FALSE, message=FALSE}
PRdata<-read.csv("../Output/Mcap2020/oxygen_P_R_calc.csv") #load data
```

Separate project specific data.  
```{r, warning=FALSE, message=FALSE}
#remove all rows of wells that did not have samples or blanks
PRdata<-PRdata[!is.na(PRdata$Type),]

#format columns
PRdata$dpf<-as.factor(PRdata$dpf)

#subset fused juvenile data frame for a different project, then save as separate file
fused_oxygen <- PRdata[which(PRdata$Lifestage=='Fused'),]
fused_oxygen <- droplevels(fused_oxygen)

#save respiration data without fused juveniles
PRdata<-PRdata[!(PRdata$Lifestage=="Fused"),]
PRdata <- droplevels(PRdata)
```

Calculate a P:R ratio.  
```{r, warning=FALSE, message=FALSE}
PRdata$ratio<-abs(PRdata$P.nmol.org.min)/abs(PRdata$R.nmol.org.min) #claculate
```

## Plotting  

Generate plots by lifestage and days post fertilization.  

### Respiration  

Generate dot plot.  

```{r, results=TRUE}
Rplot1<-ggplot(data=PRdata, aes(x=dpf, y=R.nmol.org.min, colour=Lifestage, group=interaction(Lifestage, dpf))) +
  geom_boxplot(width=0.75, position=position_dodge(0.9))+
  geom_dotplot(binaxis='y', stackdir='center', 
               position=position_dodge(0.9), dotsize=0.3, color="black")+
  theme_classic()+
  theme(text = element_text(size = 18, color="black"))+
  theme(axis.text = element_text(size = 18, color="black"))+
  theme(legend.position = "right")+
  theme(axis.title = element_text(size = 18, color="black", face="bold"))+
  ylab(expression(bold(paste("Respiration: ", mu, "mol organism"^-1, "min"^-1)))) +
  theme(legend.title=element_blank())+
  theme(legend.text = element_text(size=18))+
  theme(plot.margin=unit(c(1.5,0,0,0), "cm"))+
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 7, b = 0, l = 1))) +
  theme(axis.title.x = element_text(margin = margin(t = 7, r = 0, b = 0, l = 0))) +
  xlab("Days post fertilization"); Rplot1

ggsave("../Figures/Respiration/mcap_R.pdf", plot=Rplot1, height=8, width=10, units = c("in"), dpi=300) #output figure
```


### Photosynthesis   

```{r, results=TRUE, warning=FALSE, message=FALSE}
Pplot1<-ggplot(data=PRdata, aes(x=dpf, y=P.nmol.org.min, colour=Lifestage, group=interaction(Lifestage, dpf))) +
  geom_boxplot(width=0.75, position=position_dodge(0.9))+
  geom_dotplot(binaxis='y', stackdir='center', 
               position=position_dodge(0.9), dotsize=0.3, color="black")+
  theme_classic()+
  theme(text = element_text(size = 18, color="black"))+
  theme(axis.text = element_text(size = 18, color="black"))+
  theme(legend.position = "right")+
  theme(axis.title = element_text(size = 18, color="black", face="bold"))+
  ylab(expression(bold(paste("Photosynthesis: ", mu, "mol organism"^-1, "min"^-1)))) +
  theme(legend.title=element_blank())+
  theme(legend.text = element_text(size=18))+
  theme(plot.margin=unit(c(1.5,0,0,0), "cm"))+
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 7, b = 0, l = 1))) +
  theme(axis.title.x = element_text(margin = margin(t = 7, r = 0, b = 0, l = 0))) +
  xlab("Days post fertilization"); Pplot1

ggsave("../Figures/Respiration/mcap_P.pdf", plot=Pplot1, height=8, width=10, units = c("in"), dpi=300) #output figure
```

### P:R ratio    

```{r, results=TRUE, warning=FALSE, message=FALSE}
PRplot1<-ggplot(data=PRdata, aes(x=dpf, y=ratio, colour=Lifestage, group=interaction(Lifestage, dpf))) +
  geom_boxplot(width=0.75, position=position_dodge(0.9))+
  geom_dotplot(binaxis='y', stackdir='center', 
               position=position_dodge(0.9), dotsize=0.3, color="black")+
  theme_classic()+
  theme(text = element_text(size = 18, color="black"))+
  theme(axis.text = element_text(size = 18, color="black"))+
  theme(legend.position = "right")+
  theme(axis.title = element_text(size = 18, color="black", face="bold"))+
  ylab(expression(bold(paste("Photosynthesis:Respiration ")))) +
  theme(legend.title=element_blank())+
  theme(legend.text = element_text(size=18))+
  theme(plot.margin=unit(c(1.5,0,0,0), "cm"))+
  ylim(-10,10)+
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 7, b = 0, l = 1))) +
  theme(axis.title.x = element_text(margin = margin(t = 7, r = 0, b = 0, l = 0))) +
  xlab("Days post fertilization"); PRplot1

ggsave("../Figures/Respiration/mcap_PR.pdf", plot=PRplot1, height=8, width=10, units = c("in"), dpi=300) #output figure
```

## Analysis  

### Respiration  

Build linear mixed effect model and examine for Respiration.     
```{r, results=TRUE, warning=FALSE, message=FALSE}
Rmodel1<-lmer(R.nmol.org.min~Lifestage*dpf + (1|Run/dpf), data=PRdata) #run nested within day
summary(Rmodel1)
anova(Rmodel1, type="II")
```

Check assumptions of model for residual normality and variance.  

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(Rmodel1))
hist(residuals(Rmodel1))
leveneTest(residuals(Rmodel1)~Lifestage * dpf, data=PRdata)
```

### Photosynthesis  

Build linear mixed effect model and examine for Photosynthesis  
```{r, results=TRUE, warning=FALSE, message=FALSE}
Pmodel1<-lmer(P.nmol.org.min~Lifestage*dpf + (1|Run/dpf), data=PRdata)
summary(Pmodel1)
anova(Pmodel1, type="II")
```
 
Check assumptions of model for residual normality and variance.  

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(Pmodel1))
hist(residuals(Pmodel1))
leveneTest(residuals(Pmodel1)~Lifestage * dpf, data=PRdata)
```

### P:R  
 
Build linear mixed effect model and examine for P:R  
```{r, results=TRUE, warning=FALSE, message=FALSE}
PRdata_ratio_mod<-PRdata[-c(14, 15),] #remove outliers, seen with PRdata and removed
PRmodel1<-lmer(ratio~Lifestage*dpf + (1|Run/dpf), data=PRdata_ratio_mod)
summary(PRmodel1)
anova(PRmodel1, type="II")
```

Check assumptions of model for residual normality and variance. Need to check for outliers.    

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(PRmodel1))
hist(residuals(PRmodel1))
leveneTest(residuals(PRmodel1)~Lifestage * dpf, data=PRdata_ratio_mod)
```


### Correlations   

Correlation between P and R.   
```{r, results=TRUE, warning=FALSE, message=FALSE}
cor.test(PRdata$P.nmol.org.min, PRdata$R.nmol.org.min, method=c("spearman")) #non parametric correlation
```

