---
title: M. capitata 2020 metabolomics analysis
author: "AS Huffmyer"
date: '2021'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---
Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("vegan" %in% rownames(installed.packages()) == 'FALSE') install.packages('vegan') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("factoextra" %in% rownames(installed.packages()) == 'FALSE') install.packages('factoextra') 
if ("ggfortify" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggfortify') 
if ("naniar" %in% rownames(installed.packages()) == 'FALSE') install.packages('naniar') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
if ("mixOmics" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install("mixOmics") 

#load packages
library("ggplot2")
library('vegan')
library('factoextra')
library('ggfortify')
library('naniar')
library('cowplot')
library("mixOmics")
library("tidyverse")

```

## Generalized Work Flow

1. Data processing and exploration    
 - Import Data (positive peaks, negative peaks)  
 - Normalize to sample median   
 - Preliminary visualization of ion counts   
 - Selection of polarity and generation of final data frame ("selected_metabolites.csv")  
 - Check for polarity correlations  
 - Zero inflate values by 1000  
 - Log normalization of all ion counts of selected dataset  
 
2. Preliminary PCA and PERMANOVA statistics  
3. Preliminary PLS-DA  
4. Validation of preliminary PLS-DA  
5. sPLS-DA from PLS-DA validation  
6. sPLS-DA model validation   
7. Extraction of significant loadings  

# Data processing and exploration  
## Positive Polarity  

Load data
```{r}
data<-read.csv("Mcap2020/Data/Metabolomics/Peaks_Pos_11292020_Rutgers.csv", sep=",", na.strings=NA, header=TRUE)

#rename columns to be useful
data<- data %>% 
  rename(
    Larvae2_rep1 = Larvae.2_50uL_1,
    Larvae2_rep2 = Larvae.2_50uL_2,
    Larvae4_rep1 = Larvae.4_50uL_1,
    Larvae4_rep2 =  Larvae.4_50uL_2,
    Larvae6_rep1 = Larvae.6_15.20.Individuals_1,
    Larvae6_rep2 = Larvae.6_15.20.Individuals_2 
    )

```

Conduct normalization of samples by normalizing each metabolite ion count to the median ion count across metabolites within that sample.  

```{r}
data<- data %>%
  mutate(Larvae2_rep1_norm = Larvae2_rep1 / median(Larvae2_rep1), 
         Larvae2_rep2_norm = Larvae2_rep2 / median(Larvae2_rep2), 
         Larvae4_rep1_norm = Larvae4_rep1 / median(Larvae4_rep1), 
         Larvae4_rep2_norm = Larvae4_rep2 / median(Larvae4_rep2),
         Larvae6_rep1_norm = Larvae6_rep1 / median(Larvae6_rep1),
         Larvae6_rep2_norm = Larvae6_rep2 / median(Larvae6_rep2))
```


Mutate the data to be in long rather than wide format.  

```{r}
data <- data %>% 
  gather(Sample, IonCounts, 23:28)

data <- data[c(9, 23, 24)]
```

Add lifestage column for categorizing  

```{r}
data<-data%>%
  separate(Sample, sep="_", c("Lifestage", "Rep"), remove=FALSE)

data$Lifestage<-as.factor(data$Lifestage)
```

Create plot showing normalized ion counts (count / sample median) for each lifestage.  

```{r}
#make summary table
data_plotting<-plyr::ddply(data, c("Lifestage", "compound"), summarise,
                N    = length(IonCounts[!is.na(IonCounts)]),
                mean = mean(IonCounts, na.rm=TRUE),
                sd   = sd(IonCounts, na.rm=TRUE),
                se   = sd / sqrt(N)
)
#write summary file
data_summary<-data_plotting[c(1,2,4)]

#data_summary%>%
  #spread(Lifestage, mean) %>%
  #write_csv("Mcap2020/Output/Metabolomics/mean_norm_counts_lifestage_pos.csv")

metab_compounds<-ggplot(data_plotting, aes(y=mean, x=compound, fill=Lifestage, group=Lifestage)) + 
  geom_bar(stat="identity", position=position_dodge())+ #change to change boxplot width
  theme_classic()+
  xlab("Metabolites")+
  ylab("Normalized Ion Count")+
 # ylim(-1,100)+
  theme(
    legend.position="right", 
    panel.border=element_rect(colour="black", fill=NA, size=1),
    strip.text.x = element_text(size = 12, color = "black", face = "bold"),
    text = element_text(size = 16, color="black"), 
    axis.text.y = element_text(size=14, color="black"), 
    axis.text.x = element_text(size=12, color="black", angle=90, vjust = 1, hjust=1), 
    axis.title = element_text(face="bold"));metab_compounds

ggsave("Mcap2020/Figures/Metabolomics/compounds_fullscale_pos.pdf", plot=metab_compounds, dpi=300, width=25, height=8, units="in")

metabl_short<-metab_compounds + ylim(0,100)

ggsave("Mcap2020/Figures/Metabolomics/compounds_shortscale_pos.pdf", plot=metabl_short, dpi=300, width=25, height=8, units="in")

metabl_shortest<-metab_compounds + ylim(0,20)

ggsave("Mcap2020/Figures/Metabolomics/compounds_shortestscale_pos.pdf", plot=metabl_shortest, dpi=300, width=25, height=8, units="in")

```

Run PCA to view differences in life stages.  

Convert data to wide format.

```{r}
data_pca<-data%>%
  spread(compound, IonCounts)
```

Plot PCA with lifestage in colors.  

```{r}
scaled.pca<-prcomp(data_pca[c(4:81)],scale=TRUE, center=TRUE) 

pcaPlot1<-ggplot2::autoplot(scaled.pca, data=data_pca, frame.colour="Lifestage", loadings=FALSE,  colour="Lifestage", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=FALSE, loadings.label.size=5, loadings.label.vjust=-1, size=5) + 
  scale_colour_manual(values=c("orange", "darkgray", "purple")) +
  scale_fill_manual(values=c("orange", "darkgray", "purple")) + 
  theme_classic()+
  xlim(-0.75,0.75)+
  ylim(-0.75, 0.75)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pcaPlot1

#pcaPlot1$layers[[2]]$aes_params$size <- 1

ggsave("Mcap2020/Figures/Metabolomics/PCA_pos.pdf", plot=pcaPlot1, height=8, width=10, units = c("in"), dpi=300) #output figure
```

Read in .csv of normalized counts with trend assigned.  

Plot compounds grouping by increasing over ontogeny, decreasing, or no apparent trend.  

Add a column to distinguish increasing, decreasing, or variable trends.  

```{r}
#assign trends 
trends<-data_summary%>%
  spread(Lifestage, mean)%>%
  mutate(Trend=ifelse(Larvae2>Larvae4 & Larvae4>Larvae6, "Decreasing", 
                      ifelse(Larvae2<Larvae4 & Larvae4<Larvae6, "Increasing", "Variable")))
```


```{r}
trends<-trends%>%
  gather(Lifestage, IonCounts, 2:4)
  
metab_trends<-ggplot(trends, aes(y=IonCounts, x=compound, fill=Lifestage, group=Lifestage)) + 
  facet_wrap(~Trend, scales = "free")+
  geom_bar(stat="identity", position=position_dodge())+ #change to change boxplot width
  theme_classic()+
  xlab("Metabolites")+
  ylab("Normalized Ion Count")+
  theme(
    legend.position="right", 
    panel.border=element_rect(colour="black", fill=NA, size=1),
    strip.text.x = element_text(size = 12, color = "black", face = "bold"),
    text = element_text(size = 16, color="black"), 
    axis.text.y = element_text(size=14, color="black"), 
    axis.text.x = element_text(size=12, color="black", angle=90, vjust = 1, hjust=1), 
    axis.title = element_text(face="bold"));metab_trends

ggsave("Mcap2020/Figures/Metabolomics/trends_fullscale.pdf", plot=metab_trends, dpi=300, width=35, height=8, units="in")

trend_short<-metab_trends + ylim(0,100)

ggsave("Mcap2020/Figures/Metabolomics/trends_shortscale.pdf", plot=trend_short, dpi=300, width=35, height=8, units="in")

trend_shortest<-metab_trends + ylim(0,20)

ggsave("Mcap2020/Figures/Metabolomics/trends_shortestscale.pdf", plot=trend_shortest, dpi=300, width=35, height=8, units="in")
```



## Negative Polarity  

Load data
```{r}
data_neg<-read.csv("Mcap2020/Data/Metabolomics/Peaks_Neg_11292020_Rutgers.csv", sep=",", na.strings=NA, header=TRUE)

#rename columns to be useful
data_neg<- data_neg %>% 
  rename(
    Larvae2_rep1 = Larvae.2_50uL_1,
    Larvae2_rep2 = Larvae.2_50uL_2,
    Larvae4_rep1 = Larvae.4_50uL_1,
    Larvae4_rep2 = Larvae.4_50uL_2,
    Larvae6_rep1 = Larvae.6_15.20.Individuals_1,
    Larvae6_rep2 = Larvae.6_15.20.Individuals_2 
    )

```


Conduct normalization of samples by normalizing each metabolite ion count to the median ion count across metabolites within that sample.  

```{r}
data_neg<- data_neg %>%
  mutate(Larvae2_rep1_norm = Larvae2_rep1 / median(Larvae2_rep1), 
         Larvae2_rep2_norm = Larvae2_rep2 / median(Larvae2_rep2), 
         Larvae4_rep1_norm = Larvae4_rep1 / median(Larvae4_rep1), 
         Larvae4_rep2_norm = Larvae4_rep2 / median(Larvae4_rep2),
         Larvae6_rep1_norm = Larvae6_rep1 / median(Larvae6_rep1),
         Larvae6_rep2_norm = Larvae6_rep2 / median(Larvae6_rep2))
```


Mutate the data to be in long rather than wide format.  

```{r}
data_neg <- data_neg %>% 
  gather(Sample, IonCounts, 23:28)

data_neg <- data_neg[c(9, 23, 24)]
```

Add lifestage column for categorizing  

```{r}
data_neg<-data_neg%>%
  separate(Sample, sep="_", c("Lifestage", "Rep"), remove=FALSE)

data_neg$Lifestage<-as.factor(data_neg$Lifestage)
```

Create plot showing normalized ion counts (count / sample median) for each lifestage.  

```{r}
#make summary table
data_plotting_neg<-plyr::ddply(data_neg, c("Lifestage", "compound"), summarise,
                N    = length(IonCounts[!is.na(IonCounts)]),
                mean = mean(IonCounts, na.rm=TRUE),
                sd   = sd(IonCounts, na.rm=TRUE),
                se   = sd / sqrt(N)
)
#write summary file
data_summary_neg<-data_plotting_neg[c(1,2,4)]

#data_summary_neg%>%
  #spread(Lifestage, mean) %>%
  #write_csv("Mcap2020/Output/Metabolomics/mean_norm_counts_lifestage_neg.csv")

metab_compounds_neg<-ggplot(data_plotting_neg, aes(y=mean, x=compound, fill=Lifestage, group=Lifestage)) + 
  geom_bar(stat="identity", position=position_dodge())+ #change to change boxplot width
  theme_classic()+
  xlab("Metabolites")+
  ylab("Normalized Ion Count")+
 # ylim(-1,100)+
  theme(
    legend.position="right", 
    panel.border=element_rect(colour="black", fill=NA, size=1),
    strip.text.x = element_text(size = 12, color = "black", face = "bold"),
    text = element_text(size = 16, color="black"), 
    axis.text.y = element_text(size=14, color="black"), 
    axis.text.x = element_text(size=12, color="black", angle=90, vjust = 1, hjust=1), 
    axis.title = element_text(face="bold"));metab_compounds_neg

ggsave("Mcap2020/Figures/Metabolomics/compounds_fullscale_neg.pdf", plot=metab_compounds_neg, dpi=300, width=25, height=8, units="in")

metabl_short_neg<-metab_compounds_neg + ylim(0,100)

ggsave("Mcap2020/Figures/Metabolomics/compounds_shortscale_neg.pdf", plot=metabl_short_neg, dpi=300, width=25, height=8, units="in")

metabl_shortest_neg<-metab_compounds_neg + ylim(0,20)

ggsave("Mcap2020/Figures/Metabolomics/compounds_shortestscale_neg.pdf", plot=metabl_shortest_neg, dpi=300, width=25, height=8, units="in")

```


Run PCA to view differences in life stages.  

Convert data to wide format.

```{r}
data_pca_neg<-data_neg%>%
  spread(compound, IonCounts)
```
 
Plot PCA with lifestage in colors.  

```{r}
scaled.pca.neg<-prcomp(data_pca_neg[c(4:76)],scale=TRUE, center=TRUE) 

pcaPlot1_neg<-ggplot2::autoplot(scaled.pca.neg, data=data_pca_neg, frame.colour="Lifestage", loadings=FALSE,  colour="Lifestage", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=FALSE, loadings.label.size=5, loadings.label.vjust=-1, size=5) + 
  scale_colour_manual(values=c("orange", "darkgray", "purple")) +
  scale_fill_manual(values=c("orange", "darkgray", "purple")) + 
  theme_classic()+
  xlim(-0.75,0.75)+
  ylim(-0.75, 0.75)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pcaPlot1_neg

#pcaPlot1$layers[[2]]$aes_params$size <- 1

ggsave("Mcap2020/Figures/Metabolomics/PCA_neg.pdf", plot=pcaPlot1_neg, height=8, width=10, units = c("in"), dpi=300) #output figure
```

Read in .csv of normalized counts with trend assigned.  

Plot compounds grouping by increasing over ontogeny, decreasing, or no apparent trend.  

Add a column to distinguish increasing, decreasing, or variable trends.  
```{r}
#assign trends to neg dataset
trends_neg<-data_summary_neg%>%
  spread(Lifestage, mean)%>%
  mutate(Trend=ifelse(Larvae2>Larvae4 & Larvae4>Larvae6, "Decreasing", 
                      ifelse(Larvae2<Larvae4 & Larvae4<Larvae6, "Increasing", "Variable")))
```

```{r}
trends_neg<-trends_neg%>%
  gather(Lifestage, IonCounts, 2:4)
  
metab_trends_neg<-ggplot(trends_neg, aes(y=IonCounts, x=compound, fill=Lifestage, group=Lifestage)) + 
  facet_wrap(~Trend, scales = "free")+
  geom_bar(stat="identity", position=position_dodge())+ #change to change boxplot width
  theme_classic()+
  xlab("Metabolites")+
  ylab("Normalized Ion Count")+
  theme(
    legend.position="right", 
    panel.border=element_rect(colour="black", fill=NA, size=1),
    strip.text.x = element_text(size = 12, color = "black", face = "bold"),
    text = element_text(size = 16, color="black"), 
    axis.text.y = element_text(size=14, color="black"), 
    axis.text.x = element_text(size=12, color="black", angle=90, vjust = 1, hjust=1), 
    axis.title = element_text(face="bold"));metab_trends_neg

ggsave("Mcap2020/Figures/Metabolomics/trends_fullscale_neg.pdf", plot=metab_trends_neg, dpi=300, width=35, height=8, units="in")

trend_short_neg<-metab_trends_neg + ylim(0,100)

ggsave("Mcap2020/Figures/Metabolomics/trends_shortscale_neg.pdf", plot=trend_short_neg, dpi=300, width=35, height=8, units="in")

trend_shortest_neg<-metab_trends_neg + ylim(0,20)

ggsave("Mcap2020/Figures/Metabolomics/trends_shortestscale_neg.pdf", plot=trend_shortest_neg, dpi=300, width=35, height=8, units="in")
```

Generate a dataframe with normalized counts for positive and negative polarity datasets.  

```{r}
#merge two datasets such that mean in negative = "negative", and mean from positive = "positive". If not present, NA. 
data_summary_neg<-data_summary_neg%>%
    rename(Negative = mean)

data_summary<-data_summary%>%
    rename(Positive = mean)

#view normalized negative dataset
head(data_summary_neg)

head(data_summary)

#join datasets together
normalized_counts_pos_neg<-full_join(data_summary, data_summary_neg)

write_csv(normalized_counts_pos_neg, file="Mcap2020/Output/Metabolomics/normalized_counts_pos_neg.csv")
```

# Selecting metabolite polarity and generating final dataframe   

Use metabolite selection file to select polarity by averaging peaks across life stages for pos and neg peaks and choosing the strongest peaks.    

```{r}
#datan is positive and data_neg are negative dataframes 
data<-data%>%
  mutate(Polarity="Positive")

data_neg<-data_neg%>%
  mutate(Polarity="Negative")

data_full<-full_join(data, data_neg) #combine full positive and negative datasets

data_full<-data_full%>%
  spread(Polarity, IonCounts)
```

Check for correlations between metabolites that have both positive and negative representation. We would expect that metabolites with higher ion counts will show higher counts across polarities.  

```{r}
correlation<-data_full

#only use complete observations
correlation<-correlation[complete.cases(correlation), ]
```

Display correlations for the metabolites with shared representation.  

```{r}
corrplots<-ggplot(data=correlation, aes(x=log(1+Positive), y=log(1+Negative)))+
  geom_point()+
  facet_wrap(~compound, scales="free")+
  geom_smooth(method='lm', color='red')+
  theme_classic()+
  theme(text=element_text(size=14, face="bold"))
corrplots

ggsave("Mcap2020/Figures/Metabolomics/Correlation_plot.pdf", corrplots, width=20, height=16)
```

Relationship between positive and negative polarities are generally positive. Some of the relationships are more strong for certain metabolites and more variable for others.  

```{r}
#summarize across life stages and reps
polarity<-data_full%>%
  gather("Polarity", "IonCounts", 5:6)%>%
  group_by(Polarity, compound)%>%
  summarise(mean=mean(IonCounts))%>% #summarize across samples (lifestages)
  arrange(compound)%>% #list by compound
  replace_na(list(mean=0)) #replace NA with 0 to allow us to calculate higher values between polarities 

#add column of selected polarity based on stronger signal average
polarity<-polarity%>%
  spread(Polarity, mean)%>%
  mutate(Selection = ifelse(Negative > Positive, "Negative",
                     ifelse(Positive > Negative, "Positive", NA)))
  
```

Use the selected polarity key to select data for analysis.    

```{r}
#match selection column back onto original dataframe 
data_full$Selection<-polarity$Selection[match(data_full$compound, polarity$compound)] 

#pull out pos data
pos_peaks<-data_full%>%
  filter(Selection=="Positive")%>%
  dplyr::select(Lifestage, Rep, compound, Positive, Selection)%>%
  rename(value=Positive)
  
#pull out neg data
neg_peaks<-data_full%>%
  filter(Selection=="Negative")%>%
  dplyr::select(Lifestage, Rep, compound, Negative, Selection)%>%
  rename(value=Negative)

#rbind them together to generate a dataframe of positive and negative selected compounds
peaks_selected<-rbind(pos_peaks, neg_peaks)

peaks_selected<-peaks_selected %>%
  dplyr::select(Lifestage, Rep, compound, value) 
```

Add 1 to all values to account for 0 values and log transform.  
```{r}
hist(peaks_selected$value)

peaks_selected$log_norm_counts<-log10(1+peaks_selected$value)
hist(log(peaks_selected$log_norm_counts))
```

Output file to .csv.   

```{r}
peaks_selected%>%
  rename(norm_counts=value)%>%
  write_csv("Mcap2020/Output/Metabolomics/selected_metabolites.csv")
```

# Preliminary PCA of combined polarities  

Run PCA to view differences in life stages.   

Convert data to wide format. 

```{r}
data_pca_full<-peaks_selected%>%
  dplyr::select(Lifestage, Rep, compound, value)%>%
  spread(compound, value)
```

Plot PCA and view scree plot with lifestage in colors.   

```{r}
scaled.pca<-prcomp(data_pca_full[c(3:123)],scale=TRUE, center=TRUE) 

fviz_eig(scaled.pca)

pcaPlot_full<-ggplot2::autoplot(scaled.pca, data=data_pca_full, frame.colour="Lifestage", loadings=FALSE,  colour="Lifestage", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=FALSE, loadings.label.size=5, loadings.label.vjust=-1, size=5) + 
  scale_colour_manual(values=c("orange", "darkgray", "purple")) +
  scale_fill_manual(values=c("orange", "darkgray", "purple")) + 
  theme_classic()+
  xlim(-0.75,0.75)+
  ylim(-0.75, 0.75)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pcaPlot_full

ggsave("Mcap2020/Figures/Metabolomics/PCA_full.pdf", plot=pcaPlot_full, height=8, width=10, units = c("in"), dpi=300) #output figure
```
 
Show a panel of PCA by positive, negative, and combined datasets.   
```{r}
pca_pos<-pcaPlot1+theme(legend.position="none")
pca_neg<-pcaPlot1_neg+theme(legend.position="none")

pcaplots<-plot_grid(pca_pos, pca_neg, pcaPlot_full, labels = c("Positive", "Negative", "Combined"), ncol=3, nrow=1, rel_widths = c(0.8, 0.8, 1), label_size = 20, label_x = 0.1);pcaplots

ggsave(filename="Mcap2020/Figures/Metabolomics/All_PCAs.pdf", plot=pcaplots, dpi=300, width=20, height=5, units="in")
```  


# Preliminary PLS-DA 

Load data if starting from this point.   

```{r}
metabolites <- read.csv("Mcap2020/Output/Metabolomics/selected_metabolites.csv")
```

log_norm_count indicates normalized ion counts that are log transformed for analysis.   

Scripts from K Wong for preliminary PLS-DA.  

Generate data frames for PLS-DA.  
```{r, results=FALSE}
#https://mixomicsteam.github.io/Bookdown/plsda.html

#assigning datasets 
X <- metabolites%>%
  dplyr::select(Lifestage, Rep, compound, log_norm_counts)%>%
  spread(compound, log_norm_counts) #generate spread data frame

Y <- as.factor(X$Lifestage[1:6]) #select treatment names

X<-X[3:123] #pull only data columns

dim(X) ## display number of samples and metabolite features
```

Generate PLSDA.   
```{r}
# PLSDA without variable selection
MyResult.plsda <- plsda(X,Y) # 1 Run the method
#plotIndiv(MyResult.plsda)    # 2 Plot the samples
#plotVar(MyResult.plsda, cutoff = 0.7)    

plotIndiv(MyResult.plsda, ind.names = FALSE, legend=TRUE,ellipse = TRUE, title="PLS-DA")
```

Plot top 50 metabolite loadings contributing to 2 components. Here, comp1 separates larvae from recruit and comp2 separates larval 1 and larval 2.  
```{r, message=FALSE}
MyResult.plsda2 <- plsda(X,Y, ncomp=2) #number of components is classes-1 (3 classes of lifestages - 1 = 2)
#selectVar(MyResult.plsda2, comp=1)$value

plotLoadings(MyResult.plsda2, comp = 1, contrib = 'max', method = 'median', ndisplay = 50) #top 50 metabolites contributing to variation on component 1
plotLoadings(MyResult.plsda2, comp = 2, contrib = 'max', method = 'median', ndisplay = 50) #top 50 metabolites contributing to variation on component 2
```

# Validation of preliminary PLS-DA 

Validate PLS-DA.   
```{r}
#set.seed(30) 
#MyPerf.plsda <- perf(MyResult.plsda2, validation = "Mfold", folds=2, #fold #? 
                     #progressBar = TRUE, nrepeat = 50) # we suggest nrepeat = 50

MyPerf.plsda <- perf(MyResult.plsda2, validation = "loo") #using Loo correction since the number of groups I have is low, once more samples are available I can try using the Mfold validation

plot(MyPerf.plsda, col = color.mixo(5:7), sd = TRUE, legend.position = "horizontal")
```

NOT WORKING: 
```{r}
# Assessing optimal number of variables to keep 
list.keepX <- c(5:10,  seq(20, 100, 10))
list.keepX # to output the grid of values tested

set.seed(30) # for reproducbility in this vignette, otherwise increase nrepeat
tune.splsda.srbct <- tune.splsda(X, Y, ncomp = 4, # we suggest to push ncomp a bit more, e.g. 4
                                 validation = 'loo',
                                 measure = "BER", test.keepX = list.keepX)   # we suggest nrepeat = 50

error <- tune.splsda.srbct$error.rate
ncomp <- tune.splsda.srbct$choice.ncomp$ncomp # optimal number of components based on t-tests on the error rate
ncomp

###ncomp coming back as null for my data - too few groups? 

select.keepX <- tune.splsda.srbct$choice.keepX[1:ncomp]  # optimal number of variables to select
select.keepX

plot(tune.splsda.srbct, col = color.jet(ncomp))

```

# sPLS-DA from PLS-DA validation 

```{r}
#MyResult.splsda.final <- splsda(X, Y, ncomp = ncomp, keepX =select.keepX) - original use this once validation works 

MyResult.splsda.final <- splsda(X, Y, ncomp = 2, keepX = c(40,40)) #placeholder numbers for now until validation works

plotIndiv(MyResult.splsda.final, ind.names = FALSE, legend=TRUE,
          ellipse = TRUE, title="sPLS-DA - FINAL")

#selectVar(MyResult.splsda.final, comp=1)$value
#selectVar(MyResult.splsda.final, comp=2)$value

plotLoadings(MyResult.splsda.final, comp = 1, contrib = 'max', method = 'median')
plotLoadings(MyResult.splsda.final, comp = 2, contrib = 'max', method = 'median')

```

# sPLS-DA model validation  

```{r}
# Auroc plot of sPLS-DA
auc.splsda = auroc(MyResult.splsda.final, roc.comp = ncomp)

## Performance validation of final spls-da
perf.splsda <- perf(MyResult.splsda.final, validation = "loo", auc = TRUE) 

#Extracting Performance parameters
perf.splsda$auc
perf.splsda$error.rate
plot(perf.splsda) #Plotting Error rates

#Determining feature stability on each component 
par(mfrow=c(1,2))
plot(perf.splsda$features$stable[[1]], type = 'h', ylab = 'Stability', 
     xlab = 'Features', main = 'Comp 1', las =2)
plot(perf.splsda$features$stable[[2]], type = 'h', ylab = 'Stability', 
     xlab = 'Features', main = 'Comp 2', las =2)
#add more comps later on with more groups

```

# Extraction of significant loadings 

Select the stable features in each component. These are the values with the highest value from the comp plots above.    
```{r}
# here we match the selected variables to the stable features
ind.match1 = match(selectVar(MyResult.splsda.final, comp = 1)$name, 
                  names(perf.splsda$features$stable[[1]]))
ind.match2 = match(selectVar(MyResult.splsda.final, comp = 2)$name, 
                  names(perf.splsda$features$stable[[2]]))
#ind.match3 = match(selectVar(MyResult.splsda.final, comp = 3)$name, 
                  #names(perf.splsda$features$stable[[3]]))
#ind.match4 = match(selectVar(MyResult.splsda.final, comp = 4)$name, 
                  #names(perf.splsda$features$stable[[4]]))

#extract the frequency of selection of those selected variables
Freq1 = as.numeric(perf.splsda$features$stable[[1]][ind.match1])
Freq2 = as.numeric(perf.splsda$features$stable[[2]][ind.match2])
#Freq3 = as.numeric(perf.splsda$features$stable[[3]][ind.match3])
#Freq4 = as.numeric(perf.splsda$features$stable[[4]][ind.match4])

#Extracting the selected metabolites for each component 
comp1.select.metabolites <- data.frame(selectVar(MyResult.splsda.final, comp = 1)$value, Freq1)
comp2.select.metabolites <- data.frame(selectVar(MyResult.splsda.final, comp = 2)$value, Freq2)
#comp3.select.metabolites <- data.frame(selectVar(MyResult.splsda.final, comp = 3)$value, Freq3)
#comp4.select.metabolites <- data.frame(selectVar(MyResult.splsda.final, comp = 4)$value, Freq4)

head(comp1.select.metabolites)
head(comp2.select.metabolites)
#head(comp3.select.metabolites)
#head(comp4.select.metabolites)

```