---
title: Montipora capitata 2020 metabolomics WGCNA analysis
author: "AS Huffmyer"
date: '2021'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

options(stringsAsFactors = FALSE) #Set Strings to character
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("vegan" %in% rownames(installed.packages()) == 'FALSE') install.packages('vegan') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("factoextra" %in% rownames(installed.packages()) == 'FALSE') install.packages('factoextra') 
if ("ggfortify" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggfortify') 
if ("naniar" %in% rownames(installed.packages()) == 'FALSE') install.packages('naniar') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
if ("mixOmics" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install("mixOmics") 
if ("RVAideMemoire" %in% rownames(installed.packages()) == 'FALSE') install.packages('RVAideMemoire') 
if ("VennDiagram" %in% rownames(installed.packages()) == 'FALSE') install.packages('VennDiagram') 
if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom') 

#load packages
library("ggplot2")
library('vegan')
library('factoextra')
library('ggfortify')
library('naniar')
library('cowplot')
library("mixOmics")
library("tidyverse")
library("RVAideMemoire")
library("VennDiagram")
library("broom")

if ("genefilter" %in% rownames(installed.packages()) == 'FALSE') install.packages('genefilter') 
if ("DESeq2" %in% rownames(installed.packages()) == 'FALSE') install.packages('DESeq2') 
if ("RColorBrewer" %in% rownames(installed.packages()) == 'FALSE') install.packages('RColorBrewer') 
if ("WGCNA" %in% rownames(installed.packages()) == 'FALSE') install.packages('WGCNA') 
if ("flashClust" %in% rownames(installed.packages()) == 'FALSE') install.packages('flashClust') 
if ("gridExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('gridExtra') 
if ("ComplexHeatmap" %in% rownames(installed.packages()) == 'FALSE') install.packages('ComplexHeatmap') 
if ("goseq" %in% rownames(installed.packages()) == 'FALSE') install.packages('goseq') 
if ("dplyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('dplyr') 
if ("clusterProfiler" %in% rownames(installed.packages()) == 'FALSE') install.packages('clusterProfiler') 
if ("pheatmap" %in% rownames(installed.packages()) == 'FALSE') install.packages('pheatmap') 
if ("magrittr" %in% rownames(installed.packages()) == 'FALSE') install.packages('magrittr') 

library("genefilter")
library("DESeq2")
library("RColorBrewer")
library("WGCNA")
library("flashClust")
library("gridExtra")
library("ComplexHeatmap")
library("goseq")
library("dplyr")
library("clusterProfiler")
library("pheatmap")
library("magrittr")

```

This script uses WGCNA analysis on metabolomics data set to explore whether we can detect "modules" of genes that are correlated with lifestage as a comparison to PLS-DA and multivariate analyses used in metabolomics.Rmd script.  

# Load data  

Read in data file and create metadata with sample information as well as a count matrix with metabolite counts. Data will be read in from "selected metabolites" dataset that has already undergone polarity selection and median-centered with log normalization.  

```{r}
data<-read.csv("Mcap2020/Output/Metabolomics/selected_metabolites.csv")

metadata<-data%>%
  select(Lifestage, Rep, HPF, Development)%>%
  mutate(Sample=paste(Lifestage, Rep, sep="-"))

data<-data%>%
  mutate(Sample=paste(Lifestage, Rep, sep="-"))%>%
  select(Sample, compound, norm_counts)
```

Check that there are no metabolites with 0 counts for all samples. Should return TRUE.  

```{r}
rowSums(dplyr::count(data)) > 0
```

